### 2023-11-17 pldi24/paper.tex
Cut a page

%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0

\documentclass[acmsmall,screen,review]{acmart}

\usepackage{syntax}
\renewcommand{\syntleft}{\normalfont\itshape}
\renewcommand{\syntright}{\normalfont\itshape}

\usepackage{prftree}

\usepackage{listings}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{fancyvrb}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
%    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}

\newcounter{todos}
\newcommand{\TODO}[1]{{
  \stepcounter{todos}
  \begin{center}\large{\textcolor{red}{\textbf{TODO \arabic{todos}:} #1}}\end{center}
}}
\newcommand{\sorry}{\textcolor{red}{\textbf{sorry}}}

\newcommand{\todo}[1]{\stepcounter{todos} \textcolor{red}{TODO \arabic{todos}: #1}}

% Math fonts
\newcommand{\mc}[1]{\ensuremath{\mathcal{#1}}}
\newcommand{\mb}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\ms}[1]{\ensuremath{\mathsf{#1}}}

% Math
\newcommand{\nats}{\mathbb{N}}

% Syntax atoms
\newcommand{\lbl}[1]{{`#1}}
\newcommand{\lto}{\Rightarrow}
\newcommand{\ctt}{\ms{tt}}
\newcommand{\cff}{\ms{ff}}

% Syntax
\newcommand{\letexpr}[3]{\ensuremath{\ms{let}\;#1 = #2\;\ms{in}\;#3}}
\newcommand{\letstmt}[3]{\ensuremath{\ms{let}\;#1 = #2; #3}}
\newcommand{\brb}[2]{\ms{br}\;#1\;#2}
\newcommand{\lbrb}[2]{\brb{\lbl{#1}}{#2}}
\newcommand{\ite}[3]{\ms{if}\;#1\;\{#2\}\;\ms{else}\;\{#3\}}
\newcommand{\ewhere}[2]{\ms{then}\;#1\;\ms{where}\;#2}
\newcommand{\where}[2]{#1\;\ms{where}\;#2}
\newcommand{\wbranch}[3]{#1(#2) \lto #3}
\newcommand{\lwbranch}[3]{\wbranch{\lbl{#1}}{#2}{#3}}
\newcommand{\bsplice}[3]{#1(#2)\;\{#3\}}
%\newcommand{\lbsplice}[3]{\bsplice{\lbl{#1}}{#2}{#3}}
\newcommand{\csplits}[3]{#1 \mapsto #2;#3}
\newcommand{\cwk}[2]{#1 \mapsto #2}
\newcommand{\lwk}[2]{#1 \rightsquigarrow #2}
\newcommand{\tlin}[2]{#2 \subseteq \ms{lin}(#1)}
\newcommand{\ltlin}[3]{#3 \subseteq \ms{lin}(#1) \cap #2}
\newcommand{\thyp}[3]{#1: {#2}^{#3}}
\newcommand{\lhyp}[3]{#1[#2](#3)}
\newcommand{\llhyp}[3]{\lhyp{\lbl{#1}}{#2}{#3}}
\newcommand{\rle}[1]{{\scriptsize\textsf{#1}}}
\newcommand{\taff}{{\{\ms{a}\}}}
\newcommand{\trel}{{\{\ms{r}\}}}
\newcommand{\tint}{{\{\ms{a}, \ms{r}\}}}
\newcommand{\hasty}[5]{#1 \vdash_{#2} #3: {#4}^{#5}}
\newcommand{\haslb}[3]{#1 \vdash #2 \rhd #3}
\newcommand{\lhaslb}[3]{#1 \vdash #2 \rhd #3}
\newcommand{\issubst}[3]{#1: #2 \mapsto #3}
\newcommand{\lbsubst}[3]{#1: #2 \rightsquigarrow #3}
\newcommand{\exprletsubst}[2]{{#1};{#2}}
\newcommand{\stmtletsubst}[2]{{#1};{#2}}
\newcommand{\mhole}[1]{{#1}^?}
\newcommand{\lhole}[1]{?#1}
\newcommand{\mhasty}[6]{#1;#2 \vdash_{#3} #4: {#5}^{#6}}
\newcommand{\mhaslb}[4]{#1;#2 \vdash #3 \rhd #4}
\newcommand{\mlhaslb}[4]{#1;#2 \vdash #3 \rhd #4}
\newcommand{\tyhole}[5]{#1: #2 \mapsto_{#3} {#4}^{#5}}
\newcommand{\blkhole}[3]{#1: #2 \mapsto #3}
\newcommand{\cfghole}[3]{#1: #2 \mapsto #3}
\newcommand{\substctx}[2]{{#1}^{#2}}
\newcommand{\substlbs}[2]{{#1}^{#2}}
\newcommand{\restrictsubst}[2]{{#1}_{#2}}
\newcommand{\subsubst}[2]{#1 \subseteq #2}
\newcommand{\isrw}[3]{#1: #2 \mapsto #3}
\newcommand{\mbind}{\mathbin{{>}\hspace{-0.1em}{>}\hspace{-0.1em}{=}}}
% \newcommand{\strictlbsubst}[3]{#1: #2 \rightsquigarrow_= #3}

% Denotational semantics
\newcommand{\dnt}[1]{\llbracket{#1}\rrbracket}
\newcommand{\ednt}[1]{\left\llbracket{#1}\right\rrbracket}
\newcommand{\upg}[2]{{#1}^{\uparrow #2}}

% Weak memory
\newcommand{\bufloc}[1]{\overline{#1}}

% Branding
\newcommand{\isotopessa}{\ms{isotope_{SSA}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2023}
\acmYear{2023}
\acmDOI{XXXXXXX.XXXXXXX}

%%
%% These commands are for a JOURNAL article.
% \acmJournal{JACM}
% \acmVolume{37}
% \acmNumber{4}
% \acmArticle{111}
% \acmMonth{8}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

\begin{document}

\title{Effectful Semantics for Substructural SSA}

\author{Neel Krishnaswami}
\email{nk480@cl.cam.ac.uk}
\orcid{0000-0003-2838-5865}

\author{Jad Ghalayini}
\email{jeg74@cl.cam.ac.uk}
\orcid{0000-0002-6905-1303}

\begin{abstract}
  Static single-assignment form (SSA) is one of the most widely used
  compiler intermediate representations. In this paper, we build on
  the observation that SSA can be seen as a language of basic blocks
  with arguments (or first-order procedures tail-calling one another)
  to give a type theory for SSA programs. Notably, the type system and
  equational theory of our language is designed to support
  substituting impure terms for variables, which makes it easy to
  validate compiler optimisations such as hoisting assignments out of
  loops.

  We give a categorical axiomatisation of our type theory and show
  that our equational theory is sound with respect to it. Our
  categorical semantics generalises the notions of premonoidal
  category and Freyd category to support substructural features,
  including both support for substructural types, as well as
  F\"{u}hrmann's notion of central, copyable and discardable
  morphisms.

  Finally, we exhibit a number of concrete models of our categorical
  axiomatisation, including in particular a model of TSO-style weak
  memory based on the semantics of \citet{sparky}. This demonstrates
  that our approach is strong enough to connect machine models of
  concurrency to compiler IRs supporting a rich collection of sound
  rewrites.
\end{abstract}

\begin{CCSXML}
  <ccs2012>
  <concept>
  <concept_id>10003752.10010124.10010131.10010133</concept_id>
  <concept_desc>Theory of computation~Denotational semantics</concept_desc>
  <concept_significance>500</concept_significance>
  </concept>
  <concept>
  <concept_id>10003752.10010124.10010131.10010137</concept_id>
  <concept_desc>Theory of computation~Categorical semantics</concept_desc>
  <concept_significance>500</concept_significance>
  </concept>
  <concept>
  <concept_id>10003752.10003790.10011740</concept_id>
  <concept_desc>Theory of computation~Type theory</concept_desc>
  <concept_significance>500</concept_significance>
  </concept>
  <concept>
  <concept_id>10003752.10003790.10011742</concept_id>
  <concept_desc>Theory of computation~Separation logic</concept_desc>
  <concept_significance>300</concept_significance>
  </concept>
  </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Theory of computation~Denotational semantics}
\ccsdesc[500]{Theory of computation~Categorical semantics}
\ccsdesc[500]{Theory of computation~Type theory}
\ccsdesc[300]{Theory of computation~Separation logic}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{SSA, Categorical Semantics, Elgot Structure, Effectful Category}

% \received{20 February 2007}
% \received[revised]{12 March 2009}
% \received[accepted]{5 June 2009}

\maketitle

\section{Introduction}

Static Single Assignment, or SSA, has been a cornerstone of compiler design ever
since its introduction by \cite{ssa-original,rosen-gvn-1988} in the late 1980s.
Older compiler intermediate representations permitted the same local variable to
be assigned multiple times, whereas SSA requires each variable bindings to be
immutable -- they cannot be modified after initialisation. This makes many
dataflow analyses easier to compute, and (just as in functional languages) makes
it much easier to justify compiler optimisations in terms of equational
rewrites.

In fact, the relationship between SSA form and functional programming has been
known for decades: \citet{kelsey-ssa-cps} showed that SSA can be seen as a
subset of continuation-passing style, and \citet{appel-ssa} built on this to
observe that every SSA program can be viewed as a collection of first-order
functions which tail-call one another. This makes the equivalence of the
functional view to the traditional $\phi$-node presentation very easy to see. In
the functional style, each tail call lists its arguments. In contrast, a
$\phi$-node collects together the arguments from every callsite, which lets the
tail-calls/jumps to omit their arguments. Thus, the main difference between SSA
and a functional program is where in the program text the arguments to a tail
call are listed.

This "basic blocks with arguments" presentation of SSA has become very popular
in industry, with compiler backends such as Cranelift~\cite{cranelift},
MLIR~\cite{mlir}, and SIL~\cite{SIL} all using this representation. This tight
connection might suggest that the semantics of SSA representations is trivial:
we can just interpret an SSA program as a corresponding functional program, and
then rewrite the SSA programs using the equational theory of functional
programs.

While this is a perfectly sound thing to do, it is not strong enough to write a
good compiler. Compiler IRs like SSA are fundamentally about effectful programs,
and the equational theory of effectful functional programs is very weak. Instead
of a full substitution rule, the equational theory of impure functional
languages only justifies substituting values for variables (the $\beta$-value
rule). This weak equational rule suffices to validate some basic compiler
optimisations such as constant propagation and copy propagation, but is not
strong enough to justify many other essential compiler optimisations. For
example, hoisting a load out of a loop or conditional involves moving an
effectful term out of its original position in the program (see Equation
\ref{eqn:hoist}). As another example, dead store elimination removes an
effectful term from the program altogether.

%TODO: note that dead store elimination requires refinement, which we're not
%covering in this paper...

These examples illustrate that many optimisations require taking
\emph{effectful} terms and moving them around and substituting them for
variables. Since this is not a generally valid program transformation, we need a
finer classification of programs than merely pure and effectful to judge when a
substitution is allowed.

A useful taxonomy was introduced by \citet{fuhrmann-direct-1999}, classifying
effects based on (1) whether or not the order of effects mattered, (2) whether
it was sound to discard an effectful operation, and (3) whether it was sound to
duplicate an effectful operation. Suppose we substitute \texttt{print("hello")}
for $x$ in the program below:
\begin{equation}
\begin{BVerbatim}[baseline=c]
let x = print("hello");
let y = print("world");
let z = x;   
...
\end{BVerbatim}
\qquad \qquad \centernot\implies \qquad \qquad
\begin{BVerbatim}[baseline=c]
let y = print("world");
let x = print("hello");
...
\end{BVerbatim}
\end{equation}
In this case, substituting a term for a variable changes the order the I/O
actions are performed, which changes the program's observable behaviour. On the
other hand, suppose we have an effect which increments a counter. Then the same
transformation is sound, since \(x\) is used exactly once:
\begin{equation}
\begin{BVerbatim}[baseline=c]
let x = increment(a);
let y = increment(b);
let z = x;
...
\end{BVerbatim}
\qquad \qquad \implies \qquad \qquad
\begin{BVerbatim}[baseline=c]
let y = increment(b);
let z = increment(a);
...
\end{BVerbatim}
\end{equation}
F\"{u}hrmann called such commutative effects \emph{central}, which we also call
\emph{linear}. It can also matter whether an effect is performed at all or not.
For example, consider:
\begin{equation}
\begin{BVerbatim}[baseline=c]
let x = may_loop();
return 5 
\end{BVerbatim}
\qquad \qquad \centernot\implies \qquad \qquad
\begin{BVerbatim}[baseline=c]
return 5
\end{BVerbatim}
\end{equation}
In this example, even though $x$ is not used, the call to \texttt{may_loop()}
cannot be erased, as the observable behaviour of the program can change if it
goes into an infinite loop. On the other hand, if the effectful operation is
merely nondeterministic, then it is safe to discard results:
\begin{equation}
\begin{BVerbatim}[baseline=c]
let x = rand();
return 5 
\end{BVerbatim}
\qquad \qquad \implies \qquad \qquad
\begin{BVerbatim}[baseline=c]
return 5
\end{BVerbatim}
\end{equation}
F\"{u}hrmann called these operations \emph{discardable}. In this paper, we use
his terminology, and also call such effects \emph{affine}. Now consider the
following example: 
\begin{equation}
\begin{BVerbatim}[baseline=c]
let x = rand();
let y = x;
let z = x;
return y == z 
\end{BVerbatim}
\qquad \qquad \centernot\implies \qquad \qquad
\begin{BVerbatim}[baseline=c]
let y = rand();
let z = rand();
return y == z
\end{BVerbatim}
\end{equation}
In this case, substituting for $x$ duplicates the call to \texttt{rand()}, and
can change this from a program which always returns true to one which can return
either true or false. On the other hand, if the effect is the possibility of
nontermination, then the substitution is sound. 
\begin{equation}
\begin{BVerbatim}[baseline=c]
let x = may_loop();
let y = x;
let z = x;
return y == z 
\end{BVerbatim}
\qquad \qquad \implies \qquad \qquad
\begin{BVerbatim}[baseline=c]
let y = may_loop();
let z = may_loop();
return y == z
\end{BVerbatim}
\end{equation}
F\"{u}hrmann called such operations \emph{copyable}. In this paper, we use his
terminology, and also call such effects \emph{relevant}. With these distinction
in place, we can \emph{define} a pure term as one which is all three of
central, copyable, and discardable.

In addition to working with effects, compiler intermediate representations also
have to track accesses to memory. This information is typically gathered via
ad-hoc alias analyses, but a compositional alternative to these analyses is to
track this information via substructural type systems such as linear and affine
types, which associate aliasing and ownership information with individual
substructural values. Having aliasing information justifies important
optimisations such as loop hoisting: 
\begin{equation}
\begin{BVerbatim}[baseline=c]
while (e) {
  x = load(a);
  body
}
\end{BVerbatim}
\qquad \qquad \implies \qquad \qquad
\begin{BVerbatim}[baseline=c]
x = load(a);
while(e) {
  body
}
\end{BVerbatim}
\label{eqn:hoist}
\end{equation}
This transformation is only sound if the body of the loop \texttt{body} does not
access \texttt{a}, which is precisely the kind of frame condition that linear
types can track. Moreover, since we already want to track whether effectful
\emph{expressions} can be duplicated or discarded, the very same machinery can
be reused to track whether \emph{values} are linear/affine/relevant.

Historically, compiler writers have treated the semantics of their IR quite
informally, since the "intended model" of the language is a simple sequential
machine operating over a flat array of bytes. Unfortunately,
over time both the intended model and the actual hardware have drifted away from
this simple mental model. Compiler frameworks have to model concurrent accesses
to memory in order to compile multithreaded code, and furthermore the IR's
memory model has to be sound with respect to the actual weak memory models
supported by hardware.

As a result, it is now much harder to judge whether a particular optimisation is
sound or not. Moreover, different kinds of hardware such as x86, GPUs, and ARM
all offer different memory models. This means that we need an abstract semantics
for compiler IRs, which can permit a separation of concerns. This will let
compiler writers validate their optimisations against the abstract semantics,
and will let us separately show that a particular machine model faithfully
implements the semantics.

In this paper, we show that this is possible. We give a type system for SSA
which uses substructural types to both track effects at a fine-grained level,
and to track ownership of values. We then rigorously justify a variety of
optimisations by reference to an abstract categorical axiomatisation of this
language. We then show that our categorical axioms have a variety of concrete
models, particularly including one of TSO-style weak memory originally
introduced by \citet{sparky}. This lets us argue that a host of SSA-based
optimisations are sound with respect to the x86 memory model.

\paragraph{Contributions}

\begin{itemize}
\item We give a type system for SSA-style programs, which has a full
substitution theorem and equational theory. This language includes substructural
types, both for controlling effects and tracking ownership in values, and our
strong substitution principle makes formulating inlining and rewriting
optimisations easier than in the traditional CFG-based formulations. 
\item We give an axiomatic semantics for our language, by showing that it has a
compositional intepretation in any effectful category with an Elgot structure.
We use this structure to give generic proofs of the soundness of substitution
and rewriting properties.
\item Our abstract semantics justifies a variety of powerful optimisation
techniques, inlining, loop fusion, hoisting, and strength reduction. 
\item We show how Elgot monads give rise to instances of our categorical
semantics. We use this to derive semantics from a general trace monad, and show
that various monad transformers such as state transformers preserve the needed
Elgot structure. 
\item We use these tools to demonstrate the existence of a variety of concrete
models satisfying these categorical axioms. Starting first from simple languges
such as state plus printing, we also show how more challenging semantics such as
the TSO semantics for weak memory also fit into our framework. This thus gives
rise to an SSA-based IR with support for weak memory operations, which is fully
semantically-justified.
\item Our weak memory semantics and our results about Elgot monads are
formalised in the Lean theorem prover.
\end{itemize}

\section{SSA Syntax}

In a standard SSA-based compiler, to represent a function, we begin by
decomposing it into basic blocks, which can be viewed as \textit{label},
followed by a sequence of \textit{instructions}, each defining a
\textit{variable}, followed by a \textit{terminator}. A program written in this
manner being in "SSA form" simply means that every variable name is defined
exactly once. For example, the program in Figure \ref{fig:3-addr} has 3 basic
blocks, one of which is a distinguished \textit{entry block}, but is
\textit{not} in SSA form, since the variables \(m, n\) and \(i\) are each
modified through subsequent assignments (\(t\) is recomputed on every iteration
of the loop, but since it only has one definition, it \textit{is} in SSA form).
To convert this program to SSA form, we would traditionally use
\(\phi\)-functions, or, equivalently, annotate each basic block with arguments,
as in Figure \ref{fig:bb-arg}.

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.33\textwidth}
    \begin{BVerbatim}[baseline=t, gobble=3]
    'entry:
      m = 0
      n = 1
      brz i 'exit 'loop
    'loop:
      t = add m n
      m = n
      n = t
      i = sub i 1
      brz i 'exit 'loop
    'exit:
      ret m
    \end{BVerbatim}
    \caption{3-address code}
    \label{fig:3-addr}
  \end{subfigure}
  \begin{subfigure}[t]{0.32\textwidth}
    \begin{BVerbatim}[baseline=t, gobble=3]
      'entry:
      m0 = 0
      n0 = 1
      brz N 
        'exit (m0) 
        'loop (i, m0, n0)
    'loop(i0, m1, n1):
      m2 = n1
      n2 = add m1 n1
      i1 = sub i0 1
      brz z1 
        'exit (m2) 
        'loop (i1, m2, n2)
    'exit(m3):
      ret m3
    \end{BVerbatim}
    \caption{Basic blocks with arguments}
    \label{fig:bb-arg}
  \end{subfigure}
  \begin{subfigure}[t]{0.33\textwidth}
    \begin{BVerbatim}[baseline=t, gobble=3]
    let m0 = 0;
    let n1 = 1;
    if eq i 0 {
      br 'exit (m0)
    } else {
      br 'loop (i, m0, n0)
    } where
    'loop(i0, m1, n1) =>
      let m2 = n1;
      let n2 = add m1 n1;
      let i1 = sub i0 1;
      if eq i1 0 {
        br 'exit (m2)
      } else {
        br 'loop (i1, m2, n2)
      }
    'exit(m3) => ret m3
    \end{BVerbatim}
    \caption{Our syntax}
    \label{fig:our-syn}
  \end{subfigure}
  \caption{A simple program for computing the \texttt{i}th Fibonacci number,
  represented as 3-address code, basic-blocks with arguments, and in our SSA
  syntax} 
  \Description{}
  \label{fig:ssa-examples}
\end{figure}

For our type theoretic presentation of SSA, we will treat a very slight
generalization of the standard basic blocks with arguments syntax, in that we
will support \textit{nested scopes} and, in particular, \textit{if-statements}.
The program in Figure \ref{fig:bb-arg} translated to our syntax is given in
Figure \ref{fig:our-syn}. These do not introduce any new expressivity: an
if-statement can be viewed as syntactic sugar for a conditional branch to one of
two anonymous basic blocks, while nested scopes simply \textit{restrict} where
we can branch to. 
% In particular, in Section \ref{ssec:ssa-norm}, we will formalize
% interconversion with traditional SSA and show how it is semantically sound to
% simply merge all nested scopes. The point of the slightly modified syntax is
% simply to make some of our metatheoretic results more practical to state.

The grammar for our syntax is given in Figure \ref{fig:ssa-grammar}, while the
typing judgements themselves are given in \ref{fig:ssa-judgements}. We begin by
giving a grammar for \textit{types} \(A, B, C\), parametrized by a set of
\textit{base types} \(X\); all syntactically well-formed types are also
semantically valid. We postulate the existence of a \textit{unit type}
\(\mb{1}\), a \textit{boolean type} \(\mb{2}\), and all tensor products \(A
\otimes B\); we write \(\otimes\) to underscore the fact that types are allowed
to be substructural, assigning every type a linearity \(\ms{lin}(A)\) satisfying
the constraints in Figure \ref{fig:ssa-structural}. In particular, the rules
\rle{unit-lin} and \rle{bool-lin} state that both the unit and boolean types
must be intuitionistic; the rule \rle{pair-lin} says that the linearity of a
pair is simply the intersection of the linearity of its components. The rules
\rle{nil-lin} and \rle{cons-lin} together say that the linearity of a context is
the intersection of the linearity of its variables, where the linearity of a
variable is the intersection of the linearity of its type and its quantity
annotation.

Our grammar proper splits program fragments into three main syntactic categories:
\begin{itemize}
  \item \textit{Expressions} \(a, b, c, d, e\), which are typed via the
  judgement \(\hasty{\Gamma}{p}{a}{A}{q}\), which says that the term \(a\)
  is of type \(A\) with \textit{centrality} \(p = 1\) if \textit{central}, or
  \(p = 0\) otherwise, and \textit{quantity} \(q \subseteq \{\ms{a}, \ms{r}\}\).
  So the centrality \(p\) indicates whether it is safe to substitute this
  term for a variable, and \(q\), plus the linearity of \(A\), control how
  many times that variable can occur: if \(\taff \subseteq q\) then the variable
  is called \textit{affine} or \textit{discardable} and it is permissible to
  leave it unused, whereas if \(\trel \subseteq q\) it is called
  \textit{relevant} or \textit{duplicable} and it is permissible to use multiple
  times. An term which is central, relevant, and affine is called
  \textit{pure}. The rules for this judgement are given in Figure
  \ref{fig:ssa-term-typing}.

  \item \textit{Blocks} \(s, t\), which are typed via the judgement
  \(\haslb{\ms{L}}{t}{\Gamma}\), which says that the block \(t\) is well-formed
  with respect to the context \(\Gamma\) and only branches to
  ("\textit{targets}") labels in \(\ms{L}\). The rules for this judgement are
  given in Figure \ref{fig:ssa-block-typing}.

\item \textit{Control-flow graphs} or \textit{CFG}s \(L\), which are typed via
  the judgement \(\lhaslb{\ms{L}}{L}{\ms{K}}\), which says that \(L\) defines a
  control-flow graph with input labels \(\ms{L}\) and output labels \(\ms{K}\);
  recursive control flow is allowed. The rules for this judgement are given in
  Figure \ref{fig:ssa-block-typing}.
\end{itemize}
Here \textit{contexts} \(\Gamma\), consist of a list of variables \(x\) given a
type \(A\) annotated with a \textit{quantity} \(q \subseteq \tint\). A variable
\(x: A^q\) is called \textit{affine} if its type \(A\) is affine (i.e.
\(\tlin{A}{\taff}\)) \textit{and} \(\taff \subseteq q\), in which case it can be
freely ignored. Similarly, \(x\) is \textit{relevant} if its type \(A\) relevant
(i.e. \(\tlin{A}{\trel}\)) \textit{and} \(\trel \subseteq q\), in which case it
can be freely ignored. This is formalized in the structural judgement
\(\csplits{\Gamma}{\Delta}{\Xi}\), which says that \(\Gamma\) can be split into
subcontexts \(\Delta\) and \(\Xi\) (weakening is simply the special case
\(\cwk{\Gamma}{\Delta} \iff \csplits{\Gamma}{\Delta}{\cdot}\)), as per the
typing rules in Figure \ref{fig:ssa-structural}. In particular, the rule
\rle{split-nil} provides a base case by saying that the empty context can be
split into two empty contexts, while the rules \rle{split-left} and
\rle{split-right} send \(x: A^q\) to the left or right subcontext repsectively;
both these rules also allow us to reduce the quantity \(q\) to a quantity \(r
\subseteq q\), "forgetting" some of our capabilities. The rule \rle{split-dup}
allows the duplication of a variable which has a relevant type \(A\)
\textit{and} a relevant quantity \(q\), with the option to weaken the quantity
to \textit{different} quantities \(r\) and \(s\) on the left-hand side and
right-hand sides respectively. Finally, the rule \rle{split-aff} simply allows
us to ignore a variable \(x: A^q\) if \(A\) and \(q\) are affine.

\textit{label-sets} \(\ms{L}\) are defined to consist of a list of
\textit{labels} \(\lbl{\ell}\) annotated with a context \(\Gamma\) and an
argument type \(A\). The judgement \(\lwk{\ms{L}}{\ms{K}}\) states that
\(\ms{L}\) is a subcontext of \(\ms{K}\), i.e., that every program which targets
\(\ms{L}\) (i.e., finishes by either nontermination or branching to one of the
labels \(\llhyp{\ell}{\Gamma}{A} \in \ms{L}\) with live variables \(\Gamma\) and
an argument of type \(A\)) can also be viewed as targeting \(\ms{K}\). The
conditions for this are formalized in Figure \ref{fig:ssa-structural} as
follows: the rule \rle{join-nil} provides a base case stating that the empty
label weakens into itself. The rule \rle{join-cons} says that if \(\ms{L}\)
weakens into \(\ms{K}\) and \(\Delta\) weakens into \(\Gamma\), then
\(\lwk{\ms{L}, \llhyp{\ell}{\Gamma}{A}}{\ms{K}, \llhyp{\ell}{\Delta}{A}}\),
allowing us to "forget" any extra variables dropped by \(\cwk{\Gamma}{\Delta}\).
Finally, the rule \rle{join-zero} simply says that if \(\lwk{\ms{L}}{\ms{K}}\),
then \(\ms{L}\) also weakens to any extension of \(\ms{K}\). Note that this
judgement is "backwards," in that the "larger" context is on the \textit{right};
this corresponds to the placement of the label context on the right-hand side of
the typing judgement for blocks, and is by analogy to our semantics for blocks
and CFGs in Section \ref{sec:semantics}.

\begin{figure}
  \begin{center}
    \begin{grammar}
      <\(A, B, C\)> ::= 
      \(X\)
      \;|\; \(\mathbf{1}\)
      \;|\; \(\mathbf{2}\)
      \;|\; \(A \otimes B\)

      <\(a, b, c, e\)> ::= \(x\) 
      \;|\; \(f\;a\)
      \;|\; \((a, b)\) 
      \;|\; \(()\) 
      \;|\; \(\ctt\) 
      \;|\; \(\cff\)
      \;|\; \(\letexpr{x}{a}{e}\)
      \;|\; \(\letexpr{(x, y)}{a}{e}\)
      % \;|\; \(\lbsplice{\ell}{x: A}{t}\)

      <\(s, t\)> ::= \(\lbrb{\ell}{a}\) 
      \;|\; \(\ite{e}{s}{t}\)
      \;|\; \(\letstmt{x}{a}{t}\)
      \;|\; \(\letstmt{(x, y)}{a}{t}\)
      \;|\; \(\ewhere{t}{L}\)

      <\(L\)> ::= \(\cdot\) \;|\; \(\lwbranch{\ell}{x: A}{t}, L\)

      <\(\Gamma\)> ::= \(\cdot\) \;|\; \(\Gamma, \thyp{x}{A}{q}\)

      <\(\ms{L}\)> ::= \(\cdot\) \;|\; \(\ms{L}, \lbl{\ell}[\Gamma](x: A)\)

      <\(p\)> ::= 0 \;|\; 1

      <\(q\)> ::= \(\varnothing\) 
      \;|\; \(\taff\) 
      \;|\; \(\trel\) 
      \;|\; \(\tint\)
    \end{grammar}
  \end{center}
  \caption{Grammar for \isotopessa, parametrized over a set of instructions \(f \in \mc{I}\)}
  \Description{Grammar for isotope-SSA}
  \label{fig:ssa-grammar}
\end{figure}

\begin{figure}
  \begin{center}        
    \begingroup
    \renewcommand{\arraystretch}{1.5}
    \setlength{\tabcolsep}{2em}
    \begin{tabular}{rl}
        \multicolumn{1}{c}{Judgment} & \multicolumn{1}{c}{Meaning} \\ \hline
        \(\hasty{\Gamma}{p}{a}{A}{q}\) & \(a\) is a term of type \(A\) in
        context \(\Gamma\) with centrality \(p\) and quantity \(q\) \\
        \(\haslb{\Gamma}{t}{\ms{L}}\) & \(t\) is a block targeting label-set
        \(\ms{L}\) in context \(\Gamma\) \\
        \(\lhaslb{\ms{L}}{L}{\ms{K}}\) & The control-flow graph \(L\) sends
        input labels \(\ms{L}\) to output labels \(\ms{K}\) \\
        \(\csplits{\Gamma}{\Delta}{\Xi}\) & The context \(\Gamma\) splits into
        \(\Delta\) and \(\Xi\) \\
        \(\lwk{\ms{L}}{\ms{K}}\) & The label-set \(\ms{L}\) weakens to the
        label-set \(\ms{K}\) \\
        % \(\tlin{A}{q}\) & The type \(A\) can be used with linearity \(q\) \\
        % \(\ltlin{A}{r}{q}\) & The type \(A\) can be used with linearity \(q
        % \subseteq r\) (i.e. \(\tlin{A}{q} \land q \subseteq r\)) \\
        % \(\tlin{\Gamma}{q}\) & The context \(\Gamma\) has linearity \(q\) \\
        \(\cwk{\Gamma}{\Delta}\) & \(\Gamma\) is a weakening of \(\Delta\) (i.e.
        \(\csplits{\Gamma}{\Delta}{\cdot}\))
    \end{tabular}
    \endgroup
  \end{center}
  \caption{Typing judgements for \isotopessa}
  \Description{Typing judgements for isotope-SSA}
  \label{fig:ssa-judgements}
\end{figure}

\begin{figure}
  \begin{gather*}    
    % \prftree[r]{\rle{base-lin}}{q \subseteq \ms{lin}(X)}{\tlin{X}{q}} \qquad
    \prftree[r]{\rle{unit-lin}}{}{\tlin{\mathbf{1}}{q}} 
    \qquad
    \prftree[r]{\rle{bool-lin}}{}{\tlin{\mathbf{2}}{q}} 
    \qquad
    \prftree[r]{\rle{pair-lin}}{\tlin{A}{q}}{\tlin{B}{q}}{\tlin{A \otimes B}{q}} \\
    \prftree[r]{\rle{nil-lin}}{\tlin{\cdot}{q}} \qquad
    \prftree[r]{\rle{cons-lin}}{\ltlin{A}{r}{q}}{\tlin{\Gamma}{q}}
      {\tlin{\Gamma, \thyp{x}{A}{r}}{q}} \\
    \prftree[r]{\rle{split-nil}}{\csplits{\cdot}{\cdot}{\cdot}} \qquad
    \prftree[r]{\rle{split-left}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {r \subseteq q}
      {\csplits{\Gamma, \thyp{x}{A}{q}}{\Delta, \thyp{x}{A}{r}}{\Xi}} \qquad
    \prftree[r]{\rle{split-right}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {r \subseteq q}
      {\csplits{\Gamma, \thyp{x}{A}{q}}{\Delta}{\Xi, \thyp{x}{A}{r}}} \\
    \prftree[r]{\rle{split-dup}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\ltlin{A}{q}{\trel}}
      {r, s \subseteq q}
      {\csplits{\Gamma, \thyp{x}{A}{q}}{\Delta, \thyp{x}{A}{r}}{\Xi, \thyp{x}{A}{s}}}
      \qquad
    \prftree[r]{\rle{split-drop}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\ltlin{A}{q}{\taff}}
      {\csplits{\Gamma, \thyp{x}{A}{q}}{\Delta}{\Xi}}
      \\
    \prftree[r]{\rle{join-nil}}{\lwk{\cdot}{\cdot}} \qquad
    \prftree[r]{\rle{join-cons}}
      {\lwk{\ms{L}}{\ms{K}}}
      {\cwk{\Gamma}{\Delta}}
      {\lwk{\ms{L}, \llhyp{\ell}{\Gamma}{A}}{\ms{K}, \llhyp{\ell}{\Delta}{A}}} 
      \qquad
    \prftree[r]{\rle{join-zero}}
      {\lwk{\ms{L}}{\ms{K}}}
      {\lwk{\ms{L}}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}} 
  \end{gather*}
  \caption{Structural rules for \isotopessa}
  \Description{Structural rules for isotope-SSA}
  \label{fig:ssa-structural}
\end{figure}

We can now get to the typing judgement for expressions in Figure
\ref{fig:ssa-term-typing}. The rule \rle{var} is essentially the standard rule
for variables, with the requirement that, other than \(x\), all variables in
\(\Gamma\) are affine. Note that the quantity \(q\) of the \textit{term}
\(x\) is must satisfy \(q \subseteq q'\), where \(x: A^{q'} \in \Gamma\); this
reflects that we're allowed to \textit{forget} capabilities, but not
\textit{add} them. The \rle{app} rule is also mostly standard: given an
instruction \(f\) which takes in an argument \(A\) and has a return type of
\(B\) and a \textit{central} term \(\hasty{\Gamma}{1}{a}{A}{q}\), we
conclude that \(\hasty{\Gamma}{p}{f\;a}{B}{q}\). We, of course, require that
\(f\) itself have centrality \(p\) and quantity \(q\), i.e. that \(f \in
\mc{I}_p^q(A, B)\). The \rle{pair} rule says that we can type a pair \((a, b)\)
in \(\Gamma\) if \(\csplits{\Gamma}{\Delta}{\Xi}\) where \(\Delta\) types \(a\)
and \(\Xi\) types \(b\); the quantity of a pair is simply the intersection of
the quantity of its components (this is expressed in the rule by allowing any
quantity which types \textit{both} components), and both components must be
central. Finally, we have the rules \rle{let} and \rle{let2}, corresponding to
let-bindings and destructuring let-bindings respectively. In both cases, we
split the context \(\Gamma\) into subcontext \(\Delta, \Xi\), one of which is
used to type a central subterm \(a\), which is bound to variables \(x\) or
destructured to variables \(x, y\) respectively, in a potentially non-central
term \(e\) which is otherwise typed in \(\Delta\).

\begin{figure}
  \begin{gather*}    
    \prftree[r]{\rle{var}}
      {\cwk{\Gamma}{\thyp{x}{A}{q}}}
      {\hasty{\Gamma}{p}{x}{A}{q}} \qquad
    \prftree[r]{\rle{app}}
      {f \in \mc{I}_p^q(A, B)}
      {\hasty{\Gamma}{1}{a}{A}{q}}
      {\hasty{\Gamma}{p}{f\;a}{B}{q}} \qquad
    \prftree[r]{\rle{pair}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\hasty{\Delta}{1}{a}{A}{q}}
      {\hasty{\Xi}{1}{b}{B}{q}}
      {\hasty{\Gamma}{p}{(a, b)}{A \otimes B}{q}} \\
    \prftree[r]{\rle{unit}}
      {\cwk{\Gamma}{\cdot}}
      {\hasty{\Gamma}{p}{()}{\mb{1}}{q}} \qquad
    \prftree[r]{\rle{true}}
      {\cwk{\Gamma}{\cdot}}
      {\hasty{\Gamma}{p}{\ctt}{\mb{2}}{q}} \qquad
    \prftree[r]{\rle{false}}
      {\cwk{\Gamma}{\cdot}}
      {\hasty{\Gamma}{p}{\cff}{\mb{2}}{q}} \\
    \prftree[r]{\rle{let}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\hasty{\Delta, \thyp{x}{A}{}}{p}{e}{B}{q}}
      {\hasty{\Xi}{1}{a}{A}{q}}
      {\hasty{\Gamma}{p}{\letexpr{x}{a}{e}}{B}{q}} \qquad
    % \prftree[r]{\rle{blk}}
    %   {\haslb{\Gamma}{t}{\llhyp{\ell}{\cdot}{A}}}
    %   {\hasty{\Gamma}{0}{\lbsplice{\ell}{A}{t}}{A}{\varnothing}} 
      \\
    \prftree[r]{\rle{let2}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\hasty{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{p}{e}{C}{q}}
      {\hasty{\Xi}{1}{a}{A \otimes B}{q}}
      {\hasty{\Gamma}{p}{\letexpr{(x, y)}{a}{e}}{C}{q}}
  \end{gather*}
  \caption{Typing rules for \isotopessa expressions}
  \Description{Typing rules for isotope-SSA expressions}
  \label{fig:ssa-term-typing}
\end{figure}

The rules for block typing are given in Figure \ref{fig:ssa-block-typing}.
\rle{br} says that, given a pure term \(a: A\) (that is, central, relevant, and
affine), we can type a block \(\lbrb{\ell}{a}\) which branches to \(\ell\) with
argument \(a\), where \(\ms{L}\) contains the context \(\ell\) with an argument
type of \(A\) and an appropriate set of unused variables. The rule \rle{ite}
typechecks if-then-else statements by stating that, if \(\Gamma\) splits into
\(\Delta\) and \(\Xi\), where \(\Xi\) typechecks a pure, intuitionistic
discriminant \(e: \mb{2}\) and \(\Delta\) typechecks both \(s, t\) to branch
into \(\ms{L}\), then the whole if-statement \(\ite{e}{s}{t}\) also typechecks
as branching into \(\ms{L}\) in constext \(\Gamma\). \rle{let-blk} and
\rle{let2-blk} are quite similar to their term analogues, except that the
bound term \(e\) is allowed to be impure (otherwise, programs could only
make use of pure functions!), and the remainder of the context plus the bound
variable is used to type a block rather than a term. Finally, probably the most
interesting rule is \rle{where}, which says that, given a block \(t\) which
branches to \(\ms{L}\) in context \(\Gamma\), and labels \(L\) mapping
\(\ms{L}\) to \(\ms{K}\), \(\ewhere{t}{L}\) branches to \(\ms{K}\). Here, \(L\)
mapping \(\ms{L}\) to \(\ms{K}\) is defined using the typing rules \rle{nil-br},
which says that an empty control-flow graph simply forwards the input labels
directly to the output labels, and \rle{cons-br}, which says that if \(L\) maps
\(\ms{L}\) to either \(\ms{K}\) or the label \(\lbl{\ell}[\Gamma](A)\), and
\(t\) branches to \(\ms{L}\) in the context \(\Gamma, \thyp{x}{A}{}\), then
\(L\) maps \(\ms{L}\) to \(\ms{K}\), having (potentially recursively) defined
\(\lbl{\ell}\).

\begin{figure}
  \begin{gather*}    
    \prftree[r]{\rle{br}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\lwk{\llhyp{\ell}{\Delta}{A}}{\ms{L}}}
      {\hasty{\Xi}{1}{a}{A}{\tint}}
      {\haslb{\Gamma}{\lbrb{\ell}{a}}{\ms{L}}} 
    \\
    \prftree[r]{\rle{ite}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\hasty{\Xi}{1}{e}{\mb{2}}{\tint}}
      {\haslb{\Delta}{s}{\ms{L}}}
      {\haslb{\Delta}{t}{\ms{L}}}
      {\haslb{\Gamma}{\ite{e}{s}{t}}{\ms{L}}} 
    \\
    \prftree[r]{\rle{let-blk}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\haslb{\Delta, \thyp{x}{A}{}}{t}{\ms{L}}}
      {\hasty{\Xi}{0}{a}{A}{q}}
      {\haslb{\Gamma}{\letstmt{x}{a}{t}}{\ms{L}}} 
    \\
    \prftree[r]{\rle{let2-blk}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\haslb{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{t}{\ms{L}}}
      {\hasty{\Xi}{0}{a}{A \otimes B}{q}}
      {\haslb{\Gamma}{\letstmt{(x, y)}{a}{t}}{\ms{L}}} \qquad
    \prftree[r]{\rle{where}}
      {\haslb{\Gamma}{t}{\ms{L}}}
      {\lhaslb{\ms{L}}{L}{\ms{K}}}
      {\haslb{\Gamma}{\ewhere{t}{L}}{\ms{K}}}
    \\
    \prftree[r]{\rle{nil-br}}
      {\lhaslb{\ms{L}}{\cdot}{\ms{L}}} \qquad
    \prftree[r]{\rle{cons-br}}
      {\lhaslb{\ms{L}}{L}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}}
      {\haslb{\Gamma, \thyp{x}{A}{}}{t}{\ms{L}}}
      {\lhaslb{\ms{L}}{L, \lwbranch{\ell}{x: A}{t}}{\ms{K}}}
  \end{gather*}
  \caption{Typing rules for \isotopessa blocks}
  \Description{Typing rules for isotope-SSA blocks}
  \label{fig:ssa-block-typing}
\end{figure}

\subsection{Weakening and Substitution}

We can now state some of the basic metatheoretic properties of \isotopessa. We
begin with weakening, which we state in our setting as follows:
\begin{lemma}[Weakening] \
  \begin{itemize}
    \item If \(\cwk{\Gamma}{\Delta}\), \(q \subseteq r\), and
    \(\hasty{\Delta}{p}{a}{A}{r}\), then \(\hasty{\Gamma}{p}{a}{A}{q}\)
    \item If \(\cwk{\Gamma}{\Delta}\) and \(\haslb{\Delta}{t}{\ms{L}}\), then
    \(\haslb{\Gamma}{t}{\ms{L}}\)
    \item If \(\lwk{\ms{L}}{\ms{K}}\) and \(\haslb{\Gamma}{t}{\ms{L}}\), then \(\haslb{\Gamma}{t}{\ms{K}}\)
    % \item If \(\lwk{\ms{L}}{\ms{K}}\) and \(\lhaslb{\ms{W}}{L}{\ms{L}}\) then
    % \(\lhaslb{\ms{W}}{L}{\ms{K}}\)
  \end{itemize}
\end{lemma}
That is, we can add unused \textit{affine} variables to a context,
\textit{reduce} the quantity of an term's type, add arbitrary unused labels, and
remove \textit{affine} variables from existing labels and still keep our typing
judgements valid. This can be proved by a relatively straightforward mutual
induction.

\begin{figure}
  \begin{center}        
    \begingroup
    \renewcommand{\arraystretch}{1.5}
    \setlength{\tabcolsep}{2em}
    \begin{tabular}{rl}
        \multicolumn{1}{c}{Judgment} & \multicolumn{1}{c}{Meaning} \\ \hline
        \(\issubst{\gamma}{\Theta}{\Gamma}\) &
        \(\gamma\) is a substitution taking \(\Theta\) to \(\Gamma\) \\
        \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\) &
        \(\mc{L}\) is a label substitution taking \(\ms{L}\) to \(\ms{K}\)
    \end{tabular}
    \endgroup
  \end{center}
  \caption{Metatheoretic typing judgements for \isotopessa}
  \Description{Metatheoretic typing judgements for isotope-SSA}
  \label{fig:ssa-meta-judgements}
\end{figure}

\begin{figure}
  \begin{gather*}
    \boxed{\gamma: \ms{Var} \to \ms{Expr}}
    \\
    \prftree[r]{\rle{subst-nil}}
      {\issubst{\gamma}{\cdot}{\cdot}}
      \qquad
    \prftree[r]{\rle{subst-cons}}
      {\issubst{\gamma}{\Theta_\Gamma}{\Gamma}}
      {\hasty{\Theta_x}{1}{\gamma(x)}{A}{q}}
      {\tlin{\Theta_x}{q}}
      {\csplits{\Theta}{\Theta_\Gamma}{\Theta_x}}
      {\issubst{\gamma}{\Theta}{\Gamma, \thyp{x}{A}{q}}}
    \\
    \boxed{\mc{L}: \ms{Label} \to \ms{Expr} \to \ms{Block} \text{ s.t. for fresh } x, [a/x]\mc{L}(\lbl{\ell}, x) = \mc{L}(\lbl{\ell}, a)}
    \\
    \prftree[r]{\rle{lb-subst-nil}}
    {\lbsubst{\mc{L}}{\cdot}{\ms{K}}}
      \qquad
    \prftree[r]{\rle{lb-subst-cons}}
      {\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
      {\haslb{\Gamma, \thyp{x}{A}{}}{\mc{L}(\lbl{\ell}, x)}{\ms{K}}}
      {\lbsubst{\mc{L}}{\ms{L}, \llhyp{\ell}{\Gamma}{A}}{\ms{K}}}
  \end{gather*}
  \caption{ Typing rules for \isotopessa substitutions. We generally assume that
    for all but finitely many variables \(x\) and labels \(\lbl{\ell}\),
    \(\gamma(x) = x\) and \(\mc{L}(\lbl{\ell}, x) = \lbrb{\ell}{x}\). }
    \Description{Typing rules for isotope-SSA substitutions}
  \label{fig:ssa-subst-typing}
\end{figure}

Our main metatheoretic results will be \textit{substitution} and
\textit{label-substitution}, which will require introducing some substitution
judgements to state properly, which we do in Figure
\ref{fig:ssa-meta-judgements}. In particular, the judgement
\(\issubst{\gamma}{\Theta}{\Gamma}\) says that \(\gamma\), which is taken to be
a map from variable names to expressions, is a valid substitution taking
\(\Theta\) to \(\Gamma\). What this means is that, for each variable
\(\thyp{x}{A}{q}\) in \(\Gamma\), \(\gamma(x)\) can be interpreted as a
\textit{central} term of type \(A^q\) in a subcontext of \(\Theta\), such that
when this is done for all \(\thyp{x}{A}{q} \in \Gamma\) any non-affine variables
in \(\Theta\) are used up: this is described formally via the typing rules
\rle{subst-nil}, which says that any function \(\gamma\) takes the empty context
to itself, and \rle{subst-cons}, which says that if \(\gamma\) takes
\(\Theta_\Gamma\) to \(\Gamma\), \(\hasty{\Theta_x}{1}{\gamma(x)}{A}{q}\), \(q
\subseteq \ms{lin}(\Theta_x)\) (to avoid necessating the duplication/discarding
of substructural variables), and \(\Theta\) splits into \(\Theta_\Gamma\) and
\(\Theta_x\), then \(\issubst{\gamma}{\Theta}{\Gamma, \thyp{x}{A}{q}}\).

Given such a substitution \(\gamma\), there are two ways to use it. The first is
standard capture-avoiding substitution, which we denote \([\gamma]e\) for
expressions, \([\gamma]t\) for blocks, and \([\gamma]L\) for CFGs, as is
standard. The second is to use the substitution to generate a set of
let-bindings for a context \(\Gamma\), written
\(\exprletsubst{\substctx{\gamma}{\Gamma}}{e}\) for expressions and
\(\stmtletsubst{\substctx{\gamma}{\Gamma}}{t}\) for blocks, as defined in Figure
\ref{fig:ssa-subst-ops}. Note that this operation cannot be easily defined for
CFGs. In Section \ref{sec:semantics}, we will show that this is semantically
equivalent to regular capture-avoiding substitution, due to the fact that all
terms in a substitution are central \textit{and} that only affine computations
can be used in a term substituting for an affine variable, and only relevant
computationbs can be used in a term substituting for a relevant variable.

We can now state our substitution theorem for terms in a straightforward way: if
\(\issubst{\gamma}{\Theta}{\Gamma}\) and \(\hasty{\Gamma}{p}{e}{A}{q}\), then
\(\hasty{\Theta}{p}{[\gamma]e}{A}{q}\) and
\(\hasty{\Theta}{p}{\exprletsubst{\substctx{\gamma}{\Gamma}}{e}}{A}{q}\).
Unfortunately, the substitution theorem for blocks is a little more complicated:
the naive choice of attempting to show \(\haslb{\Gamma}{t}{\ms{L}} \implies
\haslb{\Theta}{[\gamma]t}{\ms{L}}\) does not work, since \(\ms{L}\) may contain
captured variables from \(\Gamma\) that are no longer available if we are typing
\([\gamma]t\) in the context of \(\Theta\).

\begin{figure}
  \begin{gather*}
    % \restrictsubst{\gamma}{\Gamma}(x) 
    %   =
    %   (\ms{if}\; x \in \Gamma
    %   \;\ms{then}\;\gamma(x)
    %   \;\ms{else}\;x)
    % \\
    \stmtletsubst{\substctx{\gamma}{\cdot}}{t} = t
    \qquad
    \stmtletsubst{\substctx{\gamma}{\Gamma, x: A^q}}{t} 
      = \letstmt{x}{\gamma(x)}{\stmtletsubst{\substctx{\gamma}{\Gamma}}{t}}
    \qquad
    \exprletsubst{\cdot}{e} = e
    \qquad
    \exprletsubst{\substctx{\gamma}{\Gamma, x: A^q}}{e} 
      = \letstmt{x}{\gamma(x)}{\substctx{\gamma}{\Gamma}}
    \\
    \substlbs{\gamma}{\cdot}(\lbl{\ell}, x) = \lbrb{\ell}{x}
    \qquad
    \substlbs{\gamma}{\ms{L}, \llhyp{\ell}{\Gamma}{A}}(\lbl{k}, x)
      = 
      (\ms{if}\; \lbl{k} = \lbl{\ell} 
      \;\ms{then}\;\stmtletsubst{\substctx{\gamma}{\Gamma}}{\lbrb{\ell}{x}}
      \;\ms{else}\;\substlbs{\gamma}{\ms{L}}(\lbl{\ell}, x))
    \\
    [\mc{L}](\lbrb{\ell}{a}) = \mc{L}(\lbl{\ell}, a)
    \qquad
    [\mc{L}]\ite{e}{s}{t} = \ite{e}{[\mc{L}]s}{[\mc{L}]t}
    \\
    [\mc{L}](\letstmt{x}{a}{t}) = (\letstmt{x}{a}{[\mc{L}]t})
    \qquad
    [\mc{L}](\letstmt{(x, y)}{a}{t}) = (\letstmt{(x, y)}{a}{[\mc{L}]t})
    \\
    [\mc{L}]\ewhere{t}{L} = \ewhere{[\mc{L}]t}{[\mc{L}]L}
    \\
    [\mc{L}]\cdot = \cdot
    \qquad
    [\mc{L}](\lwbranch{\ell}{x: A}{t})
    = (\lwbranch{\ell}{x: A}{[\mc{L}]t})
  \end{gather*}
  \caption{Operations on \isotopessa substitutions.}
  \Description{Operations on isotope-SSA substitutions}
  \label{fig:ssa-subst-ops}
\end{figure}

To solve this problem, we need to introduce the concept of a
\textit{label-substitution} \(\mc{L}\), which we define as a function that takes
labels \(\lbl{\ell}\) and variables \(x\) to blocks \(t\) such that, for any
fresh variable \(x\), \([a/x]\mc{L}(\lbl{\ell}, x) = \mc{L}(\lbl{\ell}, a)\). We
introduce the judgement \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\) to state that the
label-substitution \(\mc{L}\) takes the label-set \(\ms{L}\) to the label-set
\(\ms{K}\); that is, for each \(\lbl{\ell}[\Gamma](x: A) \in \ms{L}\), we have
that \(\mc{L}(\lbl{\ell}, x)\) is a block that targets \(\ms{K}\). This is
formally defined via the rules \rle{lb-subst-nil}, which states that this is
vacuously true for \(\ms{L} = \cdot\), and \rle{lb-subst-cons}, which says that
if \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\) and \(\haslb{\Gamma,
\thyp{x}{A}{}}{\mc{L}(\lbl{\ell}, x)}{\ms{K}}\) then \(\lbsubst{\mc{L}}{\ms{L},
\lhyp{\lbl{\ell}}{\Gamma}{A}}{\ms{K}}\), in Figure \ref{fig:ssa-subst-typing}.
Given a context \(\gamma\) and a label-set \(\ms{L}\), we can convert it to a
label-substitution \(\substlbs{\gamma}{\ms{L}}\) using the definition in Figure
\ref{fig:ssa-subst-ops}: using this, we can state the full substitution lemma as
follows:
\begin{lemma}[Substitution] 
  Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
  \begin{itemize}
    \item If \(\hasty{\Gamma}{p}{a}{A}{q}\), then
    \(\hasty{\Theta}{p}{[\gamma]a}{A}{q}\) and
    \(\hasty{\Theta}{p}{\exprletsubst{\substctx{\gamma}{\Gamma}}{a}}{A}{q}\)
    \item If \(\haslb{\Gamma}{t}{\ms{L}}\), then there exists \(\ms{K}\) such
    that \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) and
    \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\),
    \(\haslb{\Theta}{\stmtletsubst{\substctx{\gamma}{\Gamma}}{t}}{\ms{K}}\)
    \item If \(\lhaslb{\ms{W}}{L}{\ms{L}}\), then there exists \(\ms{K},
    \ms{W}'\) such that 
    \begin{itemize}
      \item \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) 
      \item \(\lbsubst{\substlbs{\gamma}{\ms{W}'}}{\ms{W}}{\ms{W}'}\) 
      \item \(\lhaslb{\ms{W}'}{[\gamma]L}{\ms{K}}\)
    \end{itemize}
  \end{itemize}
\end{lemma}

We can also, using the rules in Figure \ref{fig:ssa-subst-ops}, define
capture-avoiding label-substitution for blocks \([\mc{L}]t\) and CFGs
\([\mc{L}]L\), allowing us to state the following lemma:
\begin{lemma}[Label Substitution] 
  Given \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\), 
  \begin{itemize}
    \item For all \(\haslb{\Gamma}{t}{\ms{L}}\),
    \(\haslb{\Gamma}{[\mc{L}]t}{\ms{K}}\)
    \item For all \(\lhaslb{\ms{W}}{L}{\ms{L}}\),
    \(\lhaslb{\ms{W}}{[\mc{L}]L}{\ms{K}}\)
  \end{itemize}
\end{lemma}

\section{SSA Semantics}

\label{sec:semantics}

In this section, we will give a categorical semantics for \isotopessa. Starting
with Freyd categories, we will generalize to \textit{substructural effectful
categories} to support substructurality. We will then require our substructural
effectful category to have \textit{distributive coproducts} to implement
branching. Finally, we will require an \textit{Elgot structure} to implement
general, potentially nonterminating control flow.

\subsection{Substructural Categories}

We begin with some standard definitions:
\begin{definition}[Binoidal Category] 
  A \textbf{binoidal category} \(\mc{C}\) is a category equipped with a binary
  operation on objects \(\otimes: \mc{C} \times \mc{C} \to \mc{C}\) which is
  independently functorial in each argument (i.e., for each object \(A\), \(A
  \otimes -\) and \(- \otimes A\) are functors).
\end{definition}
Given \(f: \mc{C}(A, B)\), \(g: \mc{C}(A', B')\) in a binoidal category, we
write \(f \ltimes g = f \otimes A';B \otimes g\) and \(f \rtimes g = A \otimes
g;f \otimes B'\); we say \(f\) is \textbf{central} if, for all \(g\), \(f
\ltimes g = f \rtimes g\) and \(g \ltimes f = g \rtimes f\), in which case we
simply write \(f \otimes g\), \(g \otimes f\) respectively.
\begin{definition}[Premonoidal Category] 
  A \textbf{premonoidal category} is a binoidal category with an identity object
  \(I: \mc{C}\) such that there exist central natural isomorphisms \(\lambda_A:
  I \otimes A \simeq A\) (the \textbf{left unitor}), \(\rho_A: A \otimes I
  \simeq A\) (the \textbf{right unitor}), and \(\alpha_{A, B, C}: (A \otimes B)
  \otimes C \simeq A \otimes (B \otimes C)\) (the \textbf{associator})
  satisfying the \textit{pentagon} and \textit{triangle laws} (see e.g.
  \citet{maclane:71}). We say a premonoidal category is \textbf{symmetric} if it
  is equipped with a central natural isomorphism \(\sigma_{A, B}: A \otimes B
  \simeq B \otimes A\) (the \textbf{symmetry}) satisfying the \textit{hexagon
  law} (see \cite{maclane:71}). We say a premonoidal category is
  \textbf{monoidal} if every morphism is central.
\end{definition}
One issue with the notion of a premonoidal category is that it is unclear what
exactly a premonoidal functor should be: if we want them to compose, we would
like to preserve central morphisms, but trying to preserve \textit{all} central
morphisms is too strict a notion. We want, in essence, to be able to distinguish
morphisms that are truly "pure" from those which simply happen to be central. To
do that, we introduce the notion of an \textit{effectful category} as follows:
\begin{definition}[Effectful Category] 
  An \textbf{effectful category} \(\mc{V} \to \mc{C}\) is an identity on objects
  functor \(\upg{\cdot}{}: \mc{V} \to \mc{C}\) from a monoidal category
  \(\mc{V}\) to a premonoidal category \(\mc{C}\) which preserves all
  premonoidal structure (i.e. sends associators to associators, unitors to
  unitors, and tensor products to tensor products). It is \textbf{symmetric} if
  both \(\mc{V}\) and \(\mc{C}\) are, and \(\upg{\cdot}{}\) preserves the
  symmetry. A \textbf{Freyd category} is an effectful category where \(\mc{V}\)
  is Cartesian.
\end{definition}
Unfortunately, the notion of "purity" given by an effectful category only allows
us to perform relatively weak optimistions in the presence of effects. To remedy
this, we want to be able to make a finer distinction between morphisms which are
merely \textit{central} and those which are also \textit{relevant} (i.e.,
copying the input and applying them twice is the same as applying them once and
copying the output) and/or \textit{affine} (i.e., applying them and discarding
the result is the same as never doing them at all). Doing so also allows us to
add the capability of \textit{distinguishing} between objects which are affine,
relevant, neither or both, enabling us to reason about substructural type
systems categorically.

To do this, we begin by introducing the notion of a \textit{symmetric
premonoidal subcategory}, as follows:
\begin{definition}[(Symmetric) Premonoidal Subcategory]
  Given a (symmetric) premonoidal category \(\mc{C}\), we say a wide subcategory
  \(\mc{S} \subseteq \mc{C}\) is \textbf{(symmetric) premonoidal} if it is
  closed under whiskering and contains all associators, unitors, and symmetries.
  We say a subcategory is \textbf{monoidal} if, for all \(f: \mc{S}(A, B)\),
  \(g: \mc{S}(A', B')\), \(f \ltimes g = f \rtimes g\); we say it is
  \textbf{central} if this generalizes to the case where only one of \(f, g\) is
  in \(\mc{S}\) (i.e., if all morphisms in \(\mc{S}\) are central).
\end{definition} 
Note that the set of symmetric premonoidal subcategories, monoidal
subcategories, and central subcategories of a category \(\mc{C}\) are all closed
under intersections and therefore, in particular, form a complete
meet-semilattice with shared bottom element \(\ms{Skeleton}(\mc{C})\), where
\(\ms{Skeleton}(C)\) is the smallest possible symmetric premonodial subcategory.
The largest symmetric premonoidal subcategory is of course \(\mc{C}\), while the
largest central subcategory is the center of \(\mc{C}\), \(Z(\mb{C})\); the
existence of a top-element means that in these cases we may induce the structure
of a complete lattice in the usual manner.

We can now define a \textit{substructural premonoidal category}, which is simply
a premonoidal category equipped with a distinguished set of relevant and affine
morphisms
\begin{definition}[Substructural (Pre)monoidal Category]
  A \textbf{substructural (pre)monoidal category} is a tuple \((\mc{C}, \mc{A},
  \mc{R})\) where:
  \begin{itemize}
    \item \(\mc{C}\) is a symmetric premonoidal category equipped with wide
    symmetric premonoidal subcategories \(\mc{C}^{\ms{a}}, \mc{C}^{\ms{r}}\).
    We define, for \(q \subseteq \{\ms{a}, \ms{r}\}\), \(\mc{C}^q = \bigcap_{\ell \in q}\mc{C}^\ell\), where \(\mc{C}^\varnothing = \mc{C}\)
    \item \(\mc{A} \subseteq |\mc{C}|\) is a set of \textbf{affine} objects
    closed under tensor products, such that
    \begin{itemize}
      \item \(\mb{1} \in \mc{A}\)
      \item Every \(A\) in \(\mc{A}\) is equipped with a central morphism
      \(\ms{drop}(A): \mc{C}^{\{\ms{a}, \ms{r}\}}(A, \mb{1})\)
      \item \(\forall A, B \in \mc{A}, \ms{drop}(A \otimes B) = \ms{drop}(A) \otimes \ms{drop}(B);\lambda_I\)
      \item \(\forall A, B \in \mc{A}, \forall f: \mc{C}^{\ms{a}}(A, B), f;\ms{drop}(B) = \ms{drop}(A)\)
    \end{itemize}
    \item \(\mc{R} \subseteq |\mc{C}|\) is a set of \textbf{relevant} objects
    closed under tensor products, such that
    \begin{itemize}
      \item \(\mb{1} \in \mc{R}\)
      \item Every \(A\) in \(\mc{R}\) is equipped with a central morphism
      \(\ms{split}(A): \mc{C}^{\{\ms{a}, \ms{r}\}}(A, A \otimes A)\)
      \item \(\forall A, B \in \mc{R}, \ms{split}(A \otimes B) = \ms{split}(A) \otimes \ms{split}(B);\alpha;A \otimes \sigma_{A, B} \otimes B;\alpha\)
      \item Associativity: \(\forall A \in \mc{R}, \ms{split}(A);\ms{split}(A) \otimes A;\alpha = \ms{split}(A);A \otimes \ms{split}(A)\)
      \item Commutativity: \(\forall A \in \mc{R}, \ms{split}(A);\sigma = \ms{split}(A)\)
      \item Comonoid: \(\forall A \in \mc{A} \cap \mc{R}, \ms{split}(A);A \otimes \ms{drop}(A) = \rho_A^{-1} \quad \ms{split}(A);\ms{drop}(A) \otimes A = \lambda_A^{-1}\)
      \item \(\forall A, B \in \mc{R}, \forall f: \mc{C}^{\ms{r}}(A, B), f;\ms{split}(B) = \ms{split}(A);f \ltimes f = \ms{split}(A);f\rtimes f\)
    \end{itemize}
  \end{itemize}
  We call the morphisms in \(\mc{C}^{\ms{a}}\) \textbf{affine}, the
  morphisms in \(\mc{C}^{\ms{r}}\) \textbf{relevant}, and the morphisms in
  \(\mc{C}^{\{\ms{a}, \ms{r}\}}\) \textbf{pure} or \textbf{intuitionistic}. We
  call the morphisms in \(\mc{C}^\varnothing\) but not necessarily in any of the
  other \(\mc{C}^q\) \textbf{linear}. We call a substructural \textit{monoidal}
  category a \textbf{substructural Cartesian category} if \(\mc{A} = \mc{R} =
  |\mc{C}|\), in which case \(\mc{C}^{\{\ms{a}, \ms{r}\}}\) is Cartesian.
\end{definition}
A simple example of a substructural Cartesian category is the category of
relations \(\ms{Rel}\), viewed as the Kleisli category of the power-set monad
with morphisms of the form \(\ms{Rel}(A, B) = A \to \mc{P}(B)\). In particular,
we can define:
\begin{equation}
  \begin{aligned}
    \ms{Rel}^{\ms{a}}(A, B) 
    &= \{R \in \ms{Rel}(A, B) \mid \forall a, |R(a)| \geq 1\} 
    &&= A \to \mc{P}^+(B) \\
    \ms{Rel}^{\ms{r}}(A, B) 
    &= \{R \in \ms{Rel}(A, B) \mid \forall a, |R(a)| \leq 1\} 
    &&= A \to \ms{Option}(B) \\
    \implies \ms{Rel}^{\{\ms{a}, \ms{r}\}}(A, B) 
    &= \{R \in \ms{Rel}(A, B) \mid \forall a, |R(a)| = 1\} 
    &&= A \to B
  \end{aligned}
\end{equation}
Here, \(\ms{split}(A)\) is just the copy operator \(a \mapsto (a, a)\) and
\(\ms{drop}(A)\) the deletion operator \(a \mapsto ()\); it is quite trivial to
prove that these lie in \(\ms{Rel}^{\{\ms{a}, \ms{r}\}}\), being functions, and
that they satisfy the axioms of a commutative comonoid, as desired.

Note that we cannot simply have \(\ms{Rel}^{\ms{a}} = \ms{Rel}\), since, for
any morphism \(R: \ms{Rel}(A, B)\) such that \(\exists a \in A, R(a) =
\varnothing\), we have
\begin{equation}
  (R;\ms{drop}(B))(a) = \varnothing \neq \ms{drop}(A)(a) = \{()\}
\end{equation}
We will see that this corresponds to the fact that the following programs are
not semantically equal
\begin{equation}
  \ms{let}\;x = \ms{loop}(); e \not\simeq e \qquad \text{where} \qquad x \notin \ms{fv}(e)
\end{equation}
Similarly, we cannot simply have \(\ms{Rel}^{\ms{r}} = \ms{Rel}\), since,
for any morphism \(R: \ms{Rel}(A, B)\) such that \(\exists a \in A, |R(a)|
\supseteq \{b_0, b_1\}\) where \(b_0 \neq b_1\), we have that
\begin{equation}
  (R;\ms{split}(B))(a) = \{(b, b) \mid b \in R(a)\} \neq (\ms{split}(A);R)(a) = R(a) \times R(a) 
\end{equation}
since only the left-hand side contains the pair \((b_0, b_1)\), this
corresponding to the fact that the following programs are not semantically equal
\begin{equation}
  \ms{let}\;x = \ms{rand}(); (x, x) \not\simeq (\ms{rand}(), \ms{rand}())
\end{equation}
To get a substructural premonoidal category, we can apply the state monad
transformer to the power set monad to obtain the Kleisli category with morphisms
of the form \(\ms{Set}_{\ms{St}\;\mc{P}\;S}(A, B) = A \to S \to \mc{P}(B \times
S)\). Once again, it is clear that all objects have split and drop morphisms
inherited from \(\ms{Set}\), and therefore are affine and relevant.

One issue with the above definition is that, just like for regular premonoidal
categories, it is difficult to distinguish between those morphisms that are
"truly" central versus those which are just central "by accident", as we only
want to require any notion of functor we develop subsequently to preserve the
former. We can solve this the same way, by simply extending the notion of
effectful category appropriately as follows:
\begin{definition}[Susbtructural Effectful Category]
  A \textbf{substructural effectful category} \(\mc{C}\) consists of
  \begin{itemize}
    \item A substructural monoidal category \(\mc{C}_1\); we call the morphisms
    in this category \textbf{central} or \textbf{linear}
    \item A substructural premonoidal category \(\mc{C}_0\) sharing the same
    affine and relevant objects as \(\mc{C}_1\)
    \item An identity-on-objects functor \(\upg{\cdot}{}: \mc{C}_1^\varnothing
    \to \mc{C}_0^\varnothing\) preserving all substructural premonoidal
    structure, i.e., preserves all symmetries, associators, unitors, drops, and
    splits and sends \(\mc{C}_1^q\) to \(\mc{C}_0^q\). In particular,
    restricting \(\upg{\cdot}{}\) appropriately makes all \(\mc{C}_1^q \to
    \mc{C}_0^q\) into effectful categories.
  \end{itemize}
  We define \(\upg{f}{0}\) to be \(\upg{f}{}\) for morphisms in \(\mc{C}_1^q\)
  and simply \(f\) for morphisms in \(\mc{C}_0\), and \(\upg{f}{1}\) to simply
  be the identity functor on \(\mc{C}_1^q\).

  Morphisms in \(\mc{C}_p^{\ms{r}}\) are called \textbf{copyable}, morphisms
  in \(\mc{C}_p^{\ms{a}}\) are called \textbf{discardable}, while morphisms
  in \(\mc{C}_1^{\ms{a}}, \mc{C}_1^{\ms{r}}, \mc{C}_1^{\{\ms{a},
  \ms{r}\}}\) specifically are called \textbf{affine}, \textbf{relevant}, and
  \textbf{intuitionistic} or \textbf{pure} respectively.
\end{definition}
We can obtain an example of a substructural effectful category quite easily by,
considering the power-set monad above, applying the state transformer to yield
the Kliesli category with morphisms of the form \(\mc{C}_0(A, B) = A \to S \to
\mc{P}(A \times S)\). Letting \(\mc{C}_1^q = \ms{Rel}^q\) and defining
\(\upg{R}{} = \lambda a, s. R(a) \times \{s\}\), we can verify that this
satisfies the axioms of a substructural effectful category. Note that here the
\textit{central} morphisms are precisely those which do not read from or write
to the state, and therefore which commute with all other morphisms.

In general, \textit{every} premonoidal category \(\mc{C}\) can be viewed a
substructural premonoidal category, with \(\mc{C}^\varnothing = \mc{C}\) and
\(\mc{C}^{\{\ms{a}, \ms{r}\}}, \mc{C}^{\{\ms{a}\}}, \mc{C}^{\{\ms{r}\}}\) the
subcategory consisting only of the associators, unitors, and symmetry, and only
the identity object considered affine and relevant. Similarly, every
\textit{Cartesian} category \(\mc{C}\) can be viewed as a substructural
Cartesian category with all morphisms both relevant and affine; in fact, this is
the case \textit{if and only if} a category is Cartesian, hence the term
"substructural Cartesian category." We also note the following:
\begin{proposition}
  If \(\mc{C}\) is a substructural monoidal category then we can equip the
  Kleisli category \(\mc{C}_{\mb{T}}\) with the structure of a substructural
  effectful category by taking \(\ms{Aff}(\mc{C}_{\mb{T}}) = \ms{Aff}(\mc{C})\),
  \(\ms{Rel}(\mc{C}_{\mb{T}}) = \ms{Rel}(\mc{C})\), and \(\mc{C}_{\mb{T}}^q(A,
  B) = \upg{\mc{C}^q(A, B)}{}\), where \(\upg{f}{} = f;\eta\), and defining
  drops \(\upg{\ms{drop}(A)}{}\) and splits \(\upg{\ms{split}(A)}{}\). In
  particular, \(\upg{\cdot}{}: \mc{C} \to \mc{C}_{\mb{T}}\) is a substructural
  effectful category.
\end{proposition}

\subsection{Distributivity}

To add support for branching, we need to equip our category with a
\textit{distributive coproduct structure}, as follows:
\begin{definition}[Distributive (pre)monoidal category] 
  A \textbf{distributive (pre)monoidal category} is a (pre)monoidal category
  with all finite coproducts equipped with a natural family of central
  isomorphisms (the \textbf{distributor}) \(\delta_{A, B, C}: (A + B) \otimes C
  \simeq A \otimes C + B \otimes C\). This isomorphism is subject to rather
  complex coherence conditions, worked out in full for the more general case of
  \textit{rig categories} in \cite{laplaza-distributivity}, but for our purposes
  we will only make use of the following:
  \begin{equation*}
    \delta_{A, B, C}^{-1} = [\iota_0 \otimes C, \iota_1 \otimes C] 
    \qquad
    \sigma^+_{A, B} \otimes C;\delta_{B, A, C} = \delta_{A, B, C};\sigma^+_{A \otimes C, B \otimes C}
  \end{equation*}
  Note in particular that this implies that \(\iota_i \otimes C;\delta_{A, B, C}
  = \iota_i\). A substructural (pre)monodial category is \textbf{distributive}
  if it is distributive as a (pre)monoidal category, and its distributor is
  \textit{pure}. A substructural effectful category \(\mc{C}_1 \to \mc{C}_0\) is
  \textbf{distributive} if \(\mc{C}_1, \mc{C}_0\) are distributive (as a
  substructural (pre)monoidal category) and \(\upg{\cdot}{}\) strictly preserves
  all coproducts.
\end{definition}
Many (pre)monoidal categories are distributive, including of course
\(\ms{Set}\), but also categories which can be derived from monads on more
primitive categories such as \(\ms{PFun}\) and \(\ms{Rel}\), as can be deduced
from the following proposition:
\begin{proposition}
  If \(\mc{C}\) is a distributive substructural monoidal category and \(\mb{T}\)
  is a strong monad on \(\mc{C}\), then the Kleisli category \(\mc{C}_{\mb{T}}\)
  is a distributive substructural effectful category (with affine and relevant
  objects and morphisms inherited from \(\mc{C}\))
\end{proposition}
% \begin{proof}
%   Let \(\delta_{A, B, C}: \mc{C}((A + B) \otimes C, A \otimes C + B \otimes C)\) denote \(\mc{C}\)'s distributor. We will write composition in \(\mc{C}\) as "\(;\)" and composition in the Kleisli category \(\mb{T}\) as "\(\gg\)".
%   We will show that
%   \(\upg{\delta_{A, B, C}}{}: \mc{C}_{\mb{T}}((A + B) \otimes C, A \otimes C + B \otimes C)\) satisfies the desired properties of a distributor. In particular, it is:
%   \begin{itemize}
%     \item Central: this follows immediately from being the image of a pure morphism
%     \item Affine: \(\upg{\delta}{} \gg \upg{\ms{drop}}{} = \upg{(\delta;\ms{drop})}{} = \upg{\ms{drop}}{}\)
%     \item Relevant: \(\upg{\delta}{} \gg \upg{\ms{split}}{} = \upg{(\delta;\ms{split})}{} = \upg{(\ms{split};\delta \otimes \delta)}{} = \upg{\ms{split}}{} \gg \upg{\delta}{} \otimes \upg{\delta}{}\)
%     \item Left inverse: \(\upg{\delta}{} \gg \upg{[\iota_0 \otimes C, \iota_1
%     \otimes C]}{} = \upg{\delta;[\iota_0 \otimes C, \iota_1 \otimes C]}{} =
%     \ms{id}\)
%     \item Right inverse: \(\upg{[\iota_0 \otimes C, \iota_1 \otimes C]}{} \gg \upg{\delta}{} = \upg{[\iota_0 \otimes C, \iota_1 \otimes C];\delta}{} = \ms{id}\)
%     \item Symmetric: \(\upg{\sigma^+}{} \otimes C \gg \upg{\delta}{} =
%     \upg{\sigma^+ \otimes C;\delta}{} = \upg{\delta;\sigma^+}{} = \upg{\delta}{}
%     \gg \upg{\sigma^+}{}\)
%   \end{itemize}
% \end{proof}

\subsection{Elgot Categories}

The structures we've developed so far are enough to support non-looping control
flow, but to allow general, potentially nonterminating control flow, we need a
way to reason about iteration. To do so, we introduce the notion of an
\textit{Elgot category} as follows:
\begin{definition}[Elgot category] 
  An \textbf{Elgot category} is a category \(\mc{C}\) with coproducts equipped
  with an operator \(\cdot^\dagger: \mc{C}(A, B + A) \to \mc{C}(A, B)\)
  satisfying the following properties:
  \begin{itemize}
    \item \textbf{Fixpoint:} \(f;[\ms{id}, f^\dagger] = f^\dagger\)
    \item \textbf{Naturality:} \((f;g + \ms{id})^\dagger = f^\dagger;g\)
    \item \textbf{Codiagonal:} \(\forall f \in \mc{C}(A, (B + A) + A), (f^\dagger)^\dagger = (f;[\ms{id}, \iota_1])^\dagger\)
    \item \textbf{Uniformity:}
    \(
      h;f = g;\ms{id} + h \implies h;f^\dagger = g^\dagger
    \)
  \end{itemize}
  A monad \(\mb{T}\) on \(\mc{C}\) such that its Kleisli cateogry
  \(\mc{C}_{\mb{T}}\) is Elgot is called an \textbf{Elgot monad}.
\end{definition}
Some important properties which hold for all Elgot categories include:
\begin{itemize}
  \item \textbf{Dinaturality:} \((g;[\iota_0, h])^\dagger = g;[\ms{id},
  (h;[\iota_0, g])^\dagger]\)
  \item \textbf{Squaring:} \((f;[\iota_0, f])^\dagger = f^\dagger\)
\end{itemize}
Proofs can be found in Lemma 31 of \cite{goncharov-squaring}.

One important way of generating Elgot categories is through monads: a monad
whose Kliesli category is Elgot is called an \textbf{Elgot monad}. Examples of
Elgot monads include \(\ms{Option}\) (where \(\ms{None}\) is interpreted as
nontermination), the powerset monad (where the empty set is interpreted as
nontermination), and the trace monads we will explore in Section
\ref{ssec:trace-monads}.

It turns out that various monad transformers preserve Elgot structure, allowing
us to build up Elgot monads in a composable fashion. In particular,
\begin{proposition}
  If \(\mb{T}\) is an Elgot monad on \(\ms{Set}\), then \(\ms{StateT}\;S\;\mb{T}
  = \lambda A. S \to \ms{T}(A \times S)\), \(\ms{ReaderT}\;\mb{T}\;S = \lambda
  A. S \to \ms{T}(A)\), and \(\ms{WriterT}\;\mb{T}\;S = \lambda A. \ms{T}(A
  \times S)\) are all Elgot, with iteration operators lifted from \(\mb{T}\) in
  the obvious manner.
\end{proposition}

\subsection{Denotational Semantics}

We're now ready to give a semantics for an \isotopessa dialect with a set of
types \(\mc{T}\) and instructions \(\mc{I}\). We'll need:
\begin{itemize}
  \item A \textit{substructural effectful category} \(\mc{C} = \mc{C}_1 \to
  \mc{C}_0\) such that \(\mc{C}_0\) is \textit{distributive} and \textit{Elgot},
  equipped with
  \item A function \(\ms{base}(X)\) from base types \(X \in \mc{T}\) to objects
  in \(|\mc{C}|\), such that \(\ms{a} \in \ms{lin}(X) \implies \ms{base}(X) \in
  \ms{Aff}(\mc{C})\) and \(\ms{r} \in \ms{lin}(X) \implies \ms{base}(X) \in
  \ms{Rel}(\mc{C})\)
  \item A function \(\ms{inst}(f)\) from instructions \(f \in \mc{I}_p^q(A, B)\)
  to morphisms in \(\mc{C}_p^q(\dnt{A}, \dnt{B})\)
\end{itemize}

We begin by giving a semantics for types and contexts in terms of objects in
\(|\mc{C}|\) in Figure \ref{fig:ssa-type-semantics}: we map tensor products to
tensor products, the unit type to the monoidal unit, and booleans to \(\mb{1} +
\mb{1}\). Contexts are simply interpreted as the tensor products of the variable
types which appear in them; in particular, the quantity annotations have
\textit{no} effect on a context's semantics. We then give a semantics for
structural judgements in Figure \ref{fig:ssa-structural-semantics}. In
particular, we begin by interpreting the judgements that a type is affine (i.e.
\(\tlin{A}{\taff}\)) or relevant (i.e. \(\tlin{A}{\trel}\)) as the drop morphism
\(\ms{drop}(\dnt{A}): \mc{C}(\dnt{A}, I)\) and split morphism
\(\ms{drop}(\dnt{A}): \mc{C}(\dnt{A}, \dnt{A} \otimes \dnt{A})\) respectively.
We can then interpret a context split \(\csplits{\Gamma}{\Delta}{\Xi}\) as a
pure morphism \(\mc{C}^\tint_1(\dnt{\Gamma}, \dnt{\Delta} \otimes \dnt{\Xi})\)
which partitions the variables as appropriate, splitting and dropping as
necessary. We interpret a context weakening \(\cwk{\Gamma}{\Delta}\) as the
associated split \(\csplits{\Gamma}{\Delta}{\cdot}\) postcomposed with a unitor.
Finally, we interpret a label-set weakening \(\lwk{\ms{L}}{\ms{K}}\) as a
pure morphism \(\mc{C}^\tint_1(\dnt{\ms{L}}, \dnt{\ms{K}})\), where
\rle{join-nil} is the identity, \rle{join-cons} is the sum of
\(\dnt{\lwk{\ms{L}}{\ms{K}}}\) and \(\dnt{\cwk{\Gamma}{\Delta}}\), and
\rle{join-zero} is \(\dnt{\lwk{\ms{L}}{\ms{K}}}\) postcomposed with the left
injection.

\begin{figure}
  \begin{gather*}
    \boxed{\dnt{A}}: |\mc{C}| \\
    \dnt{X} = \ms{base}(X) 
    \qquad
    \dnt{\mb{1}} = \mb{1}
    \qquad
    \dnt{\mb{2}} = \mb{2} = \mb{1} + \mb{1}
    \qquad \dnt{A \otimes B} = \dnt{A} \otimes \dnt{B} \\
    \boxed{\dnt{\Gamma}}: |\mc{C}| \\
    \dnt{\cdot} = \mb{1} \qquad 
    \dnt{\Gamma, \thyp{x}{A}{q}} = \dnt{\Gamma} \otimes \dnt{A} \\
    \boxed{\dnt{\ms{L}}}: |\mc{C}| \\
    \dnt{\cdot} = \mb{0} \qquad
    \dnt{\ms{L}, \llhyp{\ell}{\Gamma}{A}} = 
      \dnt{\ms{L}} + \dnt{\Gamma} \otimes \dnt{A}
  \end{gather*}
  \caption{Semantics for \isotopessa types and contexts}
  \Description{Semantics for isotope-SSA types and contexts}
  \label{fig:ssa-type-semantics}
\end{figure}

\begin{figure}
  \begin{gather*}
    \boxed{\dnt{\tlin{A}{\taff}: \mc{C}_1^\tint(\dnt{A}, \mb{1})}}
      \\
      \dnt{\tlin{A}{\taff}} = \ms{drop}(\dnt{A})
      % \dnt{\tlin{X}{\taff}} = \ms{drop}(X) \qquad
      % \dnt{\tlin{A \otimes B}{\taff}} 
      %   = \dnt{\tlin{A}{\taff}} \otimes \dnt{\tlin{B}{\taff}}
      %   ; \lambda 
      \\
    \boxed{\dnt{\tlin{A}{\trel}}
      : \mc{C}_1^\tint(\dnt{A}, \dnt{A} \otimes \dnt{A})}
      \\
      \dnt{\tlin{A}{\trel}} = \ms{split}(\dnt{A})
      % \dnt{\tlin{X}{\trel}} = \ms{split}(X) \qquad
      % \dnt{\tlin{A \otimes B}{\trel}} = 
      %   \dnt{\tlin{A}{\trel}} \otimes \dnt{\tlin{B}{\trel}}
      %   ;\alpha;\dnt{A} \otimes \sigma \otimes \dnt{B};\alpha 
      \\
    \boxed{\dnt{\csplits{\Gamma}{\Delta}{\Xi}}
      : \mc{C}_1^\tint(\dnt{\Gamma}, \dnt{\Delta} \otimes \dnt{\Xi})}
      \\
    \dnt{\csplits{\cdot}{\cdot}{\cdot}} = \lambda^{-1} \qquad
    \dnt{\csplits
      {\Gamma, \thyp{x}{A}{q}}
      {\Delta, \thyp{x}{A}{q}}
      {\Xi}} 
      = \dnt{\csplits{\Gamma}{\Delta}{\Xi}} \otimes \dnt{A};\alpha;\dnt{\Delta} \otimes \sigma;\alpha 
      \\
    \dnt{\csplits
      {\Gamma, \thyp{x}{A}{q}}
      {\Delta}
      {\Xi, \thyp{x}{A}{q}}} 
      = \dnt{\csplits{\Gamma}{\Delta}{\Xi}} \otimes \dnt{A};\alpha 
      \\
    \dnt{\csplits
      {\Gamma, \thyp{x}{A}{q}}
      {\Delta}
      {\Xi, \thyp{x}{A}{q}}} \
      = \dnt{\Gamma} \otimes \dnt{\tlin{A}{\trel}};
        \alpha;
        \dnt{\Delta} \otimes \sigma \otimes \dnt{A};
        \alpha 
        \\
    \dnt{\csplits
      {\Gamma, \thyp{x}{A}{q}}
      {\Delta}
      {\Xi}}
      = \dnt{\csplits{\Gamma}{\Delta}{\Xi}}
        \otimes \dnt{\tlin{A}{\taff}}
      ; \rho 
      \\
    \boxed{\dnt{\cwk{\Gamma}{\Delta}}
      : \mc{C}_1^\tint(\dnt{\Gamma}, \dnt{\Delta})} 
      \\
      \dnt{\cwk{\Gamma}{\Delta}} 
      = \dnt{\csplits{\Gamma}{\Delta}{\cdot}};\rho 
      \\
    \boxed{\dnt{\lwk{\ms{L}}{\ms{K}}}
      : \mc{C}_1^\tint(\dnt{\ms{L}}, \dnt{\ms{K}})} 
      \\
    \dnt{\lwk{\cdot}{\cdot}} = \ms{id} \qquad
    \dnt{\lwk
      {\ms{L}, \llhyp{\ell}{\Gamma}{A}}
      {\ms{K}, \llhyp{\ell}{\Delta}{A}}}
      = \dnt{\lwk{\ms{L}}{\ms{K}}} + \upg{(\dnt{\cwk{\Gamma}{\Delta}} \otimes \dnt{A})}{} 
      \\
    \dnt{\lwk{\ms{L}}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}}
      = \dnt{\lwk{\ms{L}}{\ms{K}}};\iota_0
  \end{gather*}
  \caption{Semantics for \isotopessa structural judgements}
  \Description{Semantics for isotope-SSA structural judgements}
  \label{fig:ssa-structural-semantics}
\end{figure}

\begin{figure}
  \begin{gather*}
    \boxed{\dnt{\hasty{\Gamma}{p}{a}{A}{q}}
      : \mc{C}_p^q(\dnt{\Gamma}, \dnt{A})} 
      \\
    \dnt{\hasty{\Gamma}{p}{x}{A}{q}} 
      = \dnt{\cwk{\Gamma}{\thyp{x}{A}{q}}}
      \qquad
    \dnt{\hasty{\Gamma}{p}{f\;a}{B}{q}}
      = \upg{\dnt{\hasty{\Gamma}{1}{a}{A}{q}}}{p}
      ; \ms{inst}_p(f) 
      \\
    \dnt{\hasty{\Gamma}{p}{(a, b)}{A \otimes B}{q}}
      = \upg{(
        \dnt{\csplits{\Gamma}{\Delta}{\Xi}};
        \dnt{\hasty{\Delta}{1}{a}{A}{q}} \otimes
        \dnt{\hasty{\Xi}{1}{b}{B}{q}}
      )}{p} 
      \\
    \dnt{\hasty{\Gamma}{p}{()}{\mb{1}}{q}}
      = \upg{\dnt{\cwk{\Gamma}{\cdot}}}{p} 
      \\
    \dnt{\hasty{\Gamma}{p}{\ctt}{\mb{2}}{q}}
      = \upg{(\dnt{\cwk{\Gamma}{\cdot}};\ctt)}{p}
      \qquad
    \dnt{\hasty{\Gamma}{p}{\cff}{\mb{2}}{q}}
      = \upg{(\dnt{\cwk{\Gamma}{\cdot}};\cff)}{p} 
      \\
    \dnt{\hasty{\Gamma}{p}{\letexpr{x}{a}{e}}{B}{q}}
      = \upg{(
        \dnt{\csplits{\Gamma}{\Delta}{\Xi}}
        ; \dnt{\Delta} \otimes \dnt{\hasty{\Xi}{1}{a}{A}{q}}
      )}{p};\dnt{\hasty{\Delta, \thyp{x}{A}{}}{p}{e}{B}{q}} 
      \\
    \dnt{\hasty{\Gamma}{p}{\letexpr{(x, y)}{a}{e}}{C}{q}}
      = \upg{(
        \dnt{\csplits{\Gamma}{\Delta}{\Xi}}
        ; \dnt{\Delta} \otimes \dnt{\hasty{\Xi}{1}{a}{A \otimes B}{q}}
      )}{p}
      ; 
      \\ \qquad \qquad \qquad \qquad \qquad \alpha
      ; \dnt{\hasty{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{p}{e}{C}{q}}
    %   \\
    % \dnt{\hasty{\Gamma}{0}{\lbsplice{\ell}{A}{t}}{A}{\varnothing}}
    %   = \dnt{\haslb{\Gamma}{t}{\llhyp{\ell}{\cdot}{A}}};\alpha_+
  \end{gather*}
  \caption{Semantics for \isotopessa expressions}
  \Description{Semantics for isotope-SSA expressions}
  \label{fig:ssa-term-semantics}
\end{figure}

We can now interpret a term \(\hasty{\Gamma}{p}{a}{A}{q}\) as a
\textit{morphism} taking \(\dnt{\Gamma}\) into \(\dnt{A}\) with
\textit{centrality} \(p\) and \textit{quantity} \(q\) (i.e., lying in
\(\mc{C}_p^q\)). This means that, as expected, a central term becomes a central
morphism in \(\mc{C}_1\), an affine term becomes an affine morphism, and a
relevant term becomes a relevant morphism, and so, in particular, a pure term
becomes a pure morphsim. Our rules for doing so are given in Figure
\ref{fig:ssa-term-semantics}; we will abuse syntax slightly and write
\(\dnt{\text{"judgement"}}\) for \(\dnt{D}\), where \(D\) is a derivation of
"judgement"; in general, it can be proven that the semantics of any two such
derivations are equal.
% \begin{lemma}[Coherence] \
%   \begin{itemize}
%     \item Given any two derivations \(D_1: \hasty{\Gamma}{p}{a}{A}{q}\), \(D_2: \hasty{\Gamma}{p}{a}{A}{r}\), \(\dnt{D_1} = \dnt{D_2}\)
%     \item Given any two derivations \(D_1, D_2: \haslb{\Gamma}{t}{\ms{L}}\), \(\dnt{D_1} = \dnt{D_2}\)
%     \item Given any two derivations \(D_1, D_2: \csplits{\Gamma}{\Delta}{\Xi}\), \(\dnt{D_1} = \dnt{D_2}\)
%     \item Given any two derivations \(D_1, D_2: \cwk{\Gamma}{\Delta}\), \(\dnt{D_1} = \dnt{D_2}\)
%     \item Given any two derivations \(D_1, D_2: \lwk{\ms{L}}{\ms{K}}\), \(\dnt{D_1} = \dnt{D_2}\)
%   \end{itemize}
% \end{lemma}
Our treatment of \rle{var} is quite standard: the term \(x\) types in \(\Gamma\)
if all other variables in \(\Gamma\) and can be discarded, in which case the
semantics is simply doing so, i.e. \(\dnt{\cwk{\Gamma}{\thyp{x}{A}{q}}}\),
lifting to \(\mc{C}_0\) as necessary. Similarly, to type a function application
\rle{var}, we first take \(\dnt{\hasty{\Gamma}{1}{a}{A}{q}}\), which is
guarnateed to be central, lift it to the appropriate category \(\mc{C}_p\), and
then simply postcompose it with the denotation of \(f\). \rle{pair} is handled
by composing the context split \(\csplits{\Gamma}{\Delta}{\Xi}\) taking the
tensor product of the components \(\dnt{\hasty{\Delta}{1}{a}{A}{q}}\),
\(\dnt{\hasty{\Xi}{1}{b}{B}{q}}\); this is well-defined since both components
are central and \(\mc{C}_1\) is monoidal. We then lift the result to the
appropriate \(\mc{C}_p\). The semantics of \rle{unit} is simply dropping the
entire context; \(\dnt{\cwk{\Gamma}{\cdot}}\), while \rle{true}, and \rle{false}
are each handled by first dropping the context and then postcomposing with the
appropriate constant; in all cases, this is followed by lifting to \(\mc{C}_p\).
Finally, \rle{let} and \rle{let2} are handled by first splitting the context,
then passing the right hand component to the semantics of the bound term
\(\dnt{\hasty{\Xi}{1}{a}{A}{q}}\) or \(\dnt{\hasty{\Xi}{1}{a}{A \otimes
B}{q}}\), and then, lifting to \(\mc{C}_p\) as necessary, after reassociating
with \(\alpha\), passing everything to the inner term \(\dnt{\hasty{\Delta,
\thyp{x}{A}{q}}{p}{e}{B}{q}}\) or \(\dnt{\hasty{\Delta, \thyp{x}{A}{q},
\thyp{y}{B}{q}}{p}{e}{C}{q}}\). Note that this inner term is not necessarily
central, unlike the bound term!

\begin{figure}
  \begin{gather*}
    \boxed{\dnt{\haslb{\Gamma}{t}{\ms{L}}}
      : \mc{C}_0^\varnothing(\dnt{\Gamma}, \dnt{\ms{L}})} \\
    \dnt{\haslb{\Gamma}{\lbrb{\ell}{a}}{\ms{L}}}
      = \upg{(\dnt{\csplits{\Gamma}{\Delta}{\Xi}}
      ; \dnt{\Delta} \otimes \dnt{\hasty{\Xi}{1}{a}{A}{\tint}})}
      ; \dnt{\lwk{\llhyp{\ell}{\Delta}{A}}{\ms{L}}} \\
    \dnt{\haslb{\Gamma}{\ite{e}{s}{t}}{\ms{L}}}
      = \upg{(\dnt{\csplits{\Gamma}{\Delta}{\Xi}}
      ; \dnt{\Delta} \otimes \dnt{\hasty{\Xi}{1}{e}{\mb{2}}{\tint}})}
      ; \delta ;
      \\ \qquad \qquad \qquad \qquad
      [
        \dnt{\haslb{\Delta}{s}{\ms{L}}}, 
        \dnt{\haslb{\Delta}{t}{\ms{L}}} 
      ]
      \\
    \dnt{\haslb{\Gamma}{\letstmt{x}{a}{t}}{\ms{L}}}
      = \upg{\dnt{\csplits{\Gamma}{\Delta}{\Xi}}}
      ; \dnt{\Delta} \otimes \dnt{\hasty{\Gamma}{0}{a}{A}{q}}
      ; \dnt{\haslb{\Delta, \thyp{x}{A}{}}{t}{\ms{L}}}
      \\
    \dnt{\haslb{\Gamma}{\letstmt{(x, y)}{a}{t}}{\ms{L}}}
      = \upg{\dnt{\csplits{\Gamma}{\Delta}{\Xi}}}
      ; \dnt{\Delta} \otimes \dnt{\hasty{\Gamma}{0}{a}{A \otimes B}{q}}
      ; \alpha
      ; \dnt{\haslb{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{t}{\ms{L}}}
      \\
    \dnt{\haslb{\Gamma}{\ewhere{t}{L}}{\ms{K}}}
      = \dnt{\haslb{\Gamma}{t}{\ms{L}}}
      ; \dnt{\lhaslb{\ms{L}}{L}{\ms{K}}}^\dagger
      \\
    \boxed{\dnt{\lhaslb{\ms{L}}{L}{\ms{K}}}
      : \mc{C}_0^\varnothing(\dnt{\ms{L}}, \dnt{\ms{K}} + \dnt{\ms{L}})}
      \\
    \dnt{\lhaslb{\ms{L}}{L, \lwbranch{\ell}{x: A}{t}}{\ms{K}}}
      = \dnt{\lhaslb{\ms{L}}{L}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}}
      ; [\dnt{\ms{K}} + \dnt{\haslb{\Gamma, \thyp{x}{A}{}}{t}{\ms{L}}}, \iota_1]
      \\
    \dnt{\lhaslb{\ms{L}}{\cdot}{\ms{L}}}
      = \iota_0
  \end{gather*}
  \caption{Semantics for \isotopessa blocks}
  \Description{Semantics for isotope-SSA blocks}
  \label{fig:ssa-block-semantics}
\end{figure}

Similarly to terms, \textit{blocks} \(\haslb{\Gamma}{t}{\ms{L}}\) are
interpreted as \textit{impure} morphisms from \(\dnt{\Gamma}\) to \(\ms{L}\):
here \(\Gamma\) represents all the variables live on entry to the block, while
\(\ms{L}\) represents all the external blocks which \(t\) might branch to, each
annotated with their live variables and their arguments; our rules are given in
Figure \ref{fig:ssa-block-semantics}. Our semantics for break statements
\(\lbrb{\ell}{a}\) is as follows: we begin by splitting the context into live
variables which pass into dominated blocks and variables used in the argument
via \(\dnt{\csplits{\Gamma}{\Delta}{\Xi}}\). We then pass the latter into the
semantics of the argument (which is pure, and therefore must be lifted to
\(\mc{C}_0\)) \(\dnt{\hasty{\Gamma}{1}{a}{A}{\tint}}\), with the result and live
variables re-associated and passed into
\(\lwk{\llhyp{\ell}{\Delta}{A}}{\ms{L}}\) to obtain the desired target. The
semantics of if-statements are similarly given by first splitting the context
via \(\dnt{\csplits{\Gamma}{\Delta}{\Xi}}\) into variables live in the branches,
\(\Delta\), and variables appearing in the discriminant, \(\Xi\); the latter of
which are passed into the lifted pure morphism
\(\dnt{\hasty{\Delta}{1}{b}{\mb{2}}{\tint}}\). The "if" statement itself is then
simply an application of the distributor \(\delta: A \otimes (\mb{1} + \mb{1})
\to A + A\), with each branch passed to the semantics of its corresponding block
\(\dnt{\haslb{\Xi}{s}{\ms{L}}}\), \(\dnt{\haslb{\Xi}{t}{\ms{L}}}\). The
semantics of \rle{let-blk} and \rle{let2-blk} are analogous to those of
\rle{let} and \rle{let2} in the term case, except that now the bound term \(a\)
is allowed to be impure (and therefore in particular does not need to be
upgraded).

Finally, we must give semantics for \rle{where}. To do so, we begin by giving a
interpreting a control-flow graph \(\lhaslb{\ms{L}}{L}{\ms{K}}\) as a morphism
from \(\dnt{\ms{L}}\) to \(\dnt{\ms{K}} + \dnt{\ms{L}}\), where the left-hand
side of the coproduct denotes branching to one of the target labels in
\(\ms{K}\) (the \textit{output}), while the right-hand side of the coproduct
denotes a recursive call to a label in \(\ms{L}\) (the \textit{recursive
input}). In particular, \rle{nil-br} simply treats a call to one of \(\ms{L}\)
as a call to a \textit{target label}, passing it through the left injection
unchanged. On the other hand \rle{cons-br} is a little bit more complicated:
first we execute \(\dnt{\lhaslb{\ms{L}}{L}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}}\),
which will give us either an output \(\dnt{\ms{K}, \lhyp{\ell}{\Gamma}{A}}\), or
a recursive input \(\dnt{\ms{L}}\), the latter of which we can pass through
unchanged as a recursive input. In the output case, we can interpret this as
either a value of \(\dnt{\ms{K}}\) or \(\dnt{\Gamma} \otimes \dnt{A}\); the
former we can pass the value through unchanged as output; in the second case, we
can pass this in as input to \(\dnt{\haslb{\Gamma,
\thyp{x}{A}{q}}{t}{\ms{L}}}\), which will give us a value of \(\ms{L}\) we can
pass in as recursive input. Note that, even if \(t\) targets an output label in
\(\ms{K}\), we still return a \textit{recursive input}, not an output; in this
case, the \textit{next} iteration will return it as an output. We can now quite
easily state the semantics of \rle{where}: we simply precompose the semantics of
the entry block \(\dnt{\haslb{\Gamma}{t}{\ms{L}}}\) with the \textit{iterated}
semantics of the associated control-flow graph,
\(\dnt{\lhaslb{\ms{L}}{L}{\ms{K}}}^\dagger\).

\subsection{Semantic Metatheory}

Now that we have a semantics for \isotopessa programs, we can show that it
matches the syntax by showing that the syntactic and semantic notions of
substitution coincide. We begin by giving a semantics for the judgement
\(\issubst{\gamma}{\Theta}{\Gamma}\), which we will interpret as a pure morphism
taking \(\dnt{\Theta}\) to \(\dnt{\Gamma}\), defined via the rules in Figure
\ref{fig:ssa-subst-semantics}, which can simply be viewed as computing all
\(\gamma(x)\) for \(\thyp{x}{A}{q}\) in \(\Gamma\). In particular, we interpret
\rle{subst-nil} as the identity function, and \rle{subst-cons} by splitting the
context via \(\dnt{\csplits{\Theta}{\Theta_\Gamma}{\Theta_x}}\), and then
passing \(\dnt{\Theta_\Gamma}\) to the remainder of the substitution
\(\dnt{\issubst{\gamma}{\Gamma}{\Theta_\Gamma}}\) and \(\dnt{\Theta_x}\) through
\(\dnt{\hasty{\Theta_x}{1}{\gamma(x)}{A}{q}}\) to yield \(\dnt{A}\). We give a
similar semantics for label-substitutions: \rle{lb-subst-nil} is interpreted as
the zero morphism \(0_{\dnt{\ms{K}}}: \mc{C}_0^\varnothing(\mb{0},
\dnt{\ms{K}})\), while \rle{lb-subst-cons} takes in the coproduct \(\dnt{\ms{L}}
+ \dnt{\Gamma} \otimes \dnt{A}\), sending the former to
\(\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{A}}}\) and the latter to the denotation of
\(\mc{L}(\lbl{\ell}, x)\).

\begin{figure}
  \begin{gather*}
    \boxed{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}: \mc{C}_1^\varnothing(\dnt{\Theta}, \dnt{\Gamma})} 
    \\
    \dnt{\issubst{\cdot}{\cdot}{\cdot}} = \ms{id}
    \qquad
    \dnt{\issubst{\gamma}{\Theta}{\Gamma, \thyp{x}{A}{q}}}
      = \dnt{\csplits{\Theta}{\Theta_\Gamma}{\Theta_x}};\dnt{\issubst{\gamma}{\Theta_\Gamma}{\Gamma}} \otimes \dnt{\hasty{\Theta_x}{1}{a}{A}{q}}
    \\
    \boxed{\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}: \mc{C}_0^\varnothing(\dnt{\ms{L}}, \dnt{\ms{K}})} 
    \\
    \dnt{\lbsubst{\mc{L}}{\cdot}{\ms{K}}} = 0_{\ms{K}}
    \qquad
    \dnt{\lbsubst{\mc{L}}{\ms{L}, \llhyp{\ell}{\Gamma}{A}}{\ms{K}}}
    = [\dnt{\lbsubst{\ms{L}}{\ms{L}}{\ms{K}}}, \dnt{\haslb{\Gamma, \thyp{x}{A}{}}{\mc{L}(\lbl{\ell}, x)}{\ms{K}}}]
  \end{gather*}
  \caption{Semantics for \isotopessa substitutions}
  \Description{Semantics for isotope-SSA substitutions}
  \label{fig:ssa-subst-semantics}
\end{figure}

Given this, we can now state the theorem of semantic substitution, which
essentially states that the semantics of a substituted term or block is the same
as the semantics of that term or block preceded by let-bindings corresponding to
the variables in the substitution, which is itself the same as the semantics of
the original term or block precomposed with the semantics of the substitution as
given in Figure \ref{fig:ssa-subst-semantics}. Similarly, the semantics of a
\textit{label-substituted} block is the same as the semantics of the original
block \textit{postcomposed} with the semantics of that label-substitution.
Stated formally, this becomes:
\begin{theorem}[Semantic Substitution] 
  Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
  \begin{itemize}
    \item For all \(\hasty{\Gamma}{p}{a}{A}{q}\), 
    \[
      \upg{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}}{p}
      ;\dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
      = \dnt{\hasty{\Theta}{p}{\exprletsubst{\substctx{\gamma}{\Gamma}}{a}}{A}{q}}
      = \dnt{\hasty{\Theta}{p}{[\gamma]a}{A}{q}}
    \]
    \item For all 
      \(\haslb{\Gamma}{t}{\ms{L}}\), 
      \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\), 
      \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\),
    \[
      \upg{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}}{}
      ; \dnt{\haslb{\Gamma}{t}{\ms{L}}}
      = \dnt{\haslb{\Theta}{\exprletsubst{\substctx{\gamma}{\Gamma}}}{\ms{K}}} 
      ; \dnt{\lbsubst{\substlbs{\gamma}{\ms{K}}}{\ms{K}}{\ms{L}}}
      = \dnt{\haslb{\Theta}{[\gamma]t}{\ms{K}}} 
      ; \dnt{\lbsubst{\substlbs{\gamma}{\ms{K}}}{\ms{K}}{\ms{L}}}
    \]
    \item For all 
      \(\lhaslb{\ms{W}}{L}{\ms{L}}\), 
      \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\),
      \(\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}\),
      \[
        \dnt{\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}}
        ; \dnt{\lhaslb{\ms{W'}}{[\gamma]L}{\ms{K}}}
        = \dnt{\lhaslb{\ms{W}}{L}{\ms{L}}}
        ; \dnt{\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}}
        + \dnt{\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}}
      \]
  \end{itemize}  
  Similarly, given \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\),
  \begin{itemize} 
    \item For all \(\haslb{\Gamma}{t}{\ms{L}}\), we have
    \[
      \dnt{\haslb{\Gamma}{t}{\ms{L}}};\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
      = \dnt{\haslb{\Gamma}{[\mc{L}]t}{\ms{K}}}
    \]
    \item For all \(\lhaslb{\ms{W}_{\ms{L}}}{L}{\ms{L}}\), we have
    \[
      \dnt{\lhaslb{\ms{W}_{\ms{L}}}{L}{\ms{L}}};\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
      + \dnt{\lbsubst{\mc{L}}{\ms{W}_{\ms{L}}}{\ms{W}_{\ms{K}}}}
      = \dnt{\lbsubst{\mc{L}}{\ms{W}_{\ms{L}}}{\ms{W}_{\ms{K}}}};\dnt{\lhaslb{\ms{W}_{\ms{K}}}{[\mc{L}]L}{\ms{K}}}
    \]
    where all variables in \(\ms{W}\) not in \(\ms{L}\) are fresh w.r.t \(\mc{L}\).
  \end{itemize}
\end{theorem}

\subsection{Optimisations}

Because we have a denotational semantics, it is easy to validate equations
between terms, blocks, and CFGs.

Even the basic proof of the soundness of substitution already lets us prove the
soundness of many fundamental compiler optimisations, such as global value
numbering (let-introduction for pure expressions), some forms of dead code
elimination (i.e., eliminating unused affine bindings), and strength reduction
with loop hoisting (let-introduction for relevant expressions).

Substitution only applies to central terms, because those are the only terms for
which substitution is sound. Furthermore, there are many other optimisations
that follow from the beta-eta theory of the language, as well as specific
properties of model-dependent operations (such as memory operations). In the
appendix, we state a general rewriting theorem which lifts equations between
terms into a full congruence. This lets us justify optimisations such as scalar
replacement of aggregates, eliminating redundant writes, reordering independent
memory operations, and memory-to-register transformations.

Because we have a type-theoretic presentation of control-flow graphs, we also
have a substitution principle and equational theory for them. This lets us show
the soundness of inlining, loop unrolling, loop fusion, dead-code elimination,
and other optimisations which change the structure of the control-flow graph.

Note that we have not yet proved anything about the dataflow analyses used to
find the rewrites -- we have only shown that the rewrites themselves are
semantically justified.

\section{Concrete Models}

In this section, we will build on the bare-bones models given in Section
\ref{sec:semantics} to develop more sophisticated models capable of underpinning
\isotopessa dialects supporting features such as weak memory and separation
logic.

\subsection{Trace Models}

\label{ssec:trace-monads}

We will begin by giving a useful foundation for a wide class of \isotopessa
models: the "trace models" over \(\mb{Set}\), which accumulate a (potentially
infinite) nondeterministic trace of monoidal "events". These will later come in
handy for reasoning about output behaviour and weak memory.

We begin with some definitions:
\begin{definition}[Stream Action]
  A \textbf{stream action} of a monoid \(M\) on a set \(I\) equipped with a
  monoid action \(\cdot: M \times I \to I\) is a function \(\Theta: M^\omega \to
  I\) mapping (infinite) streams of \(M\) to elements of \(I\) satisfying
  \(
    m \cdot \Theta(\ell) = \Theta(m \colon \ell)
  \),
  where \(m \colon \ell\) denotes the stream formed by prefixing \(m\) to the
  stream \(\ell\).
\end{definition}
Some basic examples include:
\begin{enumerate}
  \item For any monoid \(M\), the unique map \(\Theta: M^\omega \to \mathbf{1}\)
  forms a stream action on the one element set \(\mathbf{1}\).
  % \item In particular, given monoids \(M_i\) acting on sets \(I_i\) with action
  % \(\cdot_i\) and stream actions \(\Theta_i: M_i^\omega \to I_i\), \(\Theta =
  % \langle \Theta_i \circ \ms{fmap}\;\pi_i\rangle_i: (\Pi_iM_i)^\omega \to
  % \Pi_iI_i\) is a stream action compatible with the action \(m \cdot l = (m_i
  % \cdot_i l_i)_i\).
  \item The concatenation map \(\Theta: (A^*)^\omega \to A^\omega\) is a stream
  action of \(A^*\), the set of (finite) lists of \(A\), on \(A^{\leq \omega}\),
  the set of (\textit{potentially} infinite) sequences of \(A\), where \(\ell
  \cdot s\) prepends the \(\ell\) to \(s\).
  % \item The sum map \(\Sigma: \nats^\omega \to \nats \cup \{\infty\}\) is a
  % stream action of \(\nats\) on \(\nats \cup \{\infty\}\), where \(\nats\) acts
  % on \(\nats \cup \{\infty\}\) by addition.
  \item Given a monoid \(M\) acting on sets \(I_i\) with action \(\cdot_i\) and
  stream actions \(\Theta_i: M^\omega \to I_i\), \(\Theta =
  \langle\Theta_i\rangle_i: M^\omega \to \Pi_iI_i\) is a stream action
  compatible with the action \(m \cdot l = (m \cdot_i l_i)_i\).
\end{enumerate}
We define the \textbf{stream center} of \(M\) with respect to \(I\), denoted
\(\mc{SZ}(M, I)\), as the subset of the center of \(M\), \(Z(M)\), consisting of
those elements \(m \in Z(M)\) such that for all \(i \in I\), \(m \cdot i = i\).
For instance, \(\mc{SZ}(M, \mathbf{1}) = Z(M)\), and \(\mc{SZ}(A^*, A^{\leq
\omega}) = \{[\,]\}\). Generally, \(\mc{SZ}(M, I)\) always contains
the unit of \(M\). We can now define the \textit{trace monad} of a stream action as follows:
\begin{definition}[Trace Monad]
  Given a stream action \(\Theta: M^\omega \to I\), we can define the
  \textbf{nondeterministic trace monad} \(\ms{Traces}\;\Theta\;A = \mc{P}^+(A
  \times M + I)\) with unit \(\eta_A = \lambda a.\iota_0 (a, 1)\) and
  multiplication
  \(
    \mu_A\;T = \{[\lambda (a, m). m \cdot a, \ms{id}](t) \mid t \in T\}
  \),
  where the action of \(M\) on traces is given by \(m \cdot \iota_0 (a, m') =
  \iota_0 (a, mm')\) and \(m \cdot \iota_1 t = \iota_1 (m \cdot t)\). We can
  equip this monad with an iteration operator \(f^\dagger(a) = f^\infty(a) \cup
  \bigcup_{n \in \nats}\{\iota_0 (\iota_0\;b, m) \in f_n(a)\} \cup \{\iota_1 t
  \in f_n(a)\}\), yielding an Elgot structure, where
  \[
    f_0 = f, \quad f_{n + 1} = f \gg [\upg{\iota_0}{}, f_n], \quad
    f^\infty(a) = \{\Theta (\lambda n. m_n) \mid a_0 = a \land \forall n, \exists a_{n + 1}. \iota_0 (\iota_1\;a_{n + 1}, m_n) \in f(a_n)\}
  \]
\end{definition}

We now give a few basic definitions relevant to monads over \(\mb{Set}\):
\begin{definition}[Relevant, Affine Monad]
  A monad \(\mb{T}\) on \(\ms{Set}\) is \textbf{relevant} if its Kleisli
  category satisfies \(f;\ms{split} = \ms{split};f \ltimes f = \ms{split};f
  \rtimes f\), where \(\ms{split} = \upg{\lambda a. (a, a)}{}\). \(\mb{T}\) is
  \textbf{affine} if its Kleisli category satisfies \(f;\ms{drop} = \ms{drop}\),
  where \(\ms{drop} = \upg{\lambda a.()}{}\). A monad is \textbf{intuitionistic}
  if it is both relevant and affine.
\end{definition}
\begin{definition}[Submonad] 
  A \textbf{submonad} \(\mb{S}\) of a monad \(\mb{T}\) on \(\ms{Set}\) is a
  collection of sets such that, for all \(A\), \(\mb{S}\;A \subseteq
  \mb{T}\;A\), \(\forall a \in A, \eta\;a \in \mb{S}\;A\), and, for all \(s \in
  \mb{S}\;A\), \(f: A \to \mb{S}\;B\), \(\mb{bind}\;s\;f \in \mb{S}\;A\). We say
  that a submonad of \(\mb{T}\) is \textbf{Elgot} if \(\mb{T}\) is Elgot and,
  for all \(f: A \to \mb{S}(B + A)\), \(f^\dagger: A \to \mb{S}\;B\) where
  \(f^\dagger\) is computed as in \(\mb{T}\). Note that \(\mb{S}\) is also an
  (Elgot) monad in its own right, inheriting \(\mu, \eta\) and \(\cdot^\dagger\)
  from \(\mb{T}\).
\end{definition}
Some important examples of submonads include:
\begin{enumerate}
  \item The \textbf{unit submonad} \(\eta\) of every monad \(\mb{T}\) with
  \(\eta(A) = \{\eta_A\;a \mid a \in A\}\).
  %; similarly, \(\mb{T}\) is itself a submonad of \(\mb{T}\). 
  \item Given submonads \(\mb{S}_i\) of \(\mb{T}\), their elementwise
  intersections \(\bigwedge_i\mb{S}_i = \lambda A. \bigcap_i \mb{S}_i\;A\) are
  submonads, forming a complete meet-semilattice with bottom element \(\eta\)
  and top element \(\mb{T}\),
  % ; this induces a complete lattice in the standard manner.
  \item In particular, the intersection of a relevant/affine/commutative
  submonad and another submonad is relevant/affine/commutative respectively.
  Since every commutative submonad induces a central subcategory of the Kleisli
  category of the parent monad, it follows that a triple of a relevant, affine,
  and commutative submonad make the Kleisli category of a monad \(\mb{T}\) into
  a substructural effectful category with all objects both affine and relevant.
\end{enumerate}
We can use the trace monad described above to derive some interesting examples
of submonads:
\begin{enumerate}
  \item If \(C \subseteq Z(M)\) is submonoid of \(M\), then \(A \mapsto
  \mc{P}^+(A \times C + \mb{0}), \mc{P}^+(A \times C \times I) \subseteq
  \ms{Traces}\;\Theta\;A\) are submonads
  \item In particular, if \(M_C \subseteq Z(M)\) is a central submonoid of \(M\)
  such that \(\forall m \in M_C, \forall i \in I, m \cdot i = i\), then \(A
  \mapsto \mc{P}^+(A \times M_C + \mb{0}) \subseteq \ms{Traces}\;\Theta\;A\) is
  a submonad whose Kleisli category is a central subcategory of the Kleisli
  category of \(\ms{Traces}\;\Theta\;A\). Note \(M_C = \{1\}\) is always such a
  submonoid.
  \label{item:central-submonad}
  \item Similarly, if \(M_R \subseteq M_C\) is a submonoid such that \(\forall m
  \in M_R, m \cdot m = m\), then \(A \mapsto \{S \in \mc{P}^+(A \times M_R + I)
  \mid |S| = 1\} \subseteq \ms{Traces}\;\Theta\;A\) is a \textit{relevant}
  submonad. Note \(M_R = \{1\}\) is always such a submonoid.
  \label{item:relevant-submonad}
  \item Finally, \(A \mapsto \mc{P}^+(A \times \{1\} + \mb{0}) \subseteq
  \ms{Traces}\;\Theta\;A\) is an \textit{affine} submonad.
  \label{item:affine-submonad}
\end{enumerate}
In particular, \ref{item:central-submonad}, \ref{item:relevant-submonad},
\ref{item:affine-submonad} equip the Kleisli category of
\(\ms{Traces}\;\Theta\;A\) with the structure of a substructural effectful
category given choices of \(M_C\), \(M_R\).

\subsection{Heaps, Printing, and Separation Logic}

\label{ssec:separation}

As a concrete example, we can now introduce the \textbf{printing monad}, which
is simply defined as \(\ms{Print}\;A \equiv \ms{Traces}\;\Sigma\), where
\(\Sigma: (\ms{Byte}^*)^\omega \to \ms{Byte}^{\leq \omega}\) denotes
concatenation of bytestrings into a potentially infinite bytestream. This can be
equipped with the structure of a substructural effectful category by choosing
\(M_C = M_R = \{[]\}\), giving us an \isotopessa model supporting the
instructions \(\ms{print} \in \mc{I}^\varnothing_0(\ms{Byte}^*, \mb{1})\) and
\(\ms{rand} \in \mc{I}^\taff_1(\mb{1}, A)\) for \(A\) nonempty, with standard
semantics \(\dnt{\ms{print}} = \lambda b.\{\iota_0 ((), b)\}\),
\(\dnt{\ms{rand}} = \lambda (). \{\iota_0 (a, []) \mid a \in A\}\).

We can make things a little more interesting by adding a heap to the mix,
defining our monad \(\ms{Comp} = \ms{StateT}\;\ms{Heap}\;\ms{Print}\), where
\(\ms{Heap} = \nats \rightharpoonup \nats\) is simply a partial function. The
Kleisli category of this monad can be equipped with the structure of a
substructural effectful category by simply taking the affine/relevant morphisms
to be the lifts of the affine/relevant morphisms in \(\ms{Set}_{\ms{Print}}\);
as it inherits an Elgot structure from that on \(\ms{Set}_{\ms{Print}}\), we
therefore have an \isotopessa model supporting the instructions \(\ms{set}:
\mc{I}^\varnothing_0(\nats \times \nats, \mb{1})\), \(\ms{get}:
\mc{I}^\varnothing_0(\nats, \nats)\), \(\ms{alloc}: \mc{I}^\varnothing_0(\nats,
\nats)\), and \(\ms{free}: \mc{I}^\varnothing_0(\nats, \mb{1})\). We can assign
semantics to \(\ms{set}\) in the standard fashion, with \(\dnt{\ms{set}} =
\lambda (p, v)\;h. \{\iota_0((), [p \mapsto v]h, [])\}\). \(\ms{get}\) is a bit
more tricky, since it is unclear what to do when we try to access uninitialized
memory: one option is simply to return an arbitrary value, with \(\dnt{\ms{get}}
= \lambda p\;h. \{\iota_0(v, h, []) \mid v = h\;p \lor p \notin h\}\); for now,
this will do. Finally, \(\dnt{\ms{alloc}} = \lambda v\;h.\{\iota_0\;(p, h', [])
\mid h' = h \sqcup p \mapsto v\}\) simply fills a random empty heap cell with
the provided value, returning a pointer to the cell, while \(\dnt{\ms{free}} =
\lambda p\;h. \{\iota_0\;((), h \setminus p, [])\}\)

We can use this construction to, building on the ideas in \cite{mellies-ftrs},
show how a simple separation logic based on \cite{reynolds-separation-2002}
slightly adapted to the categorical setting gives us a very rich \isotopessa
model supporting a basic notion of refinement types, while at the same time
giving us a nontrivial notion of nonlinear types.

Given a set \(A\), we can define a heap-predicate \(\Phi\) over \(A\) to be of
type \(A \to \ms{Heap} \to \ms{Prop}\). Taking as objects pairs \((A, \Phi)\),
we can define the category \(\ms{Sep}\) to have morphisms
\begin{equation}
  \ms{Sep}((A, \Phi), (B, \Psi)) = \{f: A \to \ms{Comp}\;B 
    \mid \forall \phi: \ms{Heap} \to \ms{Prop}, \ms{hoare}\;f\;(\Phi \ast \phi)\;(\Psi \ast \phi)\}
\end{equation}
where
\(
  \ms{hoare}\;f\;\Phi\;\Psi = \forall a\;h. \Phi\;a\;h \implies \forall \iota_0\;(b, h') \in f\;a\;h. \Psi\;b\;h'
\)
is the usual partial correctness triple definition and the separating
conjunction of heap-predicates is defined as usual to be 
\begin{equation}
  \begin{gathered}
  \varphi \ast \psi = \lambda h. \exists h_1, h_2. h = h_1 \sqcup h_2 \land \varphi h_1 \land \psi h_2 
  \qquad
  \varphi \ast \Phi = \lambda a. \varphi \ast \Phi\;a
  \qquad
  \Phi \ast \varphi = \lambda a. \Phi\;a \ast \varphi \\
  \Phi \ast \Psi = \lambda (a, b). \Phi\;a \ast \Psi\;b: A \times B \to \ms{Heap} \to \ms{Prop}
  \end{gathered}
\end{equation}
We can naturally define a tensor product on objects \((A, \Phi) \otimes (B,
\Psi) = (A \times B, \Phi \ast \Psi)\), verifying that, taking
\begin{itemize}
  \item Monoidal unit \((\mb{1}, \ms{emp})\), where \(\ms{emp}\) is the
  predicate which holds only for the empty heap
  \item Associators and inverse associators given by \(\alpha_{(A, \Phi), (B,
  \Psi), (C, \Theta)} = \alpha_{A, B, C}\)
  %, which are valid since \(\ast\) is associative
  \item Unitors and inverse unitors given by \(\lambda_{(A, \Phi)} =
  \lambda_A\), \(\rho_{(A, \Phi)} = \rho_A\)
  %, which are valid since \(\ms{emp}\)
  is the unit of \(\ast\)
  \item \((C, \Theta) \otimes f = C \times f\) and \(f \otimes (C, \Theta) = f
  \times C\), which are valid definitions by the frame rule
\end{itemize}
equips \(\ms{Sep}\) with the structure of a premonoidal category. We can give \(\ms{Sep}\) the structure of a substructural effectful category by taking:
\begin{itemize}
  \item Affine objects to be precisely those for which \(\ms{drop} =
  \upg{\lambda a. ()}{} \in \ms{Sep}((A, \Phi), (\mb{1}, \ms{emp}))\), and
  relevant objects to be precisely those for which \(\ms{split} = \upg{\lambda
  a. (a, a)} \in \ms{Sep}((A, \Phi), (A, \Phi * \Phi))\), which for this logic
  corresponds to pairs \((A, \Phi)\) where \(\Phi\) does not depend on the heap.
  \item Central morphisms to be those which do not print, diverge, \textit{or}
  modify the heap. % We will consider how to relax this condition later.
  \item Relevant and affine morphisms to be the lifts of relevant/affine
  morphisms in \(\ms{Print}\)
\end{itemize}
\(\ms{Sep}\) also has coproducts given by \((A, \Phi) + (B, \Psi) = (A + B, [\Phi, \Psi])\), which allow us to equip it with an Elgot structure by taking advantage of the fact that
\begin{equation}
  f \in \ms{Sep}((A, \Phi), (B + A, [\Psi, \Phi])) \implies f^\dagger \in \ms{Sep}((A, \Phi), (B, \Psi))
\end{equation}
This implies that \(\ms{Sep}\) is an \isotopessa model supporting the following
interesting operations:
\begin{itemize}
  \item \(\ms{coe}: \mc{I}^\tint_1((A, \Phi), (A, \Psi))\) where \(\forall a,
  \Phi\;a \implies \Psi\;a\), implemented by the identity morphism
  \item \(\ms{print}: \mc{I}^\varnothing_0((\ms{Byte}^*, \ms{emp}), (\mb{1},
  \ms{emp}))\) and \(\ms{rand}: \mc{I}^\taff_1((\mb{1}, \ms{emp}), (A, \ms{emp}))\) for
  \(A\) nonempty
  \item \(\ms{set}: \mc{I}^\varnothing_0((\nats \times \nats, \lambda (\ell,
  \cdot). \exists v. \ell \mapsto v), (\nats, \lambda \ell. \exists v. \ell
  \mapsto v))\) and \(\ms{get}: \mc{I}^\varnothing_0((\nats \times \nats,
  \lambda (-, \ell). \exists v. \ell \mapsto v), (\nats, \lambda \ell. \exists
  v. \ell \mapsto v))\), which are modified slightly to return the input address.
  \item \(\ms{alloc}: \mc{I}^\varnothing_0((\nats, \ms{emp}), (\nats, \lambda
  \ell. \exists v. \ell \mapsto v))\) and \(\ms{free}: \mc{I}^\varnothing_0((\nats, \lambda \ell. \exists v. \ell \mapsto v), (\mb{1}, \ms{emp}))\)
\end{itemize}
We can clearly see that this is a refinement type system in the sence of
\cite{mellies-ftrs}: in particular, there is a forgetful functor \((A, \Phi)
\mapsto A\) which is simply the identity on morphisms. Unfortunately, this means
that our notion of equality is a little too stringent, since we need to consider
equality on inputs which do not match our preconditions. Thankfully, we can fix
this problem quite easily by quotienting morphisms over the equivalence relation
\begin{equation}
  \forall f, g: \ms{Sep}((A, \Phi), (B, \Psi)).
  f \simeq g \iff \forall a, h. \Phi\;a\;h \implies f\;a\;h = g\;a\;h
\end{equation}
yielding the new category \(\ms{SComp}\). In particular, this allows us to widen
our definition of central morphisms to include all morphisms which do not print
or diverge, including in particular \(\ms{set}, \ms{get}, \ms{alloc}\) and
\(\ms{free}\), which we can in fact mark relevant and affine as well, allowing
us, for example, to make use of intermediate allocations in an otherwise pure
computation.

\subsection{Weak Memory}

Another interesting use of the trace monad to build a sophisticated \isotopessa model is to build a model of TSO-style weak memory based on \citet{sparky}. For simplicity, we will consider an infinite set of named, fixed locations \(x, y, z \in \ms{Loc}\) which are subject to concurrent modification by all processors via TSO reads, writes, and fences.

We begin with the following definitions:
\begin{definition}[Pomset] 
  A \textbf{pomset} \(\alpha\) over a set of actions \(\mc{A}\) with a
  distinguished null action "\textbf{tick}" \(\delta \in \mc{A}\) is a
  \textit{nonempty} partially-ordered \textbf{carrier set} \(P\) such that every
  \(p \in P\) has finitely many predecessors equipped with a mapping \(\alpha: P
  \to \mc{A}\) quotiented under the equivalence relation \(\alpha: P \to \mc{A}
  \simeq \alpha': Q \to \mc{A}\) if, after removing finitely many \(\delta\)
  from \(P\) and \(Q\), there exists a partial order isomorphism between them
  which respects the values of \(\alpha, \alpha'\). A pomset is \textbf{finite}
  if its carrier set is. We will write unordered pomsets using multiset
  notation, e.g. \(\{a, a, b\}\), and linearly ordered pomsets using list
  notation, e.g. \([a, a, b]\).
\end{definition}
(Finite) pomsets form a monoid under sequential composition \(\alpha;\beta\),
which is defined by the function \([\alpha, \beta]: \ms{trim}(P + Q) \to
\mc{A}\), where \(P + Q\) is given the lexicographic ordering and
\(\ms{trim}(P)\) removes all elements of \(P\) with infinitely many
predecessors. They also form a monoid under parallel composition \(\alpha ||
\beta\), which is defined by the function \([\alpha, \beta]: P + Q \to \mc{A}\)
where \(P + Q\) is given the standard partial ordering (with elements of \(P\)
and \(Q\) incomparable). Note in both cases the monoidal unit is \(\{\delta\}\),
since we only consider nonempty pomsets but allow the removal of finitely many
ticks.

Given a partially ordered set \(N\) and a family of pomsets \(\alpha_n\) for \(n
\in N\), we can define their \textit{sum} \(\Sigma_n\alpha_n\) to be given by
the function \((n, a) \mapsto \alpha_n\;a: \ms{trim}(\Sigma_nP_n) \to \mc{A}\),
where the dependent product \(\Sigma_nP_n\) is given the lexicographic order. In
particular, choosing \(N = \nats\) makes \(\Sigma: \ms{Pom}_{\ms{fin}}^\omega
\to \ms{Pom}\) into a stream action w.r.t the sequential composition monoid on
finite pomsets.

We define a \textbf{program order pomset} to be a pomset with \(\mc{A}_{\ms{PO}}
= \mc{A}_w \cup \mc{A}_r \cup \{\delta\}\), where \(\mc{A}_r\) consists of
\textit{reads} of the form \(x = v\) and \(\mc{A}_w\) consists of
\textit{writes} of the form \(x := v\) for locations \(x\) and values \(v \in
\ms{Word}\). We define the \textbf{program order monad} \(\ms{PO}\;A =
\ms{Traces}\;\Sigma\;A\), where \(\Sigma\) is taken as a stream action on finite
program order pomsets given by sequential composition for \(\mc{A}\). This
yields an \isotopessa model with support for concurrent read and write
operations \(\ms{read}_x: \mc{I}^\varnothing_0(\mb{1}, \ms{Word})\),
\(\ms{write}_x: \mc{I}^\varnothing_0(\ms{Word}, \mb{1})\) with semantics
\begin{equation}
  \dnt{\ms{read}_x} = R_x^{\ms{PO}} = \lambda (). \{(v, \{x = v\}) | v \in \ms{Word}\}
  \qquad
  \dnt{\ms{write}_x} = W_x^{\ms{PO}} =  \lambda v. \{((), \{x := v\})\}
\end{equation}
The semantics of \(\ms{read}\) in particular give a hint as to how this model
works: rather than tracking the state of the heap, we simply \textit{emit} a
pomset of the events that would have to happen for a given execution (in the
case of a read, a read event \(x = v\) whenever the read returns \(v\), and in
the case of a write, the single write event \(x := v\) for a write of \(v\)),
and later post-filter on the pomset of an entire program to get a sequentially
consistent execution. Hence, our semantics remains compositional, allowing us to
reason about each program fragment individually, while still allowing us to add
arbitrary external state (e.g. another program concurrently modifying the heap)
before finally considering the execution of an entire program.

On that note, we can define the \textit{parallel execution} of two morphisms as
follows:
\begin{equation}
  \begin{aligned}
  f_0 || f_1 = \lambda (a_0, a_1). 
  & \{\iota_0 ((b_0, b_1), \alpha_0 || \alpha_1) 
    \mid \iota_0 (b_i, \alpha_i) \in f_i\;a_i\} 
  \\ & \cup \{\iota_1 (\alpha_0 || \alpha_1) \mid (\iota_0 (b_0, \alpha_0) \in f_0\;a \lor \iota_0\;\alpha_0 \in f_0\;a_0) \land \iota_1 \alpha_1 \in f_1\;a_1\} 
  \\ & \cup \{\iota_1 (\alpha_0 || \alpha_1) \mid \iota_1\;\alpha_0 \in f_0\;a_0 \land \iota_0 (a_1, \alpha_1) \in f_1\;a_1\}: \ms{Set}_{\ms{PO}}(A \otimes A', B \otimes B')
  \end{aligned}
\end{equation}
It's tempting to try to use this as a tensor product of morphisms to obtain a
monoidal (rather than preomonoidal) category, but unfortunately, sliding still
fails: \((W_x^{\ms{PO}} || \ms{id}) ; (\ms{id} || W_y^{\ms{PO}})\) emits the
pomset \([x := a, y := b]\), while \(\ms{id} || W_y^{\ms{PO}}) ; (W_x^{\ms{PO}}
|| \ms{id})\) emits the pomset \([y := b, x := a]\). 
% If we did want this behaviour, we'd need to have a significantly more
% sophisticated model of composition which effectively takes nontermination into
% account, which we will leave to future work.

To graduate from SC concurrency to a genuine, if maximally simple, weak memory
model, we will to introduce a \textit{buffer} of write actions. We will
implement TSO ordering by buffering all our writes, in which case they are only
visible to the local thread. On the other hand, read events will first attempt
to read from the buffer, and, if there is no corresponding write in the buffer,
will read an arbitrary value, in both cases pushing an event to the global
pomset. At any point, we may choose to \textit{flush} some of the writes in
buffer to the global pomset of events. In particular, we will introduce the set
\(\mc{A}_b = \{(\bufloc{x} := v)\}\) of \textit{buffer write actions}, where an
action \(\bufloc{x} := v\) by a thread denotes adding a write \(x := v\) to the
thread's write buffer. We may then define the set of TSO actions
\(\mc{A}_{\ms{TSO}} = \mc{A}_{\ms{PO}} \cup \mc{A}_b\); a pomset over this set
of actions is called a \textbf{TSO pomset}. A buffer will be defined to be a
list of write actions \(\ms{Buf} = \mc{A}_b^*\), which we will interpret as
linear pomsets over \(\mc{A}_{\ms{TSO}}\) ordered by index (with the empty list
corresponding to \(\{\delta\}\)). In particular, we define the monad \(\ms{TSO}
= \ms{StateT}\;\ms{Buf}\;(\ms{Trace}\;\Sigma)\), where \(\Sigma\) is taken as a
the stream action of finite TSO pomsets on TSO pomsets. We can view
\(\ms{PO}\) as a submonad of this monad.

Given a buffer \(\ms{Buf}\), we can define the result of reads \([\cdot]_x:
\ms{Buf} \to \ms{Word} \sqcup \{\bot\}\) from the buffer by induction to be:
\(
  (L;\{\bufloc{x} := v\})[x] = v,
\),
\(
  (L;\{\bufloc{y} := v\})[x] = L[x]
\),
\(
  [][x] = \bot
\).
Note that writes at the \textit{end} of the buffer are prioritized, since later
writes overwrite earlier ones! The semantics of reads and writes can then be
given by
\begin{equation}
  \begin{aligned}
  \dnt{\ms{read}_x} = R_x^{\ms{TSO}} 
    &= \ms{pflush};(\lambda ()\;L. \{(v, L, \{x = v\}) \mid L[x] = v \lor L[x] = \bot\});\ms{pflush} \\
  \dnt{\ms{write}_x} = W_x^{\ms{TSO}}
    &= \ms{pflush};(\lambda v\;L. \{(v, (L;\{\bufloc{x} := v\}), \{x := v\})\});\ms{pflush}
  \end{aligned}
\end{equation}
where buffer flushing is implemented via the morphism (called \(\ms{split}\) in
\cite{sparky})
\begin{equation}
  \ms{pflush} = \lambda ()\;L. \{((), R, \alpha) | L = \alpha;R\}
  : \ms{Set}_{\ms{TSO}}(\mb{1}, \mb{1})
\end{equation}
To be able to perform synchronization, we will also need to introduce a
\(\ms{fence}: \mc{I}^\varnothing_0(\mb{1}, \mb{1})\) instruction, which simply
causes all actions before the fence to be observed before any actions after the
fence. For \(\ms{TSO}\), implementing this is as simple as flushing the buffer,
i.e., we define
\(
  \dnt{\ms{fence}} = \lambda ()\;L. \{((), [], L;\{\delta\})\}
\).

We can now define the parallel composition of morphisms in a "fork-join" style
as follows: we first flush the buffer completely (i.e., taking it and sticking
it at the beginning of our pomset), then execute \(f\) and \(g\) in parallel
with separate buffers, \textit{filtering out executions which completely flush
the buffer}. Since both threads end up with an empty buffer, the resulting
joined buffer is also empty, giving us the following:
\begin{equation}
  \begin{aligned}
    f_0 || f_1 = \lambda (a_0, a_1)\;L. 
    & \{\iota_0 ((b_0, b_1), [], L;(\alpha_0 || \alpha_1)) 
      \mid \iota_0 (b_i, [], \alpha_i) \in f_i\;a_i\;[]\} 
    \\ & \cup \{\iota_1 (\alpha_0 || \alpha_1) \mid (\iota_0 (b_0, [], \alpha_0) \in f_0\;a \lor \iota_0\;\alpha_0 \in f_0\;a_0\;[]) \land \iota_1 \alpha_1 \in f_1\;a_1\;[]\} 
    \\ & \cup \{\iota_1 (\alpha_0 || \alpha_1) \mid \iota_1\;\alpha_0 \in f_0\;a_0\;[] \land \iota_0 (a_1, [], \alpha_1) \in f_1\;a_1\;[]\}
    \end{aligned}
\end{equation}

\section{Related Work and Discussion}

For space reasons we, restrict our discussion to the most closely related work
on SSA.

Originally, \citet{kelsey-ssa-cps} showed how to embed SSA into a CPS
representation, and \citet{appel-ssa} showed informally how SSA can be seen as a
collection of mutually-tail-recursive procedures.
\citet{chakravarty-functional-ssa-2003} made Appel's observation formal, by
showing how to translate SSA into A-normal form~\cite{anf}, and used it to prove
the correctness of a constant propagation analysis. These translations were for
an untyped source and target language, and
\citet{typed-effect-ssa-rigon-torrens-vasconcellos-20} give a typed translation
from SSA into the lambda calculus, into a type-and-effect system. In contrast,
we directly track the centrality and duplicability of terms, without tracking
specific effects.

All of these papers give SSA semantics by translation -- that is, they defined
the semantics of an SSA program to be the semantics of the image of the
translation, with the idea that the simpler semantics of the functional language
can make proofs easier. The Grail language of \citet{beringer-imp-fun} makes
this idea explicit by giving both a functional and imperative semantics to a
low-level language, and showing the conditions under which they coincide.
\citet{schneider-imp-fun} prove a similar result, and further mechanise the
proof in Coq.

There have also been a number of papers which have studied the semantics of SSA
directly.

\citet{ssa-types-matsuno-ohori-06} give a type theory for what appear to be
ordinary three-address code programs. However, every well-typed program can be
placed into SSA-form by inserting $\phi$-nodes in a fully type-directed way.
This lets them model SSA without any $\phi$-nodes, lets them use the standard
semantics for three-address code.
%
\citet{hua-explicit-ssa-2010} give another type system and direct operational
semantics for SSA, and prove type safety for it.

Many operational semantics for SSA have arisen from compiler verification
efforts. \citet{barthe-compcert-ssa-2014} give an operational semantics as part
of the CompCertSSA project, and give a semantics-preserving translation from
three-address code into SSA. \citet{herklotz-gsa-2023} formalise "gated SSA" and
give semantics-preserving translations between it and ordinary SSA. Going beyond
CompCertSSA, \citet{vellum} have studied the semantics of the LLVM IR itself.

There has been much less work on denotational semantics for SSA, or directly on
its equational theory. \citet{pop-ssa-inout-2009} give an unusual denotational
model of SSA in terms of the iteration structure of a program, which they use to
better understand the loop-closing $\phi$-nodes found both in the gated SSA
representation as well as practical compilers such as GCC.
\citet{garbuzov-structural-cfg-2018} exhibit a correspondence between an
operational semantics for SSA, and an operational semantics for
call-by-push-value~\cite{cbpv}. They use the normal-form bisimulations of
\citet{lassen-bisim} to derive an equational theory for use in justifying
optimisations. Their operational semantics is effect-free, which means that it
cannot justify optimisations such as eliminating redundant writes.

Our work is the first to validate the equational theory of SSA against a weak
memory semantics. We focused on denotational models of weak memory, since they
make it easier to establish equational properties. However, we need a number of
features of the semantics which are sometimes difficult to establish in the weak
memory world. First, sequencing must be associative (i.e., $(t_1; t_2); t_3
\equiv t_1; (t_2; t_3)$).  Second, the semantics must support loops, including
the equation that a while-loop is equal to its unrolling. Finally, sums need to
be distributive (e.g., in the boolean case, $(A \otimes \mb{2}) \simeq (A +
A)$).

As a result, we focused on the TSO semantics of \citet{sparky}, which validates
the properties we need. For ARM-style concurrency, which admits load buffering,
we only found models which are "near misses": they have a subset of the
properties we need. The event structure model of
\citet{paviotti-modular-relaxed-dep-20} is formulated in terms of step-indexing,
so satisfies all of our needed properties except the loop unrolling property --
step-indexing means it is a refinement rather than an equation. The pomsets with
preconditions model of \citet{jagadeesan-pwp-20} does not exactly satisfy the
associativity of semicolon, and while the pomsets with predicate transformers
model of \citet{leaky-semicolon} does satisfy associativity, it does not yet
support loops.

We do not think it would be prohibitively difficult to extend these models to
support the required features, and we believe our results suggest that it would
be have a high payoff: those features would suffice to build a model of our IR,
giving us tight results validating the transformations compilers want against
the low-level specification of the actual machine.

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\clearpage

\appendix

% \section{Proofs}

% \subsection{Weakening and Substitution}

% \begin{theorem}[Weakening] \
%   \begin{itemize}
%     \item If \(\cwk{\Gamma}{\Delta}\) and \(q \subseteq r\),
%     \(\hasty{\Delta}{p}{a}{A}{r}\), then \(\hasty{\Gamma}{p}{a}{A}{q}\)
%     \item If \(\cwk{\Gamma}{\Delta}\), \(\haslb{\Delta}{t}{\ms{L}}\), then
%     \(\haslb{\Gamma}{t}{\ms{L}}\)
%     \item If \(\lwk{\ms{L}}{\ms{K}}\), \(\haslb{\Gamma}{t}{\ms{L}}\) then
%     \(\haslb{\Gamma}{t}{\ms{K}}\)
%   \end{itemize}
% \end{theorem}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{lemma}[Substitution Splitting] 
%   If \(\csplits{\Gamma}{\Delta}{\Xi}\) and \(\issubst{\gamma}{\Theta}{\Gamma}\),
%   then there exist \(\Theta_\Delta\), \(\Theta_\Xi\) such that:
%   \begin{itemize}
%     \item \(\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}\)
%     \item \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\)
%     \item \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\)
%   \end{itemize}
% \end{lemma}

% \begin{proof} 
%   We proceed by induction on derivations \(\csplits{\Gamma}{\Delta}{\Xi}\). In
%   the \rle{split-nil} case, we have that \(\issubst{\gamma}{\cdot}{\cdot}\) by
%   \rle{subst-nil}. For all other cases, since \(\issubst{\gamma}{\Theta}{\Gamma,
%   \thyp{x}{A}{q}}\), we have by \rle{subst-cons} that
%   \(\csplits{\Theta}{\Theta_\Gamma}{\Theta_x}\),
%   \(\issubst{\gamma}{\Theta_\Gamma}{\Gamma}\),
%   \(\hasty{\Theta_\Gamma}{1}{\gamma(x)}{A}{q}\). Since
%   \(\csplits{\Gamma}{\Delta}{\Xi}\), by induction, we have that
%   \(\csplits{\Theta_\Gamma}{\Theta_\Delta}{\Theta_\Xi}\) where \(\Theta_\Delta\)
%   and \(\Theta_\Xi\) have the desired properties above.
%   \begin{itemize}
%     \item \rle{split-left}: by left split association, we have that there exists
%     \(\Theta_{\Delta, \thyp{x}{A}{q}}\), \(\csplits{\Theta}{\Theta_{\Delta,
%     \thyp{x}{A}{q}}}{\Theta_\Xi}\), \(\csplits{\Theta_{\Delta,
%     \thyp{x}{A}{q}}}{\Theta_\Delta}{\Theta_x}\). Using the fact that
%     \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\), we can conclude by
%     \rle{subst-cons} that \(\issubst{\gamma}{\Theta_{\Delta,
%     \thyp{x}{A}{q}}}{\Delta, \thyp{x}{A}{q}}\); the desired result hence
%     follows.
%     \item \rle{split-right}: by right split association, we have that there
%     exists \(\Theta_{\Xi, \thyp{x}{A}{q}}\),
%     \(\csplits{\Theta}{\Theta_\Delta}{\Theta_{\Xi, \thyp{x}{A}{q}}}\),
%     \(\csplits{\Theta_{\Xi, \thyp{x}{A}{q}}}{\Theta_\Xi}{\Theta_x}\). Using the
%     fact that \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\), we can conclude by
%     \rle{subst-cons} that \(\issubst{\gamma}{\Theta_{\Xi, \thyp{x}{A}{q}}}{\Xi,
%     \thyp{x}{A}{q}}\); the desired result hence follows.
%     \item \rle{split-dup}: since \(\tlin{A}{\trel}\), it follows by
%     \rle{subst-cons} that \(\tlin{\Theta_x}{\trel}\), and hence that
%     \(\csplits{\Theta_x}{\Theta_x}{\Theta_x}\). Therefore, by split association,
%     we have that there exist \(\Theta_{\Delta, \thyp{x}{A}{q}}\), \(\Theta_{\Xi,
%     \thyp{x}{A}{q}}\) such that \(\csplits{\Theta}{\Theta_{\Delta,
%     \thyp{x}{A}{q}}}{\Theta_{\Xi, \thyp{x}{A}{q}}}\). Using the facts that
%     \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\),
%     \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\), we can conclude by \rle{subst-cons}
%     that \(\issubst{\gamma}{\Theta_{\Delta, \thyp{x}{A}{q}}}{\Delta,
%     \thyp{x}{A}{q}}\), \(\issubst{\gamma}{\Theta_{\Xi, \thyp{x}{A}{q}}}{\Xi,
%     \thyp{x}{A}{q}}\); the desired result hence follows.
%     \item \rle{split-drop}: since \(\tlin{A}{\taff}\), it follows by
%     \rle{subst-cons} that \(\tlin{\Theta_x}{\taff}\), and hence that
%     \(\cwk{\Theta_x}{\cdot}\). Therefore, by split association,
%     \(\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}\), giving the desired result.
%   \end{itemize}
% \end{proof}

% \begin{lemma}[Substitution] 
%   Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
%   \begin{itemize}
%     \item If \(\hasty{\Gamma}{p}{a}{A}{q}\), then \(\hasty{\Theta}{p}{[\gamma]a}{A}{q}\)
%     \item If \(\haslb{\Gamma}{t}{\ms{L}}\), then there exists \(\ms{K}\) such that \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) and \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\)
%     \item If \(\lhaslb{\ms{W}}{L}{\ms{L}}\), then there exists \(\ms{K}, \ms{W}'\) such that 
%     \begin{itemize}
%       \item \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) 
%       \item \(\lbsubst{\substlbs{\gamma}{\ms{W}'}}{\ms{W}}{\ms{W}'}\) 
%       \item \(\lhaslb{\ms{W}'}{[\gamma]L}{\ms{K}}\)
%     \end{itemize}
%   \end{itemize}
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{lemma}[Label Substitution] 
%   Given \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\), 
%   \begin{itemize}
%     \item For all \(\haslb{\Gamma}{t}{\ms{L}}\), \(\haslb{\Gamma}{[\mc{L}]t}{\ms{K}}\)
%     \item For all \(\lhaslb{\ms{W}}{L}{\ms{L}}\), \(\lhaslb{\ms{W}}{[\mc{L}]L}{\ms{K}}\)
%   \end{itemize}
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \subsection{Semantic Metatheory}

% \begin{lemma}[Semantic Upgrade]
%   For all \(\Gamma, a, A\), if \(\hasty{\Gamma}{1}{a}{A}{q}\), then
%   \[\dnt{\hasty{\Gamma}{0}{a}{A}{q}} = \upg{\dnt{\hasty{\Gamma}{1}{a}{A}{q}}}{}\]
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{lemma}[Semantic Weakening] 
%   If \(\cwk{\Gamma}{\Delta}\), then 
%   \begin{itemize}
%     \item For all \(\hasty{\Delta}{p}{a}{A}{q}\),
%     \[
%       \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
%       ; \dnt{\hasty{\Delta}{p}{a}{A}{q}}
%       = \dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
%     \]
%     \item For all \(\haslb{\Delta}{t}{\ms{L}}\),
%     \[
%       \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
%       ; \dnt{\haslb{\Delta}{t}{\ms{L}}}
%       = \dnt{\haslb{\Gamma}{t}{\ms{L}}}  
%     \]
%   \end{itemize}

%   Similarly, if \(\lwk{\ms{L}}{\ms{K}}\) and \(\haslb{\Gamma}{t}{\ms{L}}\), then
%   \[
%     \dnt{\haslb{\Gamma}{t}{\ms{L}}}
%     ; \dnt{\lwk{\ms{L}}{\ms{K}}}
%     = \dnt{\haslb{\Gamma}{t}{\ms{K}}}
%   \]
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{lemma}[Substitution Coherence] \
%   \begin{itemize}
%     \item Given any two derivations \(D_1: \issubst{\gamma}{\Theta}{\Gamma}\), \(D_2: \issubst{\gamma'}{\Theta}{\Gamma}\) such that \(\forall x \in \Theta, \gamma(x) = \gamma'(x)\), we have \(\dnt{D_1} = \dnt{D_2}\)
%     \item Given any two derivations \(D_1: \lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\), \(D_2: \lbsubst{\mc{L}'}{\ms{L}}{\ms{K}}\) such that \(\forall \lbl{\ell} \in \ms{L}, \mc{L}(x) = \mc{L}'(x)\), we have \(\dnt{D_1} = \dnt{D_2}\)
%   \end{itemize}
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{lemma}[Semantic Substitution Splitting] 
%   Given \(\csplits{\Gamma}{\Delta}{\Xi}\) and \(\issubst{\gamma}{\Delta}{\Xi}\),
%   if \(\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}\) and \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\), \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\), then
%   \[
%     \dnt{\issubst{\gamma}{\Theta}{\Gamma}};\dnt{\csplits{\Gamma}{\Delta}{\Xi}}
%     = \dnt{\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}};
%       \dnt{\issubst{\gamma}{\Theta_\Delta}{\Delta}}
%       \otimes \dnt{\issubst{\gamma}{\Theta_\Xi}{\Xi}}
%   \]
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{theorem}[Semantic Substitution] 
%   Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
%   \begin{itemize}
%     \item For all \(\hasty{\Gamma}{p}{a}{A}{q}\), 
%     \[
%       \upg{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}}{p}
%       ;\dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
%       = \dnt{\hasty{\Theta}{p}{[\gamma]a}{A}{q}}
%     \]
%     \item For all 
%       \(\haslb{\Gamma}{t}{\ms{L}}\), 
%       \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\), 
%       \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\),
%     \[
%       \upg{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}}{}
%       ; \dnt{\haslb{\Gamma}{t}{\ms{L}}}
%       = \dnt{\haslb{\Theta}{[\gamma]t}{\ms{K}}} 
%       ; \dnt{\lbsubst{\substlbs{\gamma}{\ms{K}}}{\ms{K}}{\ms{L}}}
%     \]
%     \item For all 
%       \(\lhaslb{\ms{W}}{L}{\ms{L}}\), 
%       \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\),
%       \(\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}\),
%       \[
%         \dnt{\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}}
%         ; \dnt{\lhaslb{\ms{W'}}{[\gamma]L}{\ms{K}}}
%         = \dnt{\lhaslb{\ms{W}}{L}{\ms{L}}}
%         ; \dnt{\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}}
%         + \dnt{\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}}
%       \]
%   \end{itemize}
%   Similarly, given \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\),
%   \begin{itemize} 
%     \item For all \(\haslb{\Gamma}{t}{\ms{L}}\), we have
%     \[
%       \dnt{\haslb{\Gamma}{t}{\ms{L}}};\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
%       = \dnt{\haslb{\Gamma}{[\mc{L}]t}{\ms{K}}}
%     \]
%     \item For all \(\lhaslb{\ms{W}_{\ms{L}}}{L}{\ms{L}}\), we have
%     \[
%       \dnt{\lhaslb{\ms{W}_{\ms{L}}}{L}{\ms{L}}};\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
%       + \dnt{\lbsubst{\mc{L}}{\ms{W}_{\ms{L}}}{\ms{W}_{\ms{K}}}}
%       = \dnt{\lbsubst{\mc{L}}{\ms{W}_{\ms{L}}}{\ms{W}_{\ms{K}}}};\dnt{\lhaslb{\ms{W}_{\ms{K}}}{[\mc{L}]L}{\ms{K}}}
%     \]
%     where all variables in \(\ms{W}\) not in \(\ms{L}\) are fresh w.r.t \(\mc{L}\).
%   \end{itemize}
% \end{theorem}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

\section{Rewriting}

\label{apx:rewriting}

\subsection{Syntax}

To properly be able to reason about congruence and congruence-based
optimistion, we need to introduce the concept of \textit{terms with holes},
which are defined using the grammar in Figure
\ref{fig:blocks-with-holes-grammar}. This simply extends the three main
syntactic categories with \textit{holes} of the form \(\lhole{X}\),
\(\lhole{T}\), \(\lhole{W}\). To \textit{type} terms with holes, we need to
introduce the concept of a \textit{hole-context} \(H\), which is a list of holes
annotated as corresponding to one of:
\begin{itemize}
  \item A term \(\tyhole{\lhole{X}}{\Delta}{p}{A}{q}\) of type \(A\), quantity
  \(q\), and purity \(p\) using the variables in \(\Delta\). Such a hole can be
  filled by any term \(a\) satisfying \(\hasty{\Delta}{p}{a}{A}{q}\)
  \item A block \(\blkhole{\lhole{T}}{\Delta}{\ms{L}}\) targeting \(\ms{L}\)
  using the variables in \(\Delta\). Such a hole can be filled by any block
  \(t\) satisfying \(\haslb{\Delta}{t}{\ms{L}}\).
  \item A control-flow graph \(\cfghole{\lhole{W}}{\ms{L}}{\ms{K}}\) with input
  labels \(\ms{L}\) and output labels \(\ms{K}\). Such a hole can be filled by
  any CFG satisfying \(\lhaslb{\ms{L}}{L}{\ms{K}}\).
\end{itemize}

\begin{figure}
  \begin{grammar}
    <\(\mhole{a}, \mhole{b}, \mhole{c}, \mhole{e}\)> ::= \(\lhole{X}\)
    \;|\; \(x\) 
    \;|\; \(f\;\mhole{a}\)
    \;|\; \((\mhole{a}, \mhole{b})\) 
    \;|\; \(()\) 
    \;|\; \(\ctt\) 
    \;|\; \(\cff\)
    \;|\; \(\letexpr{x}{\mhole{a}}{\mhole{e}}\)
    \;|\; \(\letexpr{(x, y)}{\mhole{a}}{\mhole{e}}\)
    % \alt \(\lbsplice{\ell}{x: A}{\mhole{t}}\)

    <\(\mhole{s}, \mhole{t}\)> ::= \(\lhole{T}\) 
    \;|\; \(\lbrb{\ell}{\mhole{a}}\) 
    \;|\; \(\ite{\mhole{e}}{\mhole{s}}{\mhole{t}}\)
    \;|\; \(\letstmt{x}{\mhole{a}}{\mhole{t}}\)
    \;|\; \(\letstmt{(x, y)}{\mhole{a}}{\mhole{t}}\)
    \;|\; \(\ewhere{\mhole{t}}{\mhole{L}}\)

    <\(\mhole{L}\)> ::= \(\lhole{W}\) \;|\; \(\cdot\) \;|\; \(\lwbranch{\ell}{x: A}{\mhole{t}}, \mhole{L}\)

    <\(H\)> ::= \(\cdot\) 
    \;|\; \(H, \tyhole{\lhole{X}}{\Gamma}{p}{A}{q}\)
    \;|\; \(H, \blkhole{\lhole{T}}{\Gamma}{\ms{L}}\)
    \;|\; \(H, \cfghole{\lhole{W}}{\ms{L}}{\ms{K}}\)
  \end{grammar}
  \caption{Grammar for \isotopessa expressions and blocks with holes}
  \Description{Grammar for isotope-SSA blocks with holes}
  \label{fig:blocks-with-holes-grammar}
\end{figure}

We introduce new typing judgements for terms, blocks, and CFGs with holes in
Figure \ref{fig:ssa-holes-judgements}. In each case, they correspond to the
judgement that, given a collection \(\mc{H}\) terms, blocks, and CFGs to fill
each hole in \(H\), would satisfy the corresponding judgement without holes in
\ref{fig:ssa-judgements}.

\TODO{note on rewrites}

\begin{figure}
  \begin{center}        
    \begingroup
    \renewcommand{\arraystretch}{1.5}
    \setlength{\tabcolsep}{2em}
    \begin{tabular}{rl}
        \multicolumn{1}{c}{Judgment} & \multicolumn{1}{c}{Meaning} \\ \hline
        \(\isrw{\mc{H}}{I}{H}\) & \(\mc{H}\) is a rewrite from \(I\) to \(H\),
        i.e. fills every hole in \(H\) modulo \(I\) \\
        \(\mhasty{H}{\Gamma}{p}{\mhole{a}}{A}{q}\) & Given
        \(\isrw{\mc{H}}{\cdot}{H}\), \(\hasty{\Gamma}{p}{[\mc{H}]\mhole{a}}{A}{q}\) \\
        \(\mhaslb{H}{\Gamma}{\mhole{t}}{\ms{L}}\) & Given
        \(\isrw{\mc{H}}{\cdot}{H}\), \(\haslb{\Gamma}{[\mc{H}]\mhole{t}}{\ms{L}}\) \\
        \(\mlhaslb{H}{\ms{L}}{\mhole{L}}{\ms{K}}\) & Given
        \(\isrw{\mc{H}}{\cdot}{H}\),
        \(\haslb{\Gamma}{[\mc{H}]\mhole{t}}{\ms{L}}\) \\
    \end{tabular}
    \endgroup
  \end{center}
  \caption{Typing judgements for \isotopessa terms, blocks, and CFGs with holes}
  \Description{Typing judgements for isotope-SSA terms, blocks, and CFGs with
  holes}
  \label{fig:ssa-holes-judgements}
\end{figure}

\begin{figure}
  \begin{gather*}    
    \prftree[r]{\rle{hole-tm}}
      {\tyhole{\lhole{X}}{\Delta}{p}{A}{q}}
      {\cwk{\Gamma}{\Delta}}
      {\mhasty{H}{\Gamma}{p}{\lhole{X}}{A}{q}} \qquad
    \prftree[r]{\rle{var}}
      {\cwk{\Gamma}{\thyp{x}{A}{q}}}
      {\mhasty{H}{\Gamma}{p}{x}{A}{q}} \qquad
    \prftree[r]{\rle{app}}
      {f \in \mc{I}_p^q(A, B)}
      {\mhasty{H}{\Gamma}{1}{\mhole{a}}{A}{q}}
      {\mhasty{H}{\Gamma}{p}{f\;\mhole{a}}{B}{q}} \\
    \prftree[r]{\rle{pair}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\mhasty{H}{\Delta}{1}{\mhole{a}}{A}{q}}
      {\mhasty{H}{\Xi}{1}{\mhole{b}}{B}{q}}
      {\mhasty{H}{\Gamma}{p}{(\mhole{a}, \mhole{b})}{A \otimes B}{q}} \\
    \prftree[r]{\rle{unit}}
      {\cwk{\Gamma}{\cdot}}
      {\mhasty{H}{\Gamma}{p}{()}{\mb{1}}{q}} \qquad
    \prftree[r]{\rle{true}}
      {\cwk{\Gamma}{\cdot}}
      {\mhasty{H}{\Gamma}{p}{\ctt}{\mb{2}}{q}} \qquad
    \prftree[r]{\rle{false}}
      {\cwk{\Gamma}{\cdot}}
      {\mhasty{H}{\Gamma}{p}{\cff}{\mb{2}}{q}} \\
    \prftree[r]{\rle{let}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\mhasty{H}{\Delta, \thyp{x}{A}{}}{p}{\mhole{e}}{B}{q}}
      {\mhasty{H}{\Xi}{1}{\mhole{a}}{A}{q}}
      {\mhasty{H}{\Gamma}{p}{\letexpr{x}{\mhole{a}}{\mhole{e}}}{B}{q}} \qquad
    % \prftree[r]{\rle{blk}}
    %   {\mhaslb{H}{\Gamma}{\mhole{t}}{\llhyp{\ell}{\cdot}{A}}}
    %   {\mhasty{H}{\Gamma}{0}{\lbsplice{\ell}{A}{\mhole{t}}}{A}{\varnothing}} 
      \\
    \prftree[r]{\rle{let2}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\mhasty{H}{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{p}{\mhole{e}}{C}{q}}
      {\mhasty{H}{\Xi}{1}{\mhole{a}}{A \otimes B}{q}}
      {\mhasty{H}{\Gamma}{p}{\letexpr{(x, y)}{\mhole{a}}{\mhole{e}}}{C}{q}}
  \end{gather*}
  \caption{Typing rules for \isotopessa terms with holes}
  \Description{Typing rules for isotope-SSA terms with holes}
  \label{fig:ssa-term-holes-typing}
\end{figure}

\begin{figure}
  \begin{gather*}    
    \prftree[r]{\rle{hole-blk}}
      {\blkhole{\lhole{T}}{\Delta}{\ms{L}} \in H}
      {\cwk{\Gamma}{\Delta}}
      {\mhaslb{H}{\Gamma}{\lhole{T}}{\ms{L}}} 
      \qquad
    \prftree[r]{\rle{br}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\lwk{\llhyp{\ell}{\Delta}{A}}{\ms{L}}}
      {\mhasty{H}{\Xi}{1}{\mhole{a}}{A}{\tint}}
      {\mhaslb{H}{\Gamma}{\lbrb{\ell}{\mhole{a}}}{\ms{L}}} 
      \\
    \prftree[r]{\rle{ite}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\mhasty{H}{\Xi}{1}{\mhole{e}}{\mb{2}}{\tint}}
      {\mhaslb{H}{\Delta}{\mhole{s}}{\ms{L}}}
      {\mhaslb{H}{\Delta}{\mhole{t}}{\ms{L}}}
      {\mhaslb{H}{\Gamma}{\ite{\mhole{e}}{\mhole{s}}{\mhole{t}}}{\ms{L}}} \\
    \prftree[r]{\rle{let-blk}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\mhaslb{H}{\Delta, \thyp{x}{A}{}}{\mhole{t}}{\ms{L}}}
      {\mhasty{H}{\Xi}{0}{\mhole{a}}{A}{q}}
      {\mhaslb{H}{\Gamma}{\letstmt{x}{\mhole{a}}{\mhole{t}}}{\ms{L}}} \\
    \prftree[r]{\rle{let2-blk}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\mhaslb{H}{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{\mhole{t}}{\ms{L}}}
      {\mhasty{H}{\Xi}{0}{\mhole{a}}{A \otimes B}{q}}
      {\mhaslb{H}{\Gamma}{\letstmt{(x, y)}{\mhole{a}}{\mhole{t}}}{\ms{L}}} \\
    \prftree[r]{\rle{where}}
      {\mhaslb{H}{\Gamma}{\mhole{t}}{\ms{L}}}
      {\mlhaslb{H}{\ms{L}}{\mhole{L}}{\ms{K}}}
      {\mhaslb{H}{\Gamma}{\ewhere{\mhole{t}}{\mhole{L}}}{\ms{K}}} 
      \qquad
    \prftree[r]{\rle{hole-cfg}}
      {\cfghole{\lhole{W}}{\ms{L}}{\ms{K}} \in H}
      {\lwk{\ms{M}}{\ms{K}}}
      {\mlhaslb{\ms{L}}{H}{\lhole{W}}{\ms{K}}}
      \\
    \prftree[r]{\rle{nil-br}}
      {\mlhaslb{H}{\ms{L}}{\cdot}{\ms{L}}} \qquad
    \prftree[r]{\rle{cons-br}}
      {\mlhaslb{H}{\ms{L}}{\mhole{L}}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}}
      {\mhaslb{H}{\Gamma, \thyp{x}{A}{}}{\mhole{t}}{\ms{L}}}
      {\mlhaslb{H}{\ms{L}}{\mhole{L}, \lwbranch{\ell}{x: A}{\mhole{t}}}{\ms{K}}}
  \end{gather*}
  \caption{Typing rules for \isotopessa blocks with holes}
  \Description{Typing rules for isotope-SSA blocks with holes}
  \label{fig:ssa-block-holes-typing}
\end{figure}

\begin{figure}
  \begin{gather*}
    \boxed{H: \ms{Hole} \to \ms{MExpr} \sqcup \ms{MBlock} \sqcup \ms{MCfg}}
    \\
    \prftree[r]{\rle{rw-nil}}
      {\isrw{\mc{H}}{H}{\cdot}}
    \qquad
    \prftree[r]{\rle{rw-tm}}
      {\isrw{\mc{H}}{I}{H}}
      {\mhasty{I}{\Gamma}{p}{\mc{H}(\lhole{X})}{A}{q}}
      {\isrw{\mc{H}}{I}{H, \tyhole{\lhole{X}}{\Gamma}{p}{A}{q}}}
    \\
    \prftree[r]{\rle{rw-blk}}
      {\isrw{\mc{H}}{I}{H}}
      {\mhaslb{I}{\Gamma}{\mc{H}(\lhole{T})}{\ms{L}}}
      {\isrw{\mc{H}}{I}{H, \blkhole{\lhole{T}}{\Gamma}{\ms{L}}}}
    \qquad
    \prftree[r]{\rle{rw-cfg}}
      {\isrw{\mc{H}}{I}{H}}
      {\mlhaslb{I}{\ms{L}}{\mc{H}(\lhole{W})}{\ms{K}}}
      {\isrw{\mc{H}}{I}{H, \cfghole{\lhole{L}}{\ms{L}}{\ms{K}}}}
  \end{gather*}
  \caption{Typing rules for \isotopessa rewrites}
  \Description{Typing rules for isotope-SSA rewrites}
  \label{fig:rewrite-typing}
\end{figure}

The typing rules for terms, blocks, and CFGs with holes are exactly the same as
the corresponding rules without holes, modulo the additional rules
\rle{hole-tm}, \rle{hole-blk}, and \rle{hole-cfg} to type the holes themselves;
they are given in full in Figures \ref{fig:ssa-term-holes-typing} and
\ref{fig:ssa-block-holes-typing}. This formalism allows us to state the
\textit{syntactic rewriting} theorem, which says, in essence, that filling holes
with valid terms yields valid terms.
\begin{theorem}[Rewriting]
  Given \(\isrw{\mc{H}}{H}{I}\), we have
  \begin{itemize}
    \item If \(\mhasty{H}{\Gamma}{p}{a}{A}{q}\), then \(\mhasty{I}{\Gamma}{p}{[\mc{H}]a}{A}{q}\)
    \item If \(\mhaslb{H}{\Gamma}{t}{\ms{L}}\), then \(\mhaslb{I}{\Gamma}{[\mc{H}]t}{\ms{L}}\)
  \end{itemize}
  Furthermore, we have that 
  \begin{itemize}
    \item \(\mhasty{\cdot}{\Gamma}{p}{a}{A}{q} \iff \hasty{\Gamma}{p}{a}{A}{q}\)
    \item \(\mhaslb{\cdot}{\Gamma}{t}{\ms{L}} \iff \haslb{\Gamma}{t}{\ms{L}}\)
  \end{itemize}
\end{theorem}

\subsection{Semantics}

We can interpret a hole context semantically as the \textit{set} of all possible
tuples of morphisms which would typecheck in place of the holes in the context,
as specified in Figure \ref{fig:holes-semantics}.
\begin{figure}
  \begin{align*}
      \\
    \dnt{H, \tyhole{\lhole{X}}{\Gamma}{p}{A}{q}}
      &= \dnt{H} \times \mc{C}_p^q(\dnt{\Gamma}, \dnt{A}) 
      \\
    \dnt{H, \blkhole{\lhole{T}}{\Gamma}{\ms{L}}}
      &= \dnt{H} \times \mc{C}_0(\dnt{\Gamma}, \dnt{\ms{L}}) 
      \\
    \dnt{H, \cfghole{\lhole{W}}{\ms{L}}{\ms{K}}}
      &= \dnt{H} \times \mc{C}_0(\dnt{\ms{L}}, \dnt{\ms{K}} + \dnt{\ms{L}}) 
      \\
    \dnt{\cdot} 
      &= \mb{1}
  \end{align*}
  \caption{Semantics for \isotopessa holes}
  \Description{Semantics for isotope-SSA holes}
  \label{fig:holes-semantics}
\end{figure}
We can now quite easily give semantics for terms, blocks, and CFGs with holes
which simply correspond to "filling in" any holes with the appropriate morphism
from the hole context, with all other rules interpreted identically to their
regular variants: this is done in Figures \ref{fig:ssa-term-holes-semantics} and
\ref{fig:ssa-block-holes-semantics}.

\begin{figure}
  \begin{gather*}
    \boxed{
      \dnt{\mhasty{H}{\Gamma}{p}{\mhole{a}}{A}{q}}
      : \dnt{H} \to \mc{C}_p^q(\dnt{\Gamma}, \dnt{A})}
      \\
    \sorry
  \end{gather*}
  \caption{Semantics for \isotopessa terms with holes}
  \Description{Semantics for isotope-SSA terms with holes}
  \label{fig:ssa-term-holes-semantics}
\end{figure}

\begin{figure}
  \begin{gather*}
    \boxed{
      \dnt{\mhaslb{H}{\Gamma}{\mhole{t}}{\ms{L}}}
      : \dnt{H} \to \mc{C}_0^\varnothing(\dnt{\Gamma}, \dnt{\ms{L}})}
      \\
    \sorry
      \\  
    \boxed{
      \dnt{\mlhaslb{H}{\ms{L}}{\mhole{L}}{\ms{K}}}
      : \dnt{H} \to \mc{C}_0^\varnothing(\dnt{\ms{L}}, \dnt{\ms{K}} + \dnt{\ms{L}})}
      \\
    \sorry
  \end{gather*}
  \caption{Semantics for \isotopessa blocks with holes}
  \Description{Semantics for isotope-SSA blocks with holes}
  \label{fig:ssa-block-holes-semantics}
\end{figure}

\begin{figure}
  \begin{gather*}
    \boxed{
      \dnt{\isrw{\mc{H}}{I}{H}}
      : \dnt{I} \to \dnt{H} 
    }
    \\
    \dnt{\isrw{\mc{H}}{I}{\cdot}}\;i = ()
    \qquad
    \dnt{\isrw{\mc{H}}{I}{H, \tyhole{\lhole{X}}{\Gamma}{p}{A}{q}}}\;i
    = (\dnt{\isrw{\mc{H}}{I}{H}}\;i, \dnt{\mhasty{I}{\Gamma}{p}{\mc{H}(\lhole{X})}{A}{q}}\;i)
    \\
    \dnt{\isrw{\mc{H}}{I}{H, \blkhole{\lhole{T}}{\Gamma}{\ms{L}}}}
    = (\dnt{\isrw{\mc{H}}{I}{H}}\;i, \dnt{\mhaslb{I}{\Gamma}{\mc{H}(\lhole{T})}{\ms{L}}}\;i)
    \\
    \dnt{\isrw{\mc{H}}{I}{H, \cfghole{\lhole{L}}{\ms{L}}{\ms{K}}}}
    = (\dnt{\isrw{\mc{H}}{I}{H}}\;i, \dnt{\mlhaslb{I}{\ms{L}}{\mc{H}(\lhole{W})}{\ms{K}}}\;i)
  \end{gather*}
  \caption{Semantics for \isotopessa rewrites}
  \Description{Semantics for isotope-SSA rewrites}
  \label{fig:rewrite-semantics}
\end{figure}

% \begin{lemma}[Rewriting Coherence] \
%   \begin{itemize}
%     \item Given any two derivations \(D_1: \mhasty{H}{\Gamma}{p}{a}{A}{q}\), \(D_2: \mhasty{\Gamma}{p}{H}{a}{A}{r}\), \(\dnt{D_1} = \dnt{D_2}\)
%     \item Given any two derivations \(D_1, D_2: \mhaslb{H}{\Gamma}{t}{\ms{L}}\), \(\dnt{D_1} = \dnt{D_2}\)
%     \item Given any two derivations \(D_1, D_2: \isrw{\mc{H}}{I}{H}\), \(\dnt{D_1} = \dnt{D_2}\)
%   \end{itemize}
% \end{lemma}

\begin{theorem}[Semantic Rewriting]
  Given \(\isrw{\mc{H}}{I}{H}\), we have that
  \begin{itemize}
    \item For all \(\mhasty{H}{\Gamma}{p}{a}{A}{q}\), \(m \in \dnt{I}\), we have
    \[
      \dnt{\mhasty{I}{\Gamma}{p}{[\mc{H}]a}{A}{q}}(m)
      = \dnt{\mhasty{H}{\Gamma}{p}{a}{A}{q}}(\dnt{\isrw{\mc{H}}{I}{H}}(m))
    \]
    \item For all \(\mhaslb{H}{\Gamma}{t}{\ms{L}}\), \(m \in \dnt{I}\), we have
    \[
      \dnt{\mhaslb{I}{\Gamma}{[\mc{H}]t}{\ms{L}}}(m)
      = \dnt{\mhaslb{H}{\Gamma}{t}{\ms{L}}}(\dnt{\isrw{\mc{H}}{I}{H}}(m))
    \]
  \end{itemize}

  Furthermore, we have that
  \[
    \dnt{\mhasty{\cdot}{\Gamma}{p}{a}{A}{q}}() = \dnt{\hasty{\Gamma}{p}{a}{A}{q}}
    \qquad
    \dnt{\mhaslb{\cdot}{\Gamma}{t}{\ms{L}}}() = \dnt{\haslb{\Gamma}{t}{\ms{L}}}
  \]
\end{theorem}

% We can further use this example to get a substructural \textit{monoidal} category by considering a primitive form of separation logic: let's introduce a grammar of predicates \(\varphi, \psi ::= \top, \ms{S}, \bot\) for programs which cannot access state, can access state, and are invalid, respectively. The separating conjunction operator on these predicates is then given by
% \begin{equation}
%   \top * \varphi = \varphi * \top = \varphi
%   \qquad
%   \bot * \varphi = \varphi * \bot = \bot
%   \qquad
%   \ms{S} * \ms{S} = \bot
% \end{equation}
% We'll then define a category \(\ms{SRel}\) with objects of the form \((A, \varphi)\) and morphisms
% \begin{equation}
%   \begin{gathered}
%     \boxed{
%       \ms{SRel}((A, \varphi), (B, \psi)) 
%       \subseteq A \to \ms{S} \to \mc{P}(A \times S) 
%     }
%     \\
%     \begin{aligned}
%       \ms{SRel}((A, \top), (B, \top)) &= \{f \in A \to S \to \mc{P}(A \times S) \mid \exists g. f = \lambda a, s. g(a) \times \{s\}\} \\
%       \ms{SRel}((A, \ms{S}), (B, \ms{S})) &= A \to S \to \mc{P}(A \times S) \\
%       \ms{SRel}((A, \varphi), (B, \psi)) &= \{\lambda a, s. \varnothing\} \quad \text{otherwise}
%     \end{aligned}
%   \end{gathered}
% \end{equation}
% Note that it is precisely the \textit{central} morphisms \(\ms{C}_0(A, B)\) which lie in \(\ms{SRel}((A, \top), (B, \top))\). We can then define the tensor product on objects as \((A, \varphi) \otimes (B, \psi) = (A \otimes B, \varphi * \psi)\), and, for \(f \in \mathsf{SRel}((A, \varphi), (B, \psi))\), and tensor product functors
% \begin{equation}
%   \begin{aligned}
%   f \otimes (C, \phi) &= f \otimes C,
%   & (C, \phi) \otimes f &= C \otimes f
%   && \text{if}\; \varphi * \phi \neq \bot, \psi * \phi \neq \bot \\
%   f \otimes (C, \phi) &= \lambda a, s. \varnothing,
%   & (C, \phi) \otimes f &=  \lambda a, s. \varnothing
%   && \text{otherwise}
%   \end{aligned} 
% \end{equation}
% Lifting associators, unitors, and symmetries from \(\mc{C}_0\) in the obvious manner (setting them to \(\lambda a, s. \varnothing\) where necessary), we can verify that \(\ms{SRel}\) is a \textit{premonoidal} category. we can see that it is in fact \textit{monoidal} by nothing that, for any \(f, g\), \(f \ltimes g = f \rtimes g\) since either \(f\) is central, \(g\) is central, or their product is the empty morphism \(\lambda a, s. \varnothing\). Hence, to get a \textit{substructural} monoidal category, we simply need to define:
% \begin{equation}
%   \begin{gathered}
%   \ms{Aff}(\ms{SRel}) = \ms{Rel}(\ms{SRel}) = \{(A, \top) | A \in \ms{Set}\} \neq |\ms{SRel}|
%   \\
%   \begin{aligned}
%     \ms{SRel}^\varnothing((A, \phi), (B, \psi))
%     &= \ms{SRel}((A, \phi), (B, \psi)) \\
%     \ms{SRel}^{\{\ms{a}\}}((A, \phi), (B, \psi)) 
%     &= \{f \in \ms{SRel}((A, \phi), (B, \psi)) \mid \forall a, s. |f(a, s)| \geq 1 \lor \phi \neq \psi \lor \phi = \bot \} \\
%     \ms{SRel}^{\{\ms{r}\}}((A, \phi), (B, \psi)) 
%     &= \{f \in \ms{SRel}((A, \phi), (B, \psi)) \mid \forall a, s. |f(a, s)| \leq 1 \lor \phi \neq \psi \lor \phi = \bot \} \\
%     \implies \ms{SRel}^{\{\ms{a}, \ms{r}\}}((A, \phi), (B, \psi)) 
%     &= \{f \in \ms{SRel}((A, \phi), (B, \psi)) \mid \forall a, s. |f(a, s)| = 1 \lor \phi \neq \psi \lor \phi = \bot \} \\
%   \end{aligned}
%   \end{gathered}
% \end{equation}
% We can verify that the associators, unitors, and symmetry lie in \(\ms{SRel}^{\{\ms{a}, \ms{r}\}}\) and that all other desired axioms are satisfied.

% \TODO{relate separation logic trick to \cite{promonad}, \cite{linear-state-usage}, \cite{mellies-ftrs}}

% \TODO{clean up, and generalize to arbitrary premonoidal categories (w/ initial object? how to deal w/ coproducts?)}

\end{document}
\endinput