### 2023-11-17 pldi24/paper.tex
Tap

%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0

\documentclass[acmsmall,screen,review]{acmart}

\usepackage{syntax}
\renewcommand{\syntleft}{\normalfont\itshape}
\renewcommand{\syntright}{\normalfont\itshape}

\usepackage{prftree}

\usepackage{listings}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{fancyvrb}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
%    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}

\newcounter{todos}
\newcommand{\TODO}[1]{{
  \stepcounter{todos}
  \begin{center}\large{\textcolor{red}{\textbf{TODO \arabic{todos}:} #1}}\end{center}
}}
\newcommand{\sorry}{\textcolor{red}{\textbf{sorry}}}

\newcommand{\todo}[1]{\stepcounter{todos} \textcolor{red}{TODO \arabic{todos}: #1}}

% Math fonts
\newcommand{\mc}[1]{\ensuremath{\mathcal{#1}}}
\newcommand{\mb}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\ms}[1]{\ensuremath{\mathsf{#1}}}

% Math
\newcommand{\nats}{\mathbb{N}}

% Syntax atoms
\newcommand{\lbl}[1]{{`#1}}
\newcommand{\lto}{\Rightarrow}
\newcommand{\ctt}{\ms{tt}}
\newcommand{\cff}{\ms{ff}}

% Syntax
\newcommand{\letexpr}[3]{\ensuremath{\ms{let}\;#1 = #2\;\ms{in}\;#3}}
\newcommand{\letstmt}[3]{\ensuremath{\ms{let}\;#1 = #2; #3}}
\newcommand{\brb}[2]{\ms{br}\;#1\;#2}
\newcommand{\lbrb}[2]{\brb{\lbl{#1}}{#2}}
\newcommand{\ite}[3]{\ms{if}\;#1\;\{#2\}\;\ms{else}\;\{#3\}}
\newcommand{\ewhere}[2]{\ms{then}\;#1\;\ms{where}\;#2}
\newcommand{\where}[2]{#1\;\ms{where}\;#2}
\newcommand{\wbranch}[3]{#1(#2) \lto #3}
\newcommand{\lwbranch}[3]{\wbranch{\lbl{#1}}{#2}{#3}}
\newcommand{\bsplice}[3]{#1(#2)\;\{#3\}}
%\newcommand{\lbsplice}[3]{\bsplice{\lbl{#1}}{#2}{#3}}
\newcommand{\csplits}[3]{#1 \mapsto #2;#3}
\newcommand{\cwk}[2]{#1 \mapsto #2}
\newcommand{\lwk}[2]{#1 \rightsquigarrow #2}
\newcommand{\tlin}[2]{#2 \subseteq \ms{lin}(#1)}
\newcommand{\ltlin}[3]{#3 \subseteq \ms{lin}(#1) \cap #2}
\newcommand{\thyp}[3]{#1: {#2}^{#3}}
\newcommand{\lhyp}[3]{#1[#2](#3)}
\newcommand{\llhyp}[3]{\lhyp{\lbl{#1}}{#2}{#3}}
\newcommand{\rle}[1]{{\scriptsize\textsf{#1}}}
\newcommand{\taff}{{\{\ms{a}\}}}
\newcommand{\trel}{{\{\ms{r}\}}}
\newcommand{\tint}{{\{\ms{a}, \ms{r}\}}}
\newcommand{\hasty}[5]{#1 \vdash_{#2} #3: {#4}^{#5}}
\newcommand{\haslb}[3]{#1 \vdash #2 \rhd #3}
\newcommand{\lhaslb}[3]{#1 \vdash #2 \rhd #3}
\newcommand{\issubst}[3]{#1: #2 \mapsto #3}
\newcommand{\lbsubst}[3]{#1: #2 \rightsquigarrow #3}
\newcommand{\exprletsubst}[2]{{#1};{#2}}
\newcommand{\stmtletsubst}[2]{{#1};{#2}}
\newcommand{\mhole}[1]{{#1}^?}
\newcommand{\lhole}[1]{?#1}
\newcommand{\mhasty}[6]{#1;#2 \vdash_{#3} #4: {#5}^{#6}}
\newcommand{\mhaslb}[4]{#1;#2 \vdash #3 \rhd #4}
\newcommand{\mlhaslb}[4]{#1;#2 \vdash #3 \rhd #4}
\newcommand{\tyhole}[5]{#1: #2 \mapsto_{#3} {#4}^{#5}}
\newcommand{\blkhole}[3]{#1: #2 \mapsto #3}
\newcommand{\cfghole}[3]{#1: #2 \mapsto #3}
\newcommand{\substctx}[2]{{#1}^{#2}}
\newcommand{\substlbs}[2]{{#1}^{#2}}
\newcommand{\restrictsubst}[2]{{#1}_{#2}}
\newcommand{\subsubst}[2]{#1 \subseteq #2}
\newcommand{\isrw}[3]{#1: #2 \mapsto #3}
\newcommand{\mbind}{\mathbin{{>}\hspace{-0.1em}{>}\hspace{-0.1em}{=}}}
% \newcommand{\strictlbsubst}[3]{#1: #2 \rightsquigarrow_= #3}

% Denotational semantics
\newcommand{\dnt}[1]{\llbracket{#1}\rrbracket}
\newcommand{\ednt}[1]{\left\llbracket{#1}\right\rrbracket}
\newcommand{\upg}[2]{{#1}^{\uparrow #2}}

% Weak memory
\newcommand{\bufloc}[1]{\overline{#1}}

% Branding
\newcommand{\isotopessa}{\ms{isotope_{SSA}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2023}
\acmYear{2023}
\acmDOI{XXXXXXX.XXXXXXX}

%%
%% These commands are for a JOURNAL article.
% \acmJournal{JACM}
% \acmVolume{37}
% \acmNumber{4}
% \acmArticle{111}
% \acmMonth{8}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

\begin{document}

\title{Effectful Semantics for Substructural SSA}

\author{Neel Krishnaswami}
\email{nk480@cl.cam.ac.uk}
\orcid{0000-0003-2838-5865}

\author{Jad Ghalayini}
\email{jeg74@cl.cam.ac.uk}
\orcid{0000-0002-6905-1303}

\begin{abstract}
  Static single-assignment form (SSA) is one of the most widely used compiler
  intermediate representations. In this paper, we build on the observation that
  SSA can be seen as a language of basic blocks with arguments (or first-order
  procedures tail-calling one another) to give a type theory for SSA programs.
  Notably, the type system and equational theory of our language is designed to
  support substituting impure terms for variables, which makes it easy to
  validate compiler optimisations such as hoisting assignments out of loops.

  We give a categorical axiomatisation of our type theory and show that our
  equational theory is sound with respect to it. Our categorical semantics
  generalises the notions of premonoidal category and Freyd category to support
  substructural features, including both support for substructural types, as
  well as F\"{u}hrmann's notion of central, copyable and discardable morphisms.
  Finally, we exhibit a number of concrete models of our categorical
  axiomatisation, including in particular a model of TSO-style weak memory based
  on the semantics of \citet{sparky}. This demonstrates that our approach is
  strong enough to connect machine models of concurrency to compiler IRs
  supporting a rich collection of sound rewrites.
\end{abstract}

\begin{CCSXML}
  <ccs2012>
  <concept>
  <concept_id>10003752.10010124.10010131.10010133</concept_id>
  <concept_desc>Theory of computation~Denotational semantics</concept_desc>
  <concept_significance>500</concept_significance>
  </concept>
  <concept>
  <concept_id>10003752.10010124.10010131.10010137</concept_id>
  <concept_desc>Theory of computation~Categorical semantics</concept_desc>
  <concept_significance>500</concept_significance>
  </concept>
  <concept>
  <concept_id>10003752.10003790.10011740</concept_id>
  <concept_desc>Theory of computation~Type theory</concept_desc>
  <concept_significance>500</concept_significance>
  </concept>
  <concept>
  <concept_id>10003752.10003790.10011742</concept_id>
  <concept_desc>Theory of computation~Separation logic</concept_desc>
  <concept_significance>300</concept_significance>
  </concept>
  </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Theory of computation~Denotational semantics}
\ccsdesc[500]{Theory of computation~Categorical semantics}
\ccsdesc[500]{Theory of computation~Type theory}
\ccsdesc[300]{Theory of computation~Separation logic}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{SSA, Categorical Semantics, Elgot Structure, Effectful Category}

% \received{20 February 2007}
% \received[revised]{12 March 2009}
% \received[accepted]{5 June 2009}

\maketitle

\section{Introduction}

Static Single Assignment, or SSA, has been a cornerstone of compiler design ever
since its introduction by \citet{ssa-original,rosen-gvn-1988} in the late 1980s.
Older compiler intermediate representations permitted the same local variable to
be assigned multiple times, whereas SSA requires each variable binding to be
immutable -- they cannot be modified after initialisation. This makes many
dataflow analyses easier to compute, and (just as in functional languages) makes
it much easier to justify compiler optimisations in terms of equational
rewrites.

In fact, the relationship between SSA form and functional programming has been
known for decades: \citet{kelsey-ssa-cps} showed that SSA can be seen as a
subset of continuation-passing style, and \citet{appel-ssa} built on this to
observe that every SSA program can be viewed as a collection of first-order
functions which tail-call one another. This makes the equivalence of the
functional view to the traditional $\phi$-node presentation very easy to see. In
the functional style, each tail call lists its arguments. In contrast, a
$\phi$-node collects together the arguments from every callsite, which lets the
tail-calls/jumps omit their arguments. Thus, the main difference between SSA and
a functional program is where in the program text the arguments to a tail call
are listed.

This "basic blocks with arguments" presentation of SSA has become very popular
in industry, with compiler backends such as Cranelift~\cite{cranelift},
MLIR~\cite{mlir}, and SIL~\cite{SIL} all using this representation. This tight
connection might suggest that the semantics of SSA representations is trivial:
we can just interpret an SSA program as a corresponding functional program, and
then rewrite SSA programs using the equational theory of functional programs.

While this is a perfectly sound thing to do, it is not strong enough to write a
good compiler. Compiler IRs like SSA are fundamentally about effectful programs,
and the equational theory of effectful functional programs is very weak. Instead
of a full substitution rule, the equational theory of impure functional
languages only justifies substituting values for variables (the $\beta$-value
rule). This weak equational rule suffices to validate some basic compiler
optimisations such as constant propagation and copy propagation, but is not
strong enough to justify many other essential compiler optimisations. For
example, hoisting a load out of a loop or conditional involves moving an
effectful term out of its original position in the program (see Equation
\ref{eqn:hoist}). As another example, dead store elimination removes an
effectful term from the program altogether.

%TODO: note that dead store elimination requires refinement, which we're not
%covering in this paper...

These examples illustrate that many optimisations require taking
\emph{effectful} terms and moving them around and substituting them for
variables. Since this is not a generally valid program transformation, we need a
finer classification of programs than merely value and expression to judge when
a substitution is allowed.

A useful taxonomy was introduced by \citet{fuhrmann-direct-1999}, classifying
effects based on (1) whether or not the order of effects mattered, (2) whether
it was sound to discard an effectful operation, and (3) whether it was sound to
duplicate an effectful operation. Suppose we substitute \texttt{print("hello")}
for $x$ in the program below:
\begin{equation}
  \begin{small}
\begin{BVerbatim}[baseline=c]
let x = print("hello");
let y = print("world");
let z = x;   
...
\end{BVerbatim}
  \end{small}
\qquad \qquad \centernot\implies \qquad \qquad
  \begin{small}
\begin{BVerbatim}[baseline=c]
let y = print("world");
let z = print("hello");
...
\end{BVerbatim}
  \end{small}
\end{equation}
In this case, substituting a term for a variable changes the order of I/O
actions, which changes the program's observable behaviour. On the other hand,
suppose we have an effect which increments a counter. Then in the following
program the same transformation is sound, since \(x\) is used exactly once:
\begin{equation}
  \begin{small}
\begin{BVerbatim}[baseline=c]
let x = increment(a);
let y = increment(b);
let z = x;
...
\end{BVerbatim}
  \end{small}
\qquad \qquad \implies \qquad \qquad
  \begin{small}
\begin{BVerbatim}[baseline=c]
let y = increment(b);
let z = increment(a);
...
\end{BVerbatim}
  \end{small}
\end{equation}
F\"{u}hrmann called such commutative effects \emph{central}, which we also call
\emph{linear}. It can also matter whether an effect is performed at all or not.
For example, consider:
\begin{equation}
  \begin{small}
\begin{BVerbatim}[baseline=c]
let x = may_loop();
return 5 
\end{BVerbatim}
  \end{small}
\qquad \qquad \centernot\implies \qquad \qquad
\begin{small}
\begin{BVerbatim}[baseline=c]
return 5
\end{BVerbatim}
\end{small}
\end{equation}
In this example, even though $x$ is not used, the call to \texttt{may_loop()}
cannot be erased, as the observable behaviour of the program can change if it
goes into an infinite loop. On the other hand, if the effectful operation is
merely nondeterministic, then it is safe to discard results:
\begin{equation}
  \begin{small}
\begin{BVerbatim}[baseline=c]
let x = nondet();
return 5 
\end{BVerbatim}
  \end{small}
\qquad \qquad \implies \qquad \qquad
  \begin{small}
\begin{BVerbatim}[baseline=c]
return 5
\end{BVerbatim}
  \end{small}
\end{equation}
F\"{u}hrmann called these operations \emph{discardable}. In this paper, we use
his terminology, and also call such effects \emph{affine}. Now consider the
following example: 
\begin{equation}
  \begin{small}
\begin{BVerbatim}[baseline=c]
let x = nondet();
let y = x;
let z = x;
return y == z 
\end{BVerbatim}
  \end{small}
\qquad \qquad \centernot\implies \qquad \qquad
  \begin{small}
\begin{BVerbatim}[baseline=c]
let y = nondet();
let z = nondet();
return y == z
\end{BVerbatim}
  \end{small}
\end{equation}
In this case, substituting for $x$ duplicates the call to \texttt{nondet()}, and
can change this from a program which always returns \texttt{true} to one which
can return either \texttt{true} or \texttt{false}. On the other hand, if the
effect is the possibility of nontermination, then the substitution is sound.
\begin{equation}
\begin{small}
\begin{BVerbatim}[baseline=c]
let x = may_loop();
let y = x;
let z = x;
return y == z 
\end{BVerbatim}
\end{small}
\qquad \qquad \implies \qquad \qquad
\begin{small}
\begin{BVerbatim}[baseline=c]
let y = may_loop();
let z = may_loop();
return y == z
\end{BVerbatim}
\end{small}
\end{equation}
F\"{u}hrmann called such operations \emph{copyable}. In this paper, we use his
terminology, and also call such effects \emph{relevant}. With these distinction
in place, we can \emph{define} a pure term as one which is all three of central,
copyable, and discardable.

In addition to working with effects, compiler intermediate representations also
have to track accesses to memory. This information is typically gathered via
ad-hoc alias analyses, but a compositional alternative to these analyses is to
track this information via substructural type systems such as linear and affine
types, which associate aliasing and ownership information with individual
values. Having aliasing information justifies important optimisations such as
loop hoisting: 
\begin{equation}
  \begin{small}
\begin{BVerbatim}[baseline=c]
while (e) {
  x = load(a);
  body
}
\end{BVerbatim}
\end{small}
\qquad \qquad \implies \qquad \qquad
\begin{small}
\begin{BVerbatim}[baseline=c]
x = load(a);
while(e) {
  body
}
\end{BVerbatim}
\end{small}
\label{eqn:hoist}
\end{equation}
This transformation is only sound if the body of the loop \texttt{body} does not
access \texttt{a}, which is precisely the kind of frame condition that linear
types can track. Moreover, since we already want to track whether effectful
\emph{expressions} can be duplicated or discarded, the very same machinery can
be reused to track whether \emph{values} are linear/affine/relevant.

Historically, compiler writers have treated the semantics of their IR quite
informally, since the "intended model" of the language is a simple sequential
machine operating over a flat array of bytes. Unfortunately, over time both the
intended model and the actual hardware have drifted away from this simple mental
model. Compiler frameworks have to model concurrent accesses to memory in order
to compile multithreaded code, and furthermore the IR's memory model has to be
sound with respect to the actual weak memory models supported by hardware.

As a result, it is now much harder to judge whether a particular optimisation is
sound or not. Moreover, different kinds of hardware such as x86, GPUs, and ARM
all offer different memory models. This means that we need an abstract semantics
for compiler IRs, which can permit a separation of concerns. Compiler writers
can validate their optimisations against the abstract semantics, and we can
separately show that a particular machine model faithfully implements the
semantics.

In this paper, we show that this is possible. We give a type system for SSA
which uses substructural types to both track effects at a fine-grained level,
and to track ownership of values. We then rigorously justify a variety of
optimisations by reference to an abstract categorical axiomatisation of this
language. We then show that our categorical axioms have a variety of concrete
models, particularly including one of TSO-style weak memory originally
introduced by \citet{sparky}. This lets us argue that a host of SSA-based
optimisations are sound with respect to the x86 memory model.

\paragraph{Contributions}

\begin{itemize}
\item We give a type system for SSA-style programs, which has a full
substitution theorem and equational theory for both expressions and control-flow
graphs. This language includes substructural types, both for controlling effects
and tracking ownership in values. 
\item We give an axiomatic semantics for our language, by showing that it has a
compositional intepretation in any distributive effectful category (a
substructural variant of a Freyd category) with an Elgot structure. We use this
structure to give generic proofs of the soundness of substitution and rewriting
properties.
\item Our abstract semantics justifies a variety of powerful optimisation
techniques, including inlining, loop fusion, hoisting, and strength reduction. 
\item We show how Elgot monads give rise to instances of our categorical
semantics. We use this to derive semantics from a general trace monad, and show
that various monad transformers such as state transformers preserve the needed
Elgot structure. 
\item We use these tools to demonstrate the existence of a variety of concrete
models satisfying these categorical axioms. Starting first from simple languages
such as state plus printing, we also show how more challenging semantics such as
the TSO semantics for weak memory also fit into our framework. This thus gives
rise to an SSA-based IR with support for weak memory operations, which is fully
semantically-justified.
% \item Our weak memory semantics and our results about Elgot monads are
% formalised in the Lean theorem prover.
\end{itemize}

\section{SSA Syntax}

In a standard SSA-based compiler, to represent a function, we begin by
decomposing it into basic blocks, which can be viewed as \textit{labels},
followed by a sequence of \textit{instructions}, each defining a
\textit{variable}, followed by a \textit{terminator}. A program written in this
manner being in "SSA form" simply means that every variable name is defined
exactly once. For example, the program in Figure \ref{fig:3-addr} has 3 basic
blocks, one of which is a distinguished \textit{entry block}, but is
\textit{not} in SSA form, since the variables \(m, n\) and \(i\) are each
modified through subsequent assignments .
% (\(t\) is recomputed on every iteration of the loop, but since it only has one
% definition, it \textit{is} in SSA form).
To convert this program to SSA form, we would traditionally use
\(\phi\)-functions, or, equivalently, annotate each basic block with arguments,
as in Figure \ref{fig:bb-arg}.

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.33\textwidth}
    \begin{small}
    \begin{BVerbatim}[baseline=t, gobble=3]
    'entry:
      m = 0
      n = 1
      brz i 'exit 'loop
    'loop:
      t = add m n
      m = n
      n = t
      i = sub i 1
      brz i 'exit 'loop
    'exit:
      ret m
    \end{BVerbatim}
    \end{small}
    \caption{3-address code}
    \label{fig:3-addr}
  \end{subfigure}
  \begin{subfigure}[t]{0.32\textwidth}
    \begin{small}
    \begin{BVerbatim}[baseline=t, gobble=3]
      'entry:
      m0 = 0
      n0 = 1
      brz N 
        'exit (m0) 
        'loop (i, m0, n0)
    'loop(i0, m1, n1):
      m2 = n1
      n2 = add m1 n1
      i1 = sub i0 1
      brz z1 
        'exit (m2) 
        'loop (i1, m2, n2)
    'exit(m3):
      ret m3
    \end{BVerbatim}
    \end{small}
    \caption{Basic blocks with arguments}
    \label{fig:bb-arg}
  \end{subfigure}
  \begin{subfigure}[t]{0.33\textwidth}
    \begin{small}
    \begin{BVerbatim}[baseline=t, gobble=3]
    let m0 = 0;
    let n1 = 1;
    if eq i 0 {
      br 'exit (m0)
    } else {
      br 'loop (i, m0, n0)
    } where
    'loop(i0, m1, n1) =>
      let m2 = n1;
      let n2 = add m1 n1;
      let i1 = sub i0 1;
      if eq i1 0 {
        br 'exit (m2)
      } else {
        br 'loop (i1, m2, n2)
      }
    'exit(m3) => ret m3
    \end{BVerbatim}
    \end{small}
    \caption{Our syntax}
    \label{fig:our-syn}
  \end{subfigure}
  \caption{A simple program for computing the \texttt{i}th Fibonacci number,
  represented as 3-address code, basic-blocks with arguments, and in our SSA
  syntax} 
  \Description{}
  \label{fig:ssa-examples}
\end{figure}

For our type theoretic presentation of SSA, we will treat a very slight
generalization of the standard basic blocks with arguments syntax, in that we
will support \textit{nested scopes} and, in particular, \textit{if-statements}.
The program in Figure \ref{fig:bb-arg} translated to our syntax is given in
Figure \ref{fig:our-syn}. This does not introduce any new expressivity: an
if-statement can be viewed as syntactic sugar for a conditional branch to one of
two anonymous basic blocks, while nested scopes simply \textit{restrict}
branching.

We parametrize our syntax by a set of \textit{base types} \(X \in \mc{T}\) and,
for each pair of types \(A, B\), a set of \textit{instructions} \(f \in
\mc{I}^q_p(A, B)\) with \textit{centrality} \(p = 1\) if \textit{central} or \(p
= 0\) otherwise and \textit{quantity} \(q \subseteq \tint\). As discussed above,
an instruction is considered \textit{central} if execution order does not
matter. The quantity \(q\) indicates whether it is safe to duplicate an
instruction, or to delete it an if its result is unused. The instruction is
\textit{affine} if \(\ms{a} \in q\), \textit{relevant}, and \(\ms{r} \in q\),
and, if both affine and relevant, \textit{intuitionistic}. An instruction which
is, central, relevant, and affine is called \textit{pure}. We assume that \(p
\leq p'\) and \(q \subseteq r\) then \(\mc{I}_p^q(A, B) \supseteq
\mc{I}_{p'}^{r}(A, B)\). The set of types is generated by base types \(X\), the
\textit{unit type} \(\mb{1}\), the \textit{boolean type} \(\mb{2}\), and tensor
products of types \(A \otimes B\). We write \(\otimes\) (rather than \(\times\))
to underscore the fact that types are allowed to be substructural. For each base
type \(X\), if we are given a linearity \(\ms{lin}(X) \subseteq \tint\), we can
lift this to all types by taking \(\ms{lin}(A \otimes B) = \ms{lin}(A) \cap
\ms{lin}(B)\) and \(\ms{lin}(\mb{1}) = \ms{lin}(\mb{2}) = \tint\). We say a type
with \(\ms{a} \in \ms{lin}(A)\) is \textit{affine}, while a type with \(\ms{r}
\in \ms{lin}(A)\) is \textit{relevant}; a type which is both affine and relevant
is called \textit{intuitionistic}. Our syntax is composed of three main
syntactic categories:
\begin{itemize}
  \item \textit{Terms} \(a, b, c, d, e\), which are typed via the judgement
  \(\hasty{\Gamma}{p}{a}{A}{q}\), which says that the term \(a\) is of type
  \(A\) with \textit{centrality} \(p\) and \textit{quantity} \(q \subseteq
  \{\ms{a}, \ms{r}\}\). Here, the centrality \(p\) indicates whether it is safe
  to substitute this term for a variable, and \(q\), plus the linearity of
  \(A\), control how many times that variable can occur: if \(\ms{a} \in q \cap
  \ms{lin}(A)\) then the variable is called \textit{affine} or
  \textit{discardable} and it is permissible to leave it unused, whereas if
  \(\ms{r} \in q \cap \ms{lin}(A)\) it is called \textit{relevant} or
  \textit{duplicable} and it is permissible to use multiple times. As for
  instructions, a term which is central, relevant, and affine is called
  \textit{pure}. The rules for this judgement are given in Figure
  \ref{fig:ssa-term-typing}.

  \item \textit{Blocks} \(s, t\), which are typed via the judgement
  \(\haslb{\Gamma}{t}{\ms{L}}\), which says that the block \(t\) is well-formed
  with respect to the context \(\Gamma\) and only branches to
  ("\textit{targets}") labels in \(\ms{L}\). The rules for this judgement are
  given in Figure \ref{fig:ssa-block-typing}.

\item \textit{Control-flow graphs} or \textit{CFG}s \(L\), which are typed via
  the judgement \(\lhaslb{\ms{L}}{L}{\ms{K}}\), which says that \(L\) defines a
  control-flow graph with input labels \(\ms{L}\) and output labels \(\ms{K}\);
  recursive control flow is allowed. The rules for this judgement are given in
  Figure \ref{fig:ssa-block-typing}.
\end{itemize}
The formal grammar for our syntax is given in Figure \ref{fig:ssa-grammar} in
the appendix, while the typing judgements themselves are given in Figure
\ref{fig:ssa-judgements}.

Here \textit{contexts} \(\Gamma\), consist of a list of variables \(x\) given a
type \(A\) annotated with a \textit{quantity} \(q \subseteq \tint\) (if \(q\) is
omitted, we assume \(q = \tint\)). The \textit{linearity} of a variable
\(\ms{lin}(\thyp{x}{A}{q})\) is defined to be \(\ms{lin}(A) \cap q\). We say a
variable \(x: A^q\) is \textit{affine} if \(\ms{a} \in
\ms{lin}(\thyp{x}{A}{q})\), in which case it can be freely ignored. Similarly,
\(x\) is \textit{relevant} if \(\ms{r} \in \ms{lin}(\thyp{x}{A}{q})\), in which
case it can be used multiple times. This is formalized in the structural
judgement \(\csplits{\Gamma}{\Delta}{\Xi}\), which says that \(\Gamma\) can be
split into subcontexts \(\Delta\) and \(\Xi\) (weakening is simply the special
case \(\cwk{\Gamma}{\Delta} \iff \csplits{\Gamma}{\Delta}{\cdot}\)), as per the
typing rules in Figure \ref{fig:ssa-structural}. In particular, the rule
\rle{split-nil} states that the empty context can be split into two empty
contexts, while the rules \rle{split-left} and \rle{split-right} send \(x: A^q\)
to the left or right subcontext respectively; both these rules also allow us to
reduce the quantity \(q\) to a quantity \(r \subseteq q\), "forgetting" some of
our capabilities. The rule \rle{split-dup} allows the duplication of a relevant
variable, with the option to weaken the quantity to \textit{different}
quantities \(r\) and \(s\) on the left and right respectively. Finally, the rule
\rle{split-aff} simply allows us to drop an affine variable.

On the other hand, \textit{label-sets} \(\ms{L}\) are defined to consist of a
list of \textit{labels} \(\lbl{\ell}\) annotated with a context \(\Gamma\)
(consisting of variables which must be live on entry to \(\lbl{\ell}\)) and an
argument type \(A\). The judgement \(\lwk{\ms{L}}{\ms{K}}\) states that
\(\ms{L}\) is a subcontext of \(\ms{K}\), i.e., has more labels with fewer live
variables, and hence that every program which targets \(\ms{L}\) can also be
viewed as targeting \(\ms{K}\). The conditions for this are formalized in Figure
\ref{fig:ssa-structural} as follows: the rule \rle{join-nil} states that the
empty label-set is a subcontext of itself. The rule \rle{join-cons} says that if
\(\ms{L}\) is a subcontext of \(\ms{K}\) and \(\Gamma\) is a weakening of
\(\Delta\), then \(\lwk{\ms{L}, \llhyp{\ell}{\Gamma}{A}}{\ms{K},
\llhyp{\ell}{\Delta}{A}}\), allowing us to "forget" any extra (affine) variables
dropped by \(\cwk{\Gamma}{\Delta}\). Finally, the rule \rle{join-zero} simply
says that if \(\lwk{\ms{L}}{\ms{K}}\), then \(\ms{L}\) is also a subcontext of
any extension of \(\ms{K}\).
% Note that this judgement is "backwards," in that the "larger" context is on
% the \textit{right}; this corresponds to the placement of the label context on
% the right-hand side of the typing judgement for blocks, and is by analogy to
% our semantics for blocks and CFGs in Section \ref{sec:semantics}.

\begin{figure}
  \begin{center}        
    \begingroup
    \renewcommand{\arraystretch}{1.5}
    \setlength{\tabcolsep}{2em}
    \begin{tabular}{rl}
        \multicolumn{1}{c}{Judgment} & \multicolumn{1}{c}{Meaning} \\ \hline
        \(\hasty{\Gamma}{p}{a}{A}{q}\) & \(a\) is a term of type \(A\) in
        context \(\Gamma\) with centrality \(p\) and quantity \(q\) \\
        \(\haslb{\Gamma}{t}{\ms{L}}\) & \(t\) is a block targeting label-set
        \(\ms{L}\) in context \(\Gamma\) \\
        \(\lhaslb{\ms{L}}{L}{\ms{K}}\) & The control-flow graph \(L\) sends
        input labels \(\ms{L}\) to output labels \(\ms{K}\) \\
        \(\csplits{\Gamma}{\Delta}{\Xi}\) & The context \(\Gamma\) splits into
        \(\Delta\) and \(\Xi\) \\
        \(\lwk{\ms{L}}{\ms{K}}\) & The label-set \(\ms{L}\) weakens to the
        label-set \(\ms{K}\) \\
        % \(\tlin{A}{q}\) & The type \(A\) can be used with linearity \(q\) \\
        % \(\ltlin{A}{r}{q}\) & The type \(A\) can be used with linearity \(q
        % \subseteq r\) (i.e. \(\tlin{A}{q} \land q \subseteq r\)) \\
        % \(\tlin{\Gamma}{q}\) & The context \(\Gamma\) has linearity \(q\) \\
        \(\cwk{\Gamma}{\Delta}\) & \(\Gamma\) is a weakening of \(\Delta\) (i.e.
        \(\csplits{\Gamma}{\Delta}{\cdot}\))
    \end{tabular}
    \endgroup
  \end{center}
  \caption{Typing judgements for \isotopessa}
  \Description{Typing judgements for isotope-SSA}
  \label{fig:ssa-judgements}
\end{figure}

\begin{figure}
  \begin{gather*}    
    % \prftree[r]{\rle{base-lin}}{q \subseteq \ms{lin}(X)}{\tlin{X}{q}} \qquad
    % \prftree[r]{\rle{unit-lin}}{}{\tlin{\mathbf{1}}{q}} 
    % \qquad
    % \prftree[r]{\rle{bool-lin}}{}{\tlin{\mathbf{2}}{q}} 
    % \qquad
    % \prftree[r]{\rle{pair-lin}}{\tlin{A}{q}}{\tlin{B}{q}}{\tlin{A \otimes B}{q}} 
    % \\
    % \prftree[r]{\rle{nil-lin}}{\tlin{\cdot}{q}} \qquad
    % \prftree[r]{\rle{cons-lin}}{\ltlin{A}{r}{q}}{\tlin{\Gamma}{q}}
    %   {\tlin{\Gamma, \thyp{x}{A}{r}}{q}} \\
    \prftree[r]{\rle{split-nil}}{\csplits{\cdot}{\cdot}{\cdot}} \qquad
    \prftree[r]{\rle{split-left}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {r \subseteq q}
      {\csplits{\Gamma, \thyp{x}{A}{q}}{\Delta, \thyp{x}{A}{r}}{\Xi}} \qquad
    \prftree[r]{\rle{split-right}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {r \subseteq q}
      {\csplits{\Gamma, \thyp{x}{A}{q}}{\Delta}{\Xi, \thyp{x}{A}{r}}} \\
    \prftree[r]{\rle{split-dup}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\ms{r} \in \ms{lin}(\thyp{x}{A}{q})}
      {r, s \subseteq q}
      {\csplits{\Gamma, \thyp{x}{A}{q}}{\Delta, \thyp{x}{A}{r}}{\Xi, \thyp{x}{A}{s}}}
      \qquad
    \prftree[r]{\rle{split-drop}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\ms{a} \in \ms{lin}(\thyp{x}{A}{q})}
      {\csplits{\Gamma, \thyp{x}{A}{q}}{\Delta}{\Xi}}
      \\
    \prftree[r]{\rle{join-nil}}{\lwk{\cdot}{\cdot}} \qquad
    \prftree[r]{\rle{join-cons}}
      {\lwk{\ms{L}}{\ms{K}}}
      {\cwk{\Gamma}{\Delta}}
      {\lwk{\ms{L}, \llhyp{\ell}{\Gamma}{A}}{\ms{K}, \llhyp{\ell}{\Delta}{A}}} 
      \qquad
    \prftree[r]{\rle{join-zero}}
      {\lwk{\ms{L}}{\ms{K}}}
      {\lwk{\ms{L}}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}} 
  \end{gather*}
  \caption{Structural rules for \isotopessa}
  \Description{Structural rules for isotope-SSA}
  \label{fig:ssa-structural}
\end{figure}

We can now get to the typing judgement for expressions in Figure
\ref{fig:ssa-term-typing}. \rle{var} states that \(\thyp{x}{A}{q}\) is
well-typed in \(\Gamma\) if all other variables are affine and \(q \subseteq
q'\), where \(\thyp{x}{A}{q'} \in \Gamma\); allowing us to \textit{forget}
capabilities, but not \textit{add} them. \rle{app} states that, given an
instruction \(f \in \mc{I}_p^q(A, B)\) and a \textit{central} term
\(\hasty{\Gamma}{1}{a}{A}{q}\), we can type \(\hasty{\Gamma}{p}{f\;a}{B}{q}\).
The \rle{pair} rule says that we can type a pair \((a, b)\) in \(\Gamma\) if
\(\csplits{\Gamma}{\Delta}{\Xi}\) where \(\Delta\) types \(a\) and \(\Xi\) types
\(b\), both as central terms. The quantity of a pair is simply the intersection
of the quantity of its components (expressed by allowing any quantity which
types \textit{both} components). Finally, \rle{let} and \rle{let2}, allow us to
type (destructuring) let-bindings. In both cases, we split the context
\(\Gamma\) into subcontexts \(\Delta, \Xi\), then, \(\Xi\) is used to type a
central term \(a\), which is bound to fresh variables appearing in a potentially
non-central term \(e\) which may also depend on variables in \(\Delta\).

\begin{figure}
  \begin{gather*}    
    \prftree[r]{\rle{var}}
      {\cwk{\Gamma}{\thyp{x}{A}{q}}}
      {\hasty{\Gamma}{p}{x}{A}{q}} \qquad
    \prftree[r]{\rle{app}}
      {f \in \mc{I}_p^q(A, B)}
      {\hasty{\Gamma}{1}{a}{A}{q}}
      {\hasty{\Gamma}{p}{f\;a}{B}{q}} \qquad
    \prftree[r]{\rle{pair}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\hasty{\Delta}{1}{a}{A}{q}}
      {\hasty{\Xi}{1}{b}{B}{q}}
      {\hasty{\Gamma}{p}{(a, b)}{A \otimes B}{q}} \\
    \prftree[r]{\rle{true}}
      {\cwk{\Gamma}{\cdot}}
      {\hasty{\Gamma}{p}{\ctt}{\mb{2}}{q}} \qquad
    \prftree[r]{\rle{false}}
      {\cwk{\Gamma}{\cdot}}
      {\hasty{\Gamma}{p}{\cff}{\mb{2}}{q}} \qquad
    \prftree[r]{\rle{let}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\hasty{\Delta, \thyp{x}{A}{}}{p}{e}{B}{q}}
      {\hasty{\Xi}{1}{a}{A}{q}}
      {\hasty{\Gamma}{p}{\letexpr{x}{a}{e}}{B}{q}} \qquad
    % \prftree[r]{\rle{blk}}
    %   {\haslb{\Gamma}{t}{\llhyp{\ell}{\cdot}{A}}}
    %   {\hasty{\Gamma}{0}{\lbsplice{\ell}{A}{t}}{A}{\varnothing}} 
      \\
    \prftree[r]{\rle{unit}}
      {\cwk{\Gamma}{\cdot}}
      {\hasty{\Gamma}{p}{()}{\mb{1}}{q}} \qquad
    \prftree[r]{\rle{let2}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\hasty{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{p}{e}{C}{q}}
      {\hasty{\Xi}{1}{a}{A \otimes B}{q}}
      {\hasty{\Gamma}{p}{\letexpr{(x, y)}{a}{e}}{C}{q}}
  \end{gather*}
  \caption{Typing rules for \isotopessa expressions}
  \Description{Typing rules for isotope-SSA expressions}
  \label{fig:ssa-term-typing}
\end{figure}

The rules for block typing are given in Figure \ref{fig:ssa-block-typing}.
Blocks do not have a type; instead, we judge which label-set \(\ms{L}\) a block
can branch to given that \(\Gamma\) is live on entry. In particular, \rle{br}
says that if \(\lwk{\lhyp{\ell}{\Delta}{A}}{\ms{L}}\), \(\Gamma\) splits into
\(\Delta\) and \(\Xi\), and \(a\) is a pure term which typechecks in \(\Xi\),
then \(\lbrb{\ell}{a}\) targets \(\ms{L}\) given \(\Gamma\) is live on entry.
\rle{ite} states that, if \(\Gamma\) splits into \(\Delta\) and \(\Xi\), \(e\)
typechecks as a pure, boolean term in \(\Xi\), and that, given all variables in
\(\Delta\) are live on entry, \(s, t\) both target \(\ms{L}\), then the
if-statement \(\ite{e}{s}{t}\) targets \(\ms{L}\) given \(\Gamma\) is live on
entry. \rle{let-blk} and \rle{let2-blk} are quite similar to their term
analogues, except that the bound term \(e\) is allowed to be impure (otherwise,
programs could only make use of pure functions!). Finally, \rle{where} says
that, given a block \(t\) which targets \(\ms{L}\), and a CFG \(L\) with inputs
\(\ms{L}\) and outputs \(\ms{K}\), \(\ewhere{t}{L}\) targets \(\ms{K}\). CFGs
are typed using the typing rules \rle{nil-br}, which says that an empty
control-flow graph simply forwards the input labels directly to the output
labels, and \rle{cons-br}, which says that if \(L\) maps \(\ms{L}\) to either
\(\ms{K}\) or \(\lbl{\ell}[\Gamma](A)\), and \(t\) targets \(\ms{L}\) given that
\(\Gamma, \thyp{x}{A}{}\) is live on entry, then \(L\) maps \(\ms{L}\) to
\(\ms{K}\), having (potentially recursively) defined \(\lbl{\ell}\).

\begin{figure}
  \begin{gather*}    
    \prftree[r]{\rle{br}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\lwk{\llhyp{\ell}{\Delta}{A}}{\ms{L}}}
      {\hasty{\Xi}{1}{a}{A}{\tint}}
      {\haslb{\Gamma}{\lbrb{\ell}{a}}{\ms{L}}} 
    \quad
    \prftree[r]{\rle{let-blk}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\haslb{\Delta, \thyp{x}{A}{}}{t}{\ms{L}}}
      {\hasty{\Xi}{0}{a}{A}{q}}
      {\haslb{\Gamma}{\letstmt{x}{a}{t}}{\ms{L}}} 
    \\
    \prftree[r]{\rle{ite}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\hasty{\Xi}{1}{e}{\mb{2}}{\tint}}
      {\haslb{\Delta}{s}{\ms{L}}}
      {\haslb{\Delta}{t}{\ms{L}}}
      {\haslb{\Gamma}{\ite{e}{s}{t}}{\ms{L}}} 
    \\
    \prftree[r]{\rle{let2-blk}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\haslb{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{t}{\ms{L}}}
      {\hasty{\Xi}{0}{a}{A \otimes B}{q}}
      {\haslb{\Gamma}{\letstmt{(x, y)}{a}{t}}{\ms{L}}} \qquad
    \prftree[r]{\rle{where}}
      {\haslb{\Gamma}{t}{\ms{L}}}
      {\lhaslb{\ms{L}}{L}{\ms{K}}}
      {\haslb{\Gamma}{\ewhere{t}{L}}{\ms{K}}}
    \\
    \prftree[r]{\rle{nil-br}}
      {\lhaslb{\ms{L}}{\cdot}{\ms{L}}} \qquad
    \prftree[r]{\rle{cons-br}}
      {\lhaslb{\ms{L}}{L}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}}
      {\haslb{\Gamma, \thyp{x}{A}{}}{t}{\ms{L}}}
      {\lhaslb{\ms{L}}{L, \lwbranch{\ell}{x: A}{t}}{\ms{K}}}
  \end{gather*}
  \caption{Typing rules for \isotopessa blocks}
  \Description{Typing rules for isotope-SSA blocks}
  \label{fig:ssa-block-typing}
\end{figure}

\subsection{Weakening and Substitution}

We can now state some of the basic metatheoretic properties of \isotopessa. We
begin with weakening, which states that we can add unused \textit{affine}
variables to a context, \textit{reduce} the quantity of an term's type, add
arbitrary unused labels, and remove \textit{affine} variables from existing
labels and still keep our typing judgements valid, i.e.
\begin{lemma}[Weakening] 
  Given \(\cwk{\Gamma}{\Delta}\) and \(q \subseteq r\),
  \(\hasty{\Delta}{p}{a}{A}{r} \implies \hasty{\Gamma}{p}{a}{A}{q}\) and
  \(\haslb{\Delta}{t}{\ms{L}} \implies \haslb{\Gamma}{t}{\ms{L}}\). Similarly,
  if \(\lwk{\ms{L}}{\ms{K}}\) then \(\haslb{\Gamma}{t}{\ms{L}} \implies
  \haslb{\Gamma}{t}{\ms{K}}\).
\end{lemma}

\begin{figure}
  \begin{center}        
    \begingroup
    \renewcommand{\arraystretch}{1.5}
    \setlength{\tabcolsep}{2em}
    \begin{tabular}{rl}
        \multicolumn{1}{c}{Judgment} & \multicolumn{1}{c}{Meaning} \\ \hline
        \(\issubst{\gamma}{\Theta}{\Gamma}\) &
        \(\gamma: \ms{Var} \to \ms{Expr}\) is a substitution taking \(\Theta\) to \(\Gamma\) \\
        \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\) &
        \(\mc{L}: \ms{Label} \to \ms{Expr} \to \ms{Block}\) is a label substitution taking \(\ms{L}\) to \(\ms{K}\)
    \end{tabular}
    \endgroup
  \end{center}
  \caption{Metatheoretic typing judgements for \isotopessa}
  \Description{Metatheoretic typing judgements for isotope-SSA}
  \label{fig:ssa-meta-judgements}
\end{figure}

\begin{figure}
  \begin{gather*}
    % \boxed{\gamma: \ms{Var} \to \ms{Expr}}
    % \\
    \prftree[r]{\rle{subst-nil}}
      {\issubst{\gamma}{\cdot}{\cdot}}
      \qquad
    \prftree[r]{\rle{subst-cons}}
      {\issubst{\gamma}{\Theta_\Gamma}{\Gamma}}
      {\hasty{\Theta_x}{1}{\gamma(x)}{A}{q}}
      {\tlin{\Theta_x}{q}}
      {\csplits{\Theta}{\Theta_\Gamma}{\Theta_x}}
      {\issubst{\gamma}{\Theta}{\Gamma, \thyp{x}{A}{q}}}
    \\
    % \boxed{\mc{L}: \ms{Label} \to \ms{Expr} \to \ms{Block} \text{ s.t. for fresh } x, [a/x]\mc{L}(\lbl{\ell}, x) = \mc{L}(\lbl{\ell}, a)}
    % \\
    \prftree[r]{\rle{lb-subst-nil}}
    {\lbsubst{\mc{L}}{\cdot}{\ms{K}}}
      \qquad
    \prftree[r]{\rle{lb-subst-cons}}
      {\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
      {\haslb{\Gamma, \thyp{x}{A}{}}{\mc{L}(\lbl{\ell}, x)}{\ms{K}}}
      {\lbsubst{\mc{L}}{\ms{L}, \llhyp{\ell}{\Gamma}{A}}{\ms{K}}}
  \end{gather*}
  \caption{ Typing rules for \isotopessa substitutions. The linearity of a
    context is defined recursively to be the intersection of the linearities of
    its variables: \(\ms{lin}(\cdot) = \tint\), \(\ms{lin}(\Gamma,
    \thyp{x}{A}{q}) = \ms{lin}(\Gamma) \cap \ms{lin}(\thyp{x}{A}{q})\). We
    assume that a substitution \(\ms{L}\) satisfies, for fresh \(x\),
    \([a/x]\mc{L}(\lbl{\ell}, x) = \mc{L}(\lbl{\ell}, a)\).
    % We generally assume that for all but finitely many variables \(x\) and
    % labels \(\lbl{\ell}\), \(\gamma(x) = x\) and \(\mc{L}(\lbl{\ell}, x) =
    % \lbrb{\ell}{x}\). 
  }
  \Description{Typing rules for isotope-SSA substitutions}
  \label{fig:ssa-subst-typing}
\end{figure}

To state our primary syntactic metatheorems, \textit{substitution} and
\textit{label-substitution}, we will need to introduce some new metatheoretic
judgements, which we do in Figure \ref{fig:ssa-meta-judgements}. In particular,
the judgement \(\issubst{\gamma}{\Theta}{\Gamma}\) says that \(\gamma\), which
is taken to be a map from variable names to expressions, takes each variable
\(\thyp{x}{A}{q}\) in \(\Gamma\) to a \textit{central} term \(\gamma(x)\). Due
to linearity, the resources \(\Theta\) must be split into subcontexts
\(\Theta_x\) which allow us to interpret each \(\gamma(x)\) as a term of type
\(A\) with quantity \(q\). This is described formally via the typing rules
\rle{subst-nil}, which says that any function \(\gamma\) takes the empty context
to itself, and \rle{subst-cons}, which says that if \(\gamma\) takes
\(\Theta_\Gamma\) to \(\Gamma\), \(\hasty{\Theta_x}{1}{\gamma(x)}{A}{q}\), \(q
\subseteq \ms{lin}(\Theta_x)\) (to avoid necessating the duplication/dropping of
substructural variables), and \(\Theta\) splits into \(\Theta_\Gamma\) and
\(\Theta_x\), then \(\issubst{\gamma}{\Theta}{\Gamma, \thyp{x}{A}{q}}\).

Given such a substitution \(\gamma\), there are two ways to use it. The first is
standard capture-avoiding substitution, which we denote \([\gamma]e\) for
expressions, \([\gamma]t\) for blocks, and \([\gamma]L\) for CFGs, as is
standard. The second is to use the substitution to generate a set of
let-bindings for a context \(\Gamma\), written
\(\exprletsubst{\substctx{\gamma}{\Gamma}}{e}\) for expressions and
\(\stmtletsubst{\substctx{\gamma}{\Gamma}}{t}\) for blocks, as defined in Figure
\ref{fig:ssa-subst-ops}.
%
% Note that this operation cannot be easily defined for CFGs. 
%
% In Section \ref{sec:semantics}, we will show that this is semantically
% equivalent to regular capture-avoiding substitution, due to the fact that all
% terms in a substitution are central \textit{and} that only affine computations
% can be used in a term substituting for an affine variable, and only relevant
% computationbs can be used in a term substituting for a relevant variable.
%
With these definitions, we can now state substitution for terms in a
straightforward way: if \(\issubst{\gamma}{\Theta}{\Gamma}\) and
\(\hasty{\Gamma}{p}{e}{A}{q}\), then \(\hasty{\Theta}{p}{[\gamma]e}{A}{q}\) and
\(\hasty{\Theta}{p}{\exprletsubst{\substctx{\gamma}{\Gamma}}{e}}{A}{q}\).

\begin{figure}
  \begin{gather*}
    % \restrictsubst{\gamma}{\Gamma}(x) 
    %   =
    %   (\ms{if}\; x \in \Gamma
    %   \;\ms{then}\;\gamma(x)
    %   \;\ms{else}\;x)
    % \\
    \stmtletsubst{\substctx{\gamma}{\cdot}}{t} = t
    \qquad
    \stmtletsubst{\substctx{\gamma}{\Gamma, x: A^q}}{t} 
      = \letstmt{x}{\gamma(x)}{\stmtletsubst{\substctx{\gamma}{\Gamma}}{t}}
    \qquad
    \exprletsubst{\cdot}{e} = e
    \qquad
    \exprletsubst{\substctx{\gamma}{\Gamma, x: A^q}}{e} 
      = \letstmt{x}{\gamma(x)}{\substctx{\gamma}{\Gamma}}
    \\
    \substlbs{\gamma}{\cdot}(\lbl{\ell}, x) = \lbrb{\ell}{x}
    \qquad
    \substlbs{\gamma}{\ms{L}, \llhyp{\ell}{\Gamma}{A}}(\lbl{k}, x)
      = 
      (\ms{if}\; \lbl{k} = \lbl{\ell} 
      \;\ms{then}\;\stmtletsubst{\substctx{\gamma}{\Gamma}}{\lbrb{\ell}{x}}
      \;\ms{else}\;\substlbs{\gamma}{\ms{L}}(\lbl{\ell}, x))
    \\
    [\mc{L}](\lbrb{\ell}{a}) = \mc{L}(\lbl{\ell}, a)
    \qquad
    [\mc{L}]\ite{e}{s}{t} = \ite{e}{[\mc{L}]s}{[\mc{L}]t}
    \\
    [\mc{L}](\letstmt{x}{a}{t}) = (\letstmt{x}{a}{[\mc{L}]t})
    \qquad
    [\mc{L}](\letstmt{(x, y)}{a}{t}) = (\letstmt{(x, y)}{a}{[\mc{L}]t})
    \\
    [\mc{L}]\ewhere{t}{L} = \ewhere{[\mc{L}]t}{[\mc{L}]L}
    \\
    [\mc{L}]\cdot = \cdot
    \qquad
    [\mc{L}](\lwbranch{\ell}{x: A}{t})
    = (\lwbranch{\ell}{x: A}{[\mc{L}]t})
  \end{gather*}
  \caption{Operations on \isotopessa substitutions.}
  \Description{Operations on isotope-SSA substitutions}
  \label{fig:ssa-subst-ops}
\end{figure}

Unfortunately, substitution for blocks is a little more complicated: the naive
choice of attempting to show \(\haslb{\Gamma}{t}{\ms{L}} \implies
\haslb{\Theta}{[\gamma]t}{\ms{L}}\) does not work, since \(\ms{L}\) may require
variables from \(\Gamma\) to be live which simply do not exist in \(\Theta\). To
solve this problem, we need to bind the terms in \(\gamma\) to variables with
the correct names for the live variable sets of desired targets. To do this, we
introduce the concept of a \textit{label-substitution} \(\mc{L}\), which we
define as a function that takes labels \(\lbl{\ell}\) and variables \(x\) to
blocks \(t\) such that, for any fresh variable \(x\), \([a/x]\mc{L}(\lbl{\ell},
x) = \mc{L}(\lbl{\ell}, a)\). We introduce the judgement
\(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\) to mean that for each
\(\lbl{\ell}[\Gamma](x: A) \in \ms{L}\), we have that \(\mc{L}(\lbl{\ell}, x)\)
targets \(\ms{K}\). Formally, \rle{lb-subst-nil} states that this is vacuously
true for \(\ms{L} = \cdot\), while \rle{lb-subst-cons} says that if
\(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\) and \(\haslb{\Gamma,
\thyp{x}{A}{}}{\mc{L}(\lbl{\ell}, x)}{\ms{K}}\) then \(\lbsubst{\mc{L}}{(\ms{L},
\lhyp{\lbl{\ell}}{\Gamma}{A})}{\ms{K}}\). Given a context \(\gamma\) and a
label-set \(\ms{L}\), we can define an associated label-substitution
\(\substlbs{\gamma}{\ms{L}}\) using the definition in Figure
\ref{fig:ssa-subst-ops}: using this, we can state the full substitution lemma as
follows:
\begin{lemma}[Substitution] 
  Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
  \begin{itemize}
    \item If \(\hasty{\Gamma}{p}{a}{A}{q}\), then
    \(\hasty{\Theta}{p}{[\gamma]a}{A}{q}\) and
    \(\hasty{\Theta}{p}{\exprletsubst{\substctx{\gamma}{\Gamma}}{a}}{A}{q}\)
    \item If \(\haslb{\Gamma}{t}{\ms{L}}\), then there exists \(\ms{K}\) such
    that \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) and
    \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\) and
    \(\haslb{\Theta}{\stmtletsubst{\substctx{\gamma}{\Gamma}}{t}}{\ms{K}}\)
    \item If \(\lhaslb{\ms{W}}{L}{\ms{L}}\), then there exists \(\ms{K},
    \ms{W}'\) such that \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\),
    \(\lbsubst{\substlbs{\gamma}{\ms{W}'}}{\ms{W}}{\ms{W}'}\), and
    \(\lhaslb{\ms{W}'}{[\gamma]L}{\ms{K}}\)
  \end{itemize}
  Similarly, given \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\), we have that for all
  \(\haslb{\Gamma}{t}{\ms{L}}\), \(\haslb{\Gamma}{[\mc{L}]t}{\ms{K}}\)
\end{lemma}

\section{SSA Semantics}

\label{sec:semantics}

In this section, we will give a categorical semantics for \isotopessa. Starting
with Freyd categories, we will generalize to \textit{substructural effectful
categories} to support substructurality. We will then require our substructural
effectful category to have \textit{distributive coproducts} to implement
branching. Finally, we will require an \textit{Elgot structure} to implement
potentially nonterminating control flow.

\subsection{Substructural Categories}

We begin with some standard definitions from the literature:
\begin{definition}[Binoidal Category] 
  A \textbf{binoidal category} \(\mc{C}\) is a category equipped with a binary
  operation on objects \(\otimes: \mc{C} \times \mc{C} \to \mc{C}\) which is
  independently functorial in each argument (i.e., for each object \(A\), \(A
  \otimes -\) and \(- \otimes A\) are functors). 
  Given \(f: \mc{C}(A, B)\), \(g: \mc{C}(A', B')\) in a binoidal category, we
  write \(f \ltimes g = f \otimes A';B \otimes g\) and \(f \rtimes g = A \otimes
  g;f \otimes B'\); we say \(f\) is \textbf{central} if, for all \(g\), \(f
  \ltimes g = f \rtimes g\) and \(g \ltimes f = g \rtimes f\), in which case we
  simply write \(f \otimes g\), \(g \otimes f\) respectively.
\end{definition}
\begin{definition}[Premonoidal Category] 
  A \textbf{premonoidal category} is a binoidal category with an identity object
  \(I: \mc{C}\) such that there exist central natural isomorphisms \(\lambda_A:
  I \otimes A \simeq A\) (the \textbf{left unitor}), \(\rho_A: A \otimes I
  \simeq A\) (the \textbf{right unitor}), and \(\alpha_{A, B, C}: (A \otimes B)
  \otimes C \simeq A \otimes (B \otimes C)\) (the \textbf{associator})
  satisfying the \textit{pentagon} and \textit{triangle laws} (see e.g.
  \citet{maclane:71}). We say a premonoidal category is \textbf{symmetric} if it
  is equipped with a central natural isomorphism \(\sigma_{A, B}: A \otimes B
  \simeq B \otimes A\) (the \textbf{symmetry}) satisfying the \textit{hexagon
  law} (see \cite{maclane:71}). We say a premonoidal category is
  \textbf{monoidal} if every morphism is central.
\end{definition}
One issue with the notion of a premonoidal category is that it is unclear what
exactly a premonoidal functor should be: if we want them to compose, we need to
preserve central morphisms, but trying to preserve \textit{all} central
morphisms rules out many important examples. We therefore would to be able to
distinguish morphisms that are truly "pure" from those which are simply
"accidentally" central. To do that, we introduce the notion of an
\textit{effectful category} as follows:
\begin{definition}[Effectful Category] 
  An \textbf{effectful category} \(\mc{V} \to \mc{C}\) is an identity on objects
  functor \(\upg{\cdot}{}: \mc{V} \to \mc{C}\) from a monoidal category
  \(\mc{V}\) to a premonoidal category \(\mc{C}\) which preserves all
  premonoidal structure (i.e. sends associators to associators, unitors to
  unitors, and tensor products to tensor products). It is \textbf{symmetric} if
  both \(\mc{V}\) and \(\mc{C}\) are, and \(\upg{\cdot}{}\) preserves the
  symmetry. A \textbf{Freyd category} is an effectful category where \(\mc{V}\)
  is Cartesian.
\end{definition}
As is, this definition presents us with two extreme options: we can either
assume \(\mc{V}\) is simply monoidal, and hence that "pure" morphisms are merely
central, or that \(\mc{V}\) is Cartesian, and hence that all pure morphisms are
truly \textit{pure}. We would like, however, to make a finer distinction between
morphisms which are merely \textit{central} and those which are also
\textit{relevant} and/or \textit{affine}, as defined in the introduction. Doing
so also allows us to add the capability of define what it means for an
\textit{object} to be affine or relevant as well. With this aim in mind, we
introduce the notion of a \textit{symmetric premonoidal subcategory}, as
follows:
\begin{definition}[(Symmetric) Premonoidal Subcategory]
  Given a (symmetric) premonoidal category \(\mc{C}\), we say a wide subcategory
  \(\mc{S} \subseteq \mc{C}\) is \textbf{(symmetric) premonoidal} if it is
  closed under whiskering and contains all associators, unitors, and symmetries.
  We say a subcategory is \textbf{monoidal} if, for all \(f: \mc{S}(A, B)\),
  \(g: \mc{S}(A', B')\), \(f \ltimes g = f \rtimes g\); we say it is
  \textbf{central} if this generalizes to the case where only one of \(f, g\) is
  in \(\mc{S}\) (i.e., if all morphisms in \(\mc{S}\) are central).
\end{definition} 
Note that the symmetric premonoidal subcategories, monoidal subcategories, and
central subcategories of a category \(\mc{C}\) are all closed under
intersections of hom-sets, forming a complete meet-semilattice.
%
% and therefore, in particular, form a complete meet-semilattice with shared
% bottom element \(\ms{Skeleton}(\mc{C})\), where \(\ms{Skeleton}(C)\) is the
% smallest possible symmetric premonodial subcategory. The largest symmetric
% premonoidal subcategory is of course \(\mc{C}\), while the largest central
% subcategory is the center of \(\mc{C}\), \(Z(\mb{C})\).
%
% ; the existence of a top-element means that in these cases we may induce the
% structure of a complete lattice in the usual manner.
%
We can now define a \textit{substructural premonoidal category}, which is simply
a premonoidal category equipped with premonoidal subcategories of relevant and
affine morphisms:
\begin{definition}[Substructural (Pre)monoidal Category]
  A \textbf{substructural (pre)monoidal category} is a symmetric premonoidal
  category \(\mc{C}\) equipped with
  \begin{itemize}
    \item Wide symmetric premonoidal subcategories \(\mc{C}^{\ms{a}},
    \mc{C}^{\ms{r}}\). We define \(\mc{C}^q = \bigcap_{\ell \in q}\mc{C}^\ell\),
    where \(\mc{C}^\varnothing = \mc{C}\)
    \item A set of \textbf{affine} objects \(\ms{Aff}(\mc{C}) = \mc{A} \subseteq
    |\mc{C}|\) closed under tensors, such that \(\mb{1} \in \mc{A}\) and
    \begin{itemize}
      \item Every \(A\) in \(\mc{A}\) is equipped with a central morphism
      \(\ms{drop}(A): \mc{C}^{\{\ms{a}, \ms{r}\}}(A, \mb{1})\) satisfying
      \item \(\forall A, B \in \mc{A}. \ms{drop}(A \otimes B) = \ms{drop}(A) \otimes \ms{drop}(B);\lambda_I\), \(\forall f: \mc{C}^{\ms{a}}(A, B). f;\ms{drop}(B) = \ms{drop}(A)\)
    \end{itemize}
    \item A set of \textbf{relevant} objects \(\ms{Rel}(\mc{C}) = \mc{R}
    \subseteq |\mc{C}|\) closed under tensors, such that \(\mb{1} \in \mc{R}\)
    and
    \begin{itemize}
      \item Every \(A\) in \(\mc{R}\) is equipped with a central morphism
      \(\ms{split}(A): \mc{C}^{\{\ms{a}, \ms{r}\}}(A, A \otimes A)\) satisfying
      \item \(\forall A, B \in \mc{R}, \ms{split}(A \otimes B) = \ms{split}(A) \otimes \ms{split}(B);\alpha;A \otimes \sigma_{A, B} \otimes B;\alpha\)
      \item Associativity: \(\forall A \in \mc{R}, \ms{split}(A);\ms{split}(A) \otimes A;\alpha = \ms{split}(A);A \otimes \ms{split}(A)\)
      \item Commutativity: \(\forall A \in \mc{R}, \ms{split}(A);\sigma = \ms{split}(A)\)
      \item Comonoid: \(\forall A \in \mc{A} \cap \mc{R}, \ms{split}(A);A \otimes \ms{drop}(A) = \rho_A^{-1} \quad \ms{split}(A);\ms{drop}(A) \otimes A = \lambda_A^{-1}\)
      \item \(\forall A, B \in \mc{R}, \forall f: \mc{C}^{\ms{r}}(A, B), f;\ms{split}(B) = \ms{split}(A);f \ltimes f = \ms{split}(A);f\rtimes f\)
    \end{itemize}
  \end{itemize}
  We call the morphisms in \(\mc{C}^{\ms{a}}\) \textbf{affine}, the morphisms in
  \(\mc{C}^{\ms{r}}\) \textbf{relevant}, and the morphisms in
  \(\mc{C}^{\{\ms{a}, \ms{r}\}}\) \textbf{intuitionistic}. 
  % We call the morphisms in \(\mc{C}^\varnothing\) but not necessarily in any
  % of the other \(\mc{C}^q\) \textbf{linear}. 
  We call a substructural \textit{monoidal} category a \textbf{substructural
  Cartesian category} if \(\mc{A} = \mc{R} = |\mc{C}|\), in which case
  \(\mc{C}^{\{\ms{a}, \ms{r}\}}\) is Cartesian.
\end{definition}

\paragraph{Example}

A simple example of a substructural Cartesian category is the category of
relations \(\ms{Rel}\), viewed as the Kleisli category of the power-set monad
with morphisms of the form \(\ms{Rel}(A, B) = A \to \mc{P}(B)\). In particular,
we can define:
\begin{equation}
  \begin{aligned}
    \ms{Rel}^{\ms{a}}(A, B) 
    &= \{R \in \ms{Rel}(A, B) \mid \forall a, |R(a)| \geq 1\} 
    &&= A \to \mc{P}^+(B) \\
    \ms{Rel}^{\ms{r}}(A, B) 
    &= \{R \in \ms{Rel}(A, B) \mid \forall a, |R(a)| \leq 1\} 
    &&= A \to \ms{Option}(B) \\
    \implies \ms{Rel}^{\{\ms{a}, \ms{r}\}}(A, B) 
    &= \{R \in \ms{Rel}(A, B) \mid \forall a, |R(a)| = 1\} 
    &&= A \to B
  \end{aligned}
\end{equation}
Here, \(\ms{split}(A)\) is just the copy operator \(a \mapsto (a, a)\) and
\(\ms{drop}(A)\) the deletion operator \(a \mapsto ()\); it is quite trivial to
prove that these lie in \(\ms{Rel}^{\{\ms{a}, \ms{r}\}}\), being functions, and
that they satisfy the axioms of a commutative comonoid, as desired.

Note that we cannot simply have \(\ms{Rel}^{\ms{a}} = \ms{Rel}\), since, for any
morphism \(R: \ms{Rel}(A, B)\) such that \(\exists a \in A, R(a) =
\varnothing\), we have \((R;\ms{drop}(B))(a) = \varnothing \neq \ms{drop}(A)(a)
= \{()\} \). We will see that this corresponds to the fact that "\(\ms{let}\;x =
\ms{loop}(); e\)" is not semantically equal to "\(e\)" even where \(x \notin
\ms{fv}(e)\). Similarly, we cannot simply have \(\ms{Rel}^{\ms{r}} = \ms{Rel}\),
since, for any \(R: \ms{Rel}(A, B)\) such that \(\exists a \in A, |R(a)| > 1\),
\(
  (R;\ms{split}(B))(a) = \{(b, b) \mid b \in R(a)\} \neq (\ms{split}(A);R)(a) = R(a) \times R(a) 
\). This corresponds
to the fact that the program "\(\ms{let}\;x = \ms{nondet}(); (x, x)\)" is not
semantically equal to "\((\ms{nondet}(), \ms{nondet}())\)".

% To get a substructural premonoidal category, we can apply the state monad
% transformer to the power set monad to obtain the Kleisli category with morphisms
% of the form \(\ms{Set}_{\ms{St}\;\mc{P}\;S}(A, B) = A \to S \to \mc{P}(B \times
% S)\). It is clear that all objects have split and drop morphisms inherited from
% \(\ms{Set}\), and therefore are affine and relevant.

Just like for regular premonoidal categories, we would like to distinguish
between those morphisms that are "truly" central versus those which are just
central "by accident". We can do this the same way, by simply extending the
notion of effectful category appropriately as follows:
\begin{definition}[Susbtructural Effectful Category]
  An effectful category \(\mc{C}_1 \to \mc{C}_0\) is said to be
  \textbf{substructural} if \(\mc{C}_0, \mc{C}_1\) are substructural and
  \(\upg{\cdot}{}: \mc{C}_1 \to \mc{C}_0\) preserves all drops and splits and
  sends \(\mc{C}_1^q\) to \(\mc{C}_0^q\). In particular, restricting
  \(\upg{\cdot}{}\) appropriately makes all \(\mc{C}_1^q \to \mc{C}_0^q\) into
  effectful categories.  We call a morphism in \(\mc{C}_1\) \textbf{central},
  and a central intuitionistic morphism \textbf{pure}.
  % Morphisms in \(\mc{C}_p^{\ms{r}}\) are called \textbf{copyable}, morphisms
  % in \(\mc{C}_p^{\ms{a}}\) are called \textbf{discardable}, while morphisms in
  % \(\mc{C}_1^{\ms{a}}, \mc{C}_1^{\ms{r}}, \mc{C}_1^{\{\ms{a}, \ms{r}\}}\)
  % specifically are called \textbf{affine}, \textbf{relevant}, and
  % \textbf{intuitionistic} or \textbf{pure} respectively.
\end{definition}

\paragraph{Example}

Applying the FP-style state transformer to the power-set monad yields the
Kliesli category with morphisms of the form \(\mc{C}_0(A, B) = A \to S \to
\mc{P}(A \times S)\). Taking \(\mc{C}_1\) to be the original category
\(\ms{Rel}\), where \(\upg{R}{} = \lambda a, s. R(a) \times \{s\}\) is the
standard lifting operator, and taking \(\mc{C}_1^q = \ms{Rel}^q\), we can verify
that this gives us the axioms of a substructural effectful category. Note that
here the \textit{central} morphisms are precisely those which do not read from
or write to the state, and therefore which commute with all other morphisms. In
general,
\begin{proposition}
  If \(\mc{C}\) is a substructural monoidal category and \(\mb{T}\) is a strong
  monad on \(\mc{C}\) then we can make the Kleisli category \(\mc{C}_{\mb{T}}\)
  into a substructural effectful category \(\upg{\cdot}{}: \mc{C} \to
  \mc{C}_{\mb{T}}\) by taking \(\ms{Aff}(\mc{C}_{\mb{T}}) = \ms{Aff}(\mc{C})\),
  \(\ms{Rel}(\mc{C}_{\mb{T}}) = \ms{Rel}(\mc{C})\), and \(\mc{C}_{\mb{T}}^q(A,
  B) = \upg{\mc{C}^q(A, B)}{}\), where \(\upg{f}{} = f;\eta\), and defining
  drops \(\upg{\ms{drop}(A)}{}\) and splits \(\upg{\ms{split}(A)}{}\).
\end{proposition}
Moreover, \textit{every} premonoidal category \(\mc{C}\) can be viewed a
substructural premonoidal category, taking \(\mc{C}^{\ms{a}}\) and
\(\mc{C}^{\ms{r}}\) to be the subcategory consisting only of the associators,
unitors, and symmetry, and only the identity object to be affine and relevant.
Similarly, every \textit{Cartesian} category \(\mc{C}\) can be viewed as a
substructural Cartesian category with all morphisms and objects both relevant
and affine; in fact, this is the case \textit{if and only if} a category is
Cartesian.

\subsection{Distributivity}

To support branching, we need a bit more structure:
\begin{definition}[Distributive (pre)monoidal category] 
  A \textbf{distributive (pre)monoidal category} is a (pre)monoidal category
  with all finite coproducts equipped with a natural family of central
  isomorphisms (the \textbf{distributor}) \(\delta_{A, B, C}: (A + B) \otimes C
  \simeq A \otimes C + B \otimes C\). This isomorphism is subject to rather
  complex coherence conditions, worked out in full for the more general case of
  \textit{rig categories} by \citet{laplaza-distributivity}, but for our
  purposes we will only make use of the following:
  \begin{equation*}
    \delta_{A, B, C}^{-1} = [\iota_0 \otimes C, \iota_1 \otimes C] 
    \qquad
    \sigma^+_{A, B} \otimes C;\delta_{B, A, C} = \delta_{A, B, C};\sigma^+_{A \otimes C, B \otimes C}
  \end{equation*}
  Note in particular that this implies that \(\iota_i \otimes C;\delta_{A, B, C}
  = \iota_i\). A substructural (pre)monodial category is \textbf{distributive}
  if it is distributive as a (pre)monoidal category, and its distributor is
  \textit{pure}. A substructural effectful category \(\mc{C}_1 \to \mc{C}_0\) is
  \textbf{distributive} if \(\mc{C}_1, \mc{C}_0\) are distributive (as a
  substructural (pre)monoidal category) and \(\upg{\cdot}{}\) strictly preserves
  all coproducts.
\end{definition}
Many (pre)monoidal categories are distributive, including of course
\(\ms{Set}\), but also categories which can be derived from monads on more
primitive categories such as \(\ms{PFun}\) and \(\ms{Rel}\), as can be deduced
from the following proposition:
\begin{proposition}
  If \(\mc{C}\) is a distributive substructural monoidal category and \(\mb{T}\)
  is a strong monad on \(\mc{C}\), then the Kleisli category \(\mc{C}_{\mb{T}}\)
  is a distributive substructural effectful category (with affine and relevant
  objects and morphisms lifted from \(\mc{C}\)).
\end{proposition}
% \begin{proof}
%   Let \(\delta_{A, B, C}: \mc{C}((A + B) \otimes C, A \otimes C + B \otimes C)\) denote \(\mc{C}\)'s distributor. We will write composition in \(\mc{C}\) as "\(;\)" and composition in the Kleisli category \(\mb{T}\) as "\(\gg\)".
%   We will show that
%   \(\upg{\delta_{A, B, C}}{}: \mc{C}_{\mb{T}}((A + B) \otimes C, A \otimes C + B \otimes C)\) satisfies the desired properties of a distributor. In particular, it is:
%   \begin{itemize}
%     \item Central: this follows immediately from being the image of a pure morphism
%     \item Affine: \(\upg{\delta}{} \gg \upg{\ms{drop}}{} = \upg{(\delta;\ms{drop})}{} = \upg{\ms{drop}}{}\)
%     \item Relevant: \(\upg{\delta}{} \gg \upg{\ms{split}}{} = \upg{(\delta;\ms{split})}{} = \upg{(\ms{split};\delta \otimes \delta)}{} = \upg{\ms{split}}{} \gg \upg{\delta}{} \otimes \upg{\delta}{}\)
%     \item Left inverse: \(\upg{\delta}{} \gg \upg{[\iota_0 \otimes C, \iota_1
%     \otimes C]}{} = \upg{\delta;[\iota_0 \otimes C, \iota_1 \otimes C]}{} =
%     \ms{id}\)
%     \item Right inverse: \(\upg{[\iota_0 \otimes C, \iota_1 \otimes C]}{} \gg \upg{\delta}{} = \upg{[\iota_0 \otimes C, \iota_1 \otimes C];\delta}{} = \ms{id}\)
%     \item Symmetric: \(\upg{\sigma^+}{} \otimes C \gg \upg{\delta}{} =
%     \upg{\sigma^+ \otimes C;\delta}{} = \upg{\delta;\sigma^+}{} = \upg{\delta}{}
%     \gg \upg{\sigma^+}{}\)
%   \end{itemize}
% \end{proof}

\subsection{Elgot Categories}

The structures we've developed so far are enough to support non-looping control
flow, but to allow general, potentially nonterminating control flow, we need a
way to reason about iteration. To do so, we introduce the notion of an
\textit{Elgot category} as follows:
\begin{definition}[Elgot category] 
  An \textbf{Elgot category} is a category \(\mc{C}\) with coproducts equipped
  with an operator \(\cdot^\dagger: \mc{C}(A, B + A) \to \mc{C}(A, B)\)
  satisfying the following properties:
  \begin{itemize}
    \item \textbf{Fixpoint:} \(f;[\ms{id}, f^\dagger] = f^\dagger\)
    \item \textbf{Naturality:} \((f;g + \ms{id})^\dagger = f^\dagger;g\)
    \item \textbf{Codiagonal:} \(\forall f \in \mc{C}(A, (B + A) + A), (f^\dagger)^\dagger = (f;[\ms{id}, \iota_1])^\dagger\)
    \item \textbf{Uniformity:}
    \(
      h;f = g;\ms{id} + h \implies h;f^\dagger = g^\dagger
    \)
  \end{itemize}
  % A monad \(\mb{T}\) on \(\mc{C}\) such that its Kleisli cateogry
  % \(\mc{C}_{\mb{T}}\) is Elgot is called an \textbf{Elgot monad}.
\end{definition}
% Some important properties which hold for all Elgot categories include:
% \begin{itemize}
%   \item \textbf{Dinaturality:} \((g;[\iota_0, h])^\dagger = g;[\ms{id},
%   (h;[\iota_0, g])^\dagger]\)
%   \item \textbf{Squaring:} \((f;[\iota_0, f])^\dagger = f^\dagger\)
% \end{itemize}
% Proofs can be found in Lemma 31 of \cite{goncharov-squaring}.
A monad \(\mb{T}\) whose Kliesli category \(\mb{C}_{\mb{T}}\) is Elgot is called
an \textbf{Elgot monad}. Examples of Elgot monads include \(\ms{Option}\) (where
nontermination is \(\ms{None}\)), the powerset monad (where nontermination is
the empty set), which can both be viewed as special cases of the trace monads we
will explore in Section \ref{ssec:trace-monads}. It turns out that various monad
transformers preserve Elgot structure, allowing us to build up Elgot monads in a
composable fashion via a monad stack. In particular,
\begin{proposition}
  If \(\mb{T}\) is an Elgot monad on \(\ms{Set}\), then \(\ms{StateT}\;S\;\mb{T}
  = \lambda A. S \to \ms{T}(A \times S)\), \(\ms{ReaderT}\;\mb{T}\;S = \lambda
  A. S \to \ms{T}(A)\), and \(\ms{WriterT}\;\mb{T}\;S = \lambda A. \ms{T}(A
  \times S)\) are all Elgot, with iteration operators lifted from \(\mb{T}\) in
  the obvious manner.
\end{proposition}

\subsection{Denotational Semantics}

We can now give a semantics for an \isotopessa dialect with types
\(\mc{T}\) and instructions \(\mc{I}\). We'll need:
\begin{itemize}
  \item A \textit{substructural effectful category} \(\mc{C} = \mc{C}_1 \to
  \mc{C}_0\) such that \(\mc{C}_0\) is \textit{distributive} and \textit{Elgot}
  \item A function \(\ms{base}(X)\) from base types \(X \in \mc{T}\) to objects
  in \(|\mc{C}|\), such that \(\ms{a} \in \ms{lin}(X) \implies \ms{base}(X) \in
  \ms{Aff}(\mc{C})\) and \(\ms{r} \in \ms{lin}(X) \implies \ms{base}(X) \in
  \ms{Rel}(\mc{C})\)
  \item A function \(\ms{inst}(f)\) from instructions \(f \in \mc{I}_p^q(A, B)\)
  to morphisms in \(\mc{C}_p^q(\dnt{A}, \dnt{B})\)
\end{itemize}
We begin by giving a semantics for types and contexts in terms of objects in
Figure \ref{fig:ssa-type-semantics}: we map tensor products to tensor products,
the unit type to the monoidal unit, and booleans to \(\mb{1} + \mb{1}\).
Contexts are simply interpreted as the tensor products of the variable types
which appear in them; in particular, quantity annotations have \textit{no}
effect on a context's semantics. We proceed to give semantics for structural
judgements in Figure \ref{fig:ssa-structural-semantics}. In particular, we
interpret a context split \(\csplits{\Gamma}{\Delta}{\Xi}\) as a pure morphism
\(\mc{C}^\tint_1(\dnt{\Gamma}, \dnt{\Delta} \otimes \dnt{\Xi})\) which
partitions the variables as appropriate, splitting and dropping as necessary; a
weakening \(\cwk{\Gamma}{\Delta}\) is interpreted as the associated split
\(\dnt{\csplits{\Gamma}{\Delta}{\cdot}}\) postcomposed with a unitor. Finally, a
label-set weakening \(\lwk{\ms{L}}{\ms{K}}\) is interpreted as a pure morphism
\(\mc{C}^\tint_1(\dnt{\ms{L}}, \dnt{\ms{K}})\), where \rle{join-nil} is the
identity, \rle{join-cons} is the sum of \(\dnt{\lwk{\ms{L}}{\ms{K}}}\) and
\(\dnt{\cwk{\Gamma}{\Delta}}\), and \rle{join-zero} is
\(\dnt{\lwk{\ms{L}}{\ms{K}}}\) postcomposed with the left injection.

\begin{figure}
  \begin{align*}
    \boxed{\dnt{A}: |\mc{C}|} & \qquad  
    \dnt{X} = \ms{base}(X) 
    \qquad
    \dnt{\mb{1}} = \mb{1}
    \qquad
    \dnt{\mb{2}} = \mb{2} = \mb{1} + \mb{1}
    \qquad \dnt{A \otimes B} = \dnt{A} \otimes \dnt{B} \\
    \boxed{\dnt{\Gamma}: |\mc{C}|} & \qquad 
    \dnt{\cdot} = \mb{1} \qquad 
    \dnt{\Gamma, \thyp{x}{A}{q}} = \dnt{\Gamma} \otimes \dnt{A} 
    \\
    \boxed{\dnt{\ms{L}}: |\mc{C}|} & \qquad 
    \dnt{\cdot} = \mb{0} \qquad
    \dnt{\ms{L}, \llhyp{\ell}{\Gamma}{A}} = 
      \dnt{\ms{L}} + \dnt{\Gamma} \otimes \dnt{A}
  \end{align*}
  \caption{Semantics for \isotopessa types and contexts}
  \Description{Semantics for isotope-SSA types and contexts}
  \label{fig:ssa-type-semantics}
\end{figure}

\begin{figure}
  \begin{gather*}
    % \boxed{\dnt{\tlin{A}{\taff}: \mc{C}_1^\tint(\dnt{A}, \mb{1})}}
    %   \\
    %   \dnt{\tlin{A}{\taff}} = \ms{drop}(\dnt{A})
    %   % \dnt{\tlin{X}{\taff}} = \ms{drop}(X) \qquad
    %   % \dnt{\tlin{A \otimes B}{\taff}} 
    %   %   = \dnt{\tlin{A}{\taff}} \otimes \dnt{\tlin{B}{\taff}}
    %   %   ; \lambda 
    %   \\
    % \boxed{\dnt{\tlin{A}{\trel}}
    %   : \mc{C}_1^\tint(\dnt{A}, \dnt{A} \otimes \dnt{A})}
    %   \\
    %   \dnt{\tlin{A}{\trel}} = \ms{split}(\dnt{A})
    %   % \dnt{\tlin{X}{\trel}} = \ms{split}(X) \qquad
    %   % \dnt{\tlin{A \otimes B}{\trel}} = 
    %   %   \dnt{\tlin{A}{\trel}} \otimes \dnt{\tlin{B}{\trel}}
    %   %   ;\alpha;\dnt{A} \otimes \sigma \otimes \dnt{B};\alpha 
    %   \\
    \boxed{\dnt{\csplits{\Gamma}{\Delta}{\Xi}}
      : \mc{C}_1^\tint(\dnt{\Gamma}, \dnt{\Delta} \otimes \dnt{\Xi})}
      \\
    \dnt{\csplits{\cdot}{\cdot}{\cdot}} = \lambda^{-1} \qquad
    \dnt{\csplits
      {\Gamma, \thyp{x}{A}{q}}
      {\Delta, \thyp{x}{A}{q}}
      {\Xi}} 
      = \dnt{\csplits{\Gamma}{\Delta}{\Xi}} \otimes \dnt{A};\alpha;\dnt{\Delta} \otimes \sigma;\alpha 
      \\
    \dnt{\csplits
      {\Gamma, \thyp{x}{A}{q}}
      {\Delta}
      {\Xi, \thyp{x}{A}{q}}} 
      = \dnt{\csplits{\Gamma}{\Delta}{\Xi}} \otimes \dnt{A};\alpha 
      \\
    \dnt{\csplits
      {\Gamma, \thyp{x}{A}{q}}
      {\Delta}
      {\Xi, \thyp{x}{A}{q}}} \
      = \dnt{\Gamma} \otimes \ms{split}(\dnt{A});
        \alpha;
        \dnt{\Delta} \otimes \sigma \otimes \dnt{A};
        \alpha 
        \\
    \dnt{\csplits
      {\Gamma, \thyp{x}{A}{q}}
      {\Delta}
      {\Xi}}
      = \dnt{\csplits{\Gamma}{\Delta}{\Xi}}
        \otimes \ms{drop}(\dnt{A})
      ; \rho 
      \\
    \boxed{\dnt{\cwk{\Gamma}{\Delta}}
      : \mc{C}_1^\tint(\dnt{\Gamma}, \dnt{\Delta})} 
      \qquad \qquad
      \dnt{\cwk{\Gamma}{\Delta}} 
      = \dnt{\csplits{\Gamma}{\Delta}{\cdot}};\rho 
      \\
    \boxed{\dnt{\lwk{\ms{L}}{\ms{K}}}
      : \mc{C}_1^\tint(\dnt{\ms{L}}, \dnt{\ms{K}})} 
      \\
    \dnt{\lwk{\cdot}{\cdot}} = \ms{id} \qquad
    \dnt{\lwk
      {\ms{L}, \llhyp{\ell}{\Gamma}{A}}
      {\ms{K}, \llhyp{\ell}{\Delta}{A}}}
      = \dnt{\lwk{\ms{L}}{\ms{K}}} + \upg{(\dnt{\cwk{\Gamma}{\Delta}} \otimes \dnt{A})}{} 
      \\
    \dnt{\lwk{\ms{L}}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}}
      = \dnt{\lwk{\ms{L}}{\ms{K}}};\iota_0
  \end{gather*}
  \caption{Semantics for \isotopessa structural judgements}
  \Description{Semantics for isotope-SSA structural judgements}
  \label{fig:ssa-structural-semantics}
\end{figure}

\begin{figure}
  \begin{gather*}
    \boxed{\dnt{\hasty{\Gamma}{p}{a}{A}{q}}
      : \mc{C}_p^q(\dnt{\Gamma}, \dnt{A})} 
      \\
    \dnt{\hasty{\Gamma}{p}{x}{A}{q}} 
      = \dnt{\cwk{\Gamma}{\thyp{x}{A}{q}}}
      \qquad
    \dnt{\hasty{\Gamma}{p}{f\;a}{B}{q}}
      = \upg{\dnt{\hasty{\Gamma}{1}{a}{A}{q}}}{p}
      ; \ms{inst}_p(f) 
      \\
    \dnt{\hasty{\Gamma}{p}{(a, b)}{A \otimes B}{q}}
      = \upg{(
        \dnt{\csplits{\Gamma}{\Delta}{\Xi}};
        \dnt{\hasty{\Delta}{1}{a}{A}{q}} \otimes
        \dnt{\hasty{\Xi}{1}{b}{B}{q}}
      )}{p} 
      \\
    \dnt{\hasty{\Gamma}{p}{()}{\mb{1}}{q}}
      = \upg{\dnt{\cwk{\Gamma}{\cdot}}}{p} 
      \quad
    \dnt{\hasty{\Gamma}{p}{\ctt}{\mb{2}}{q}}
      = \upg{\dnt{\cwk{\Gamma}{\cdot}}}{p};\iota_0
      \quad
    \dnt{\hasty{\Gamma}{p}{\cff}{\mb{2}}{q}}
      = \upg{\dnt{\cwk{\Gamma}{\cdot}}}{p};\iota_1 
      \\
    \dnt{\hasty{\Gamma}{p}{\letexpr{x}{a}{e}}{B}{q}}
      = \upg{(
        \dnt{\csplits{\Gamma}{\Delta}{\Xi}}
        ; \dnt{\Delta} \otimes \dnt{\hasty{\Xi}{1}{a}{A}{q}}
      )}{p};\dnt{\hasty{\Delta, \thyp{x}{A}{}}{p}{e}{B}{q}} 
      \\
    \dnt{\hasty{\Gamma}{p}{\letexpr{(x, y)}{a}{e}}{C}{q}}
      = \upg{(
        \dnt{\csplits{\Gamma}{\Delta}{\Xi}}
        ; \dnt{\Delta} \otimes \dnt{\hasty{\Xi}{1}{a}{A \otimes B}{q}}
      )}{p}
      ; 
      \\ \qquad \qquad \qquad \qquad \qquad \alpha
      ; \dnt{\hasty{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{p}{e}{C}{q}}
    %   \\
    % \dnt{\hasty{\Gamma}{0}{\lbsplice{\ell}{A}{t}}{A}{\varnothing}}
    %   = \dnt{\haslb{\Gamma}{t}{\llhyp{\ell}{\cdot}{A}}};\alpha_+
  \end{gather*}
  \caption{Semantics for \isotopessa expressions}
  \Description{Semantics for isotope-SSA expressions}
  \label{fig:ssa-term-semantics}
\end{figure}

We can now interpret a term \(\hasty{\Gamma}{p}{a}{A}{q}\) as a
\textit{morphism} taking \(\dnt{\Gamma}\) into \(\dnt{A}\) with
\textit{centrality} \(p\) and \textit{quantity} \(q\) (i.e., lying in
\(\mc{C}_p^q\)); our rules for doing so are given in Figure
\ref{fig:ssa-term-semantics}. We will abuse syntax slightly and write
\(\dnt{\text{"judgement"}}\) for \(\dnt{D}\), where \(D\) is a derivation of
"judgement"; in general, it can be proven that the semantics of any two such
derivations are equal. We also define the syntax sugar \(\upg{f}{p}\), where
\(\upg{f}{0}\) is \(\upg{f}{}\) for morphisms in \(\mc{C}_1\) and simply \(f\)
for morphisms in \(\mc{C}_0\), and \(\upg{f}{1}\) is the identity functor on
\(\mc{C}_1\).

Our treatment of \rle{var} is quite standard: semantics of \(x\) is simply
discarding all other variables in \(\Gamma\), i.e.
\(\dnt{\cwk{\Gamma}{\thyp{x}{A}{q}}}\). Similarly, to type a function
application \rle{app}, we first take \(\dnt{\hasty{\Gamma}{1}{a}{A}{q}}\), which
is guaranteed to be central, lift to \(\mc{C}_p\), and then simply postcompose
with \(f\). \rle{pair} is handled by precomposing the context split
\(\csplits{\Gamma}{\Delta}{\Xi}\) with the tensor product of the components
\(\dnt{\hasty{\Delta}{1}{a}{A}{q}}\), \(\dnt{\hasty{\Xi}{1}{b}{B}{q}}\); this is
well-defined since both components are central and \(\mc{C}_1\) is monoidal. The
semantics of \rle{unit} is simply dropping the entire context, while \rle{true}
and \rle{false} are interpreted as first dropping the context and then injecting
into the appropriate branch of the boolean type. Finally, \rle{let} and
\rle{let2} are handled by first splitting the context, then passing the right
hand component to the semantics of the bound term
\(\dnt{\hasty{\Xi}{1}{a}{A}{q}}\) or \(\dnt{\hasty{\Xi}{1}{a}{A \otimes
B}{q}}\), and then, lifting to \(\mc{C}_p\) as necessary, after reassociating
with \(\alpha\), passing everything to the inner term \(\dnt{\hasty{\Delta,
\thyp{x}{A}{q}}{p}{e}{B}{q}}\) or \(\dnt{\hasty{\Delta, \thyp{x}{A}{q},
\thyp{y}{B}{q}}{p}{e}{C}{q}}\).
% Note that this inner term is not necessarily central, unlike the bound term!

\begin{figure}
  \begin{gather*}
    \boxed{\dnt{\haslb{\Gamma}{t}{\ms{L}}}
      : \mc{C}_0^\varnothing(\dnt{\Gamma}, \dnt{\ms{L}})} \\
    \dnt{\haslb{\Gamma}{\lbrb{\ell}{a}}{\ms{L}}}
      = \upg{(\dnt{\csplits{\Gamma}{\Delta}{\Xi}}
      ; \dnt{\Delta} \otimes \dnt{\hasty{\Xi}{1}{a}{A}{\tint}})}{}
      ; \dnt{\lwk{\llhyp{\ell}{\Delta}{A}}{\ms{L}}} \\
    \dnt{\haslb{\Gamma}{\ite{e}{s}{t}}{\ms{L}}}
      = \upg{(\dnt{\csplits{\Gamma}{\Delta}{\Xi}}{}
      ; \dnt{\Delta} \otimes \dnt{\hasty{\Xi}{1}{e}{\mb{2}}{\tint}})}
      ; \delta ;
      \\ \qquad \qquad \qquad \qquad
      [
        \dnt{\haslb{\Delta}{s}{\ms{L}}}, 
        \dnt{\haslb{\Delta}{t}{\ms{L}}} 
      ]
      \\
    \dnt{\haslb{\Gamma}{\letstmt{x}{a}{t}}{\ms{L}}}
      = \upg{\dnt{\csplits{\Gamma}{\Delta}{\Xi}}}{}
      ; \dnt{\Delta} \otimes \dnt{\hasty{\Gamma}{0}{a}{A}{q}}
      ; \dnt{\haslb{\Delta, \thyp{x}{A}{}}{t}{\ms{L}}}
      \\
    \dnt{\haslb{\Gamma}{\letstmt{(x, y)}{a}{t}}{\ms{L}}}
      = \upg{\dnt{\csplits{\Gamma}{\Delta}{\Xi}}}{}
      ; \dnt{\Delta} \otimes \dnt{\hasty{\Gamma}{0}{a}{A \otimes B}{q}}
      ; \alpha
      ; \dnt{\haslb{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{t}{\ms{L}}}
      \\
    \dnt{\haslb{\Gamma}{\ewhere{t}{L}}{\ms{K}}}
      = \dnt{\haslb{\Gamma}{t}{\ms{L}}}
      ; \dnt{\lhaslb{\ms{L}}{L}{\ms{K}}}^\dagger
      \\
    \boxed{\dnt{\lhaslb{\ms{L}}{L}{\ms{K}}}
      : \mc{C}_0^\varnothing(\dnt{\ms{L}}, \dnt{\ms{K}} + \dnt{\ms{L}})}
      \\
    \dnt{\lhaslb{\ms{L}}{L, \lwbranch{\ell}{x: A}{t}}{\ms{K}}}
      = \dnt{\lhaslb{\ms{L}}{L}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}}
      ; [\dnt{\ms{K}} + \dnt{\haslb{\Gamma, \thyp{x}{A}{}}{t}{\ms{L}}}, \iota_1]
      \\
    \dnt{\lhaslb{\ms{L}}{\cdot}{\ms{L}}}
      = \iota_0
  \end{gather*}
  \caption{Semantics for \isotopessa blocks}
  \Description{Semantics for isotope-SSA blocks}
  \label{fig:ssa-block-semantics}
\end{figure}

We interpret \textit{blocks} \(\haslb{\Gamma}{t}{\ms{L}}\) as \textit{impure}
morphisms from \(\dnt{\Gamma}\) to \(\dnt{\ms{L}}\). \(\Gamma\) represents all
the variables live on entry to the block, while \(\ms{L}\) represents all valid
targets, each annotated with live variables and an argument type; our semantics
are given in Figure \ref{fig:ssa-block-semantics}. \rle{br} is interpreted by
first splitting the context via \(\dnt{\csplits{\Gamma}{\Delta}{\Xi}}\), and
then using \(\Xi\) to compute the target's argument. The result, along with the
unchanged live variables \(\Delta\), is then passed to
\(\lwk{\llhyp{\ell}{\Delta}{A}}{\ms{L}}\) to obtain the desired target.
Similarly, we interpret \rle{ite} by first splitting the context via
\(\dnt{\csplits{\Gamma}{\Delta}{\Xi}}\) and using \(\Xi\) to compute the
discriminant. The "if" statement itself is then simply an application of the
distributor \(\delta: \dnt{\Delta} \otimes (\mb{1} + \mb{1}) \to \dnt{\Delta} +
\dnt{\Delta}\); we pass the left of the coproduct to the true branch
\(\dnt{\haslb{\Xi}{s}{\ms{L}}}\) and the right to the false branch
\(\dnt{\haslb{\Xi}{t}{\ms{L}}}\). The semantics of \rle{let-blk} and
\rle{let2-blk} are analogous to those of \rle{let} and \rle{let2}, except that
now the bound term \(a\) is allowed to be impure.

Finally, we must give semantics for \rle{where}. To do so, we interpret a
control-flow graph \(\lhaslb{\ms{L}}{L}{\ms{K}}\) as a morphism from
\(\dnt{\ms{L}}\) to \(\dnt{\ms{K}} + \dnt{\ms{L}}\), where the left of the
coproduct denotes branching to one of the labels in \(\ms{K}\) (the
\textit{output}), while the right of the coproduct denotes a recursive call to a
label in \(\ms{L}\) (the \textit{recursive input}). In particular, \rle{nil-br}
simply forwards all calls to one of \(\ms{L}\) to the output unchanged.
\rle{cons-br} is a little bit more complicated: first we execute
\(\dnt{\lhaslb{\ms{L}}{L}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}}\), which yields
either an output \(\dnt{\ms{K}, \llhyp{\ell}{\Gamma}{A}}\) or a recursive input
\(\dnt{\ms{L}}\), which we pass through unchanged as a recursive input. In the
former case, we can interpret this as either a value of \(\dnt{\ms{K}}\), which
we pass through unchanged as output, or \(\dnt{\Gamma} \otimes \dnt{A}\); which
we can pass as input to \(\dnt{\haslb{\Gamma, \thyp{x}{A}{q}}{t}{\ms{L}}}\),
giving us a value of \(\ms{L}\) we can use as recursive input. Note that, even
if \(t\) yields an output label in \(\ms{K}\), we still return a
\textit{recursive input}, not an output; in this case, the \textit{next}
iteration will return it as an output. We can now quite easily state the
semantics of \rle{where}: we simply precompose the semantics of the entry block
\(\dnt{\haslb{\Gamma}{t}{\ms{L}}}\) with the \textit{iterated} semantics of the
associated control-flow graph, \(\dnt{\lhaslb{\ms{L}}{L}{\ms{K}}}^\dagger\).

\subsection{Semantic Metatheory}

Now that we have a semantics for \isotopessa programs, we can show that it
matches the syntax by showing that the syntactic and semantic notions of
substitution coincide. We begin by giving a semantics for
\(\issubst{\gamma}{\Theta}{\Gamma}\) in Figure \ref{fig:ssa-subst-semantics},
which we will interpret as a pure morphism taking \(\dnt{\Theta}\) to
\(\dnt{\Gamma}\). This morphism can simply be viewed as computing all
\(\gamma(x)\) for \(\thyp{x}{A}{q}\) in \(\Gamma\); in particular, we interpret
\rle{subst-cons} by splitting the context via
\(\dnt{\csplits{\Theta}{\Theta_\Gamma}{\Theta_x}}\), passing
\(\dnt{\Theta_\Gamma}\) to the remainder of the substitution
\(\dnt{\issubst{\gamma}{\Gamma}{\Theta_\Gamma}}\) to compute the rest of the
\(\gamma(y)\) in \(\Gamma\) recursively, and using \(\dnt{\Theta_x}\) to compute
\(\gamma(x)\). Similarly, we interpret label-substitutions as morphisms from
\(\dnt{\ms{L}}\) to \(\dnt{\ms{K}}\) which, for each input label in \(\ms{L}\),
performs a computation targeting \(\ms{K}\). In particular, \rle{lb-subst-nil}
is interpreted as the zero morphism \(0_{\dnt{\ms{K}}}:
\mc{C}_0^\varnothing(\mb{0}, \dnt{\ms{K}})\), while \rle{lb-subst-cons} takes in
the coproduct \(\dnt{\ms{L}} + \dnt{\Gamma} \otimes \dnt{A}\), sending the
former to \(\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{A}}}\) and the latter to the
denotation of \(\mc{L}(\lbl{\ell}, x)\).

\begin{figure}
  \begin{gather*}
    \boxed{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}: \mc{C}_1^\varnothing(\dnt{\Theta}, \dnt{\Gamma})} 
    \\
    \dnt{\issubst{\cdot}{\cdot}{\cdot}} = \ms{id}
    \qquad
    \dnt{\issubst{\gamma}{\Theta}{\Gamma, \thyp{x}{A}{q}}}
      = \dnt{\csplits{\Theta}{\Theta_\Gamma}{\Theta_x}};\dnt{\issubst{\gamma}{\Theta_\Gamma}{\Gamma}} \otimes \dnt{\hasty{\Theta_x}{1}{a}{A}{q}}
    \\
    \boxed{\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}: \mc{C}_0^\varnothing(\dnt{\ms{L}}, \dnt{\ms{K}})} 
    \\
    \dnt{\lbsubst{\mc{L}}{\cdot}{\ms{K}}} = 0_{\ms{K}}
    \qquad
    \dnt{\lbsubst{\mc{L}}{\ms{L}, \llhyp{\ell}{\Gamma}{A}}{\ms{K}}}
    = [\dnt{\lbsubst{\ms{L}}{\ms{L}}{\ms{K}}}, \dnt{\haslb{\Gamma, \thyp{x}{A}{}}{\mc{L}(\lbl{\ell}, x)}{\ms{K}}}]
  \end{gather*}
  \caption{Semantics for \isotopessa substitutions}
  \Description{Semantics for isotope-SSA substitutions}
  \label{fig:ssa-subst-semantics}
\end{figure}

We can now state the theorem of semantic substitution, which essentially states
that the semantics of a substituted term or block is the same as the semantics
of that term or block preceded by let-bindings corresponding to the variables in
the substitution, which is the same as the semantics of the original term or
block precomposed with the semantics of the substitution. Similarly, the
semantics of a \textit{label-substituted} block is the same as the semantics of
the original block \textit{postcomposed} with the semantics of that
label-substitution. Stated formally, we obtain:
\begin{theorem}[Semantic Substitution] 
  Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
  \begin{itemize}
    \item For all \(\hasty{\Gamma}{p}{a}{A}{q}\), 
    \(
      \upg{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}}{p}
      ;\dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
      = \dnt{\hasty{\Theta}{p}{\exprletsubst{\substctx{\gamma}{\Gamma}}{a}}{A}{q}}
      = \dnt{\hasty{\Theta}{p}{[\gamma]a}{A}{q}}
    \)
    \item For all 
      \(\haslb{\Gamma}{t}{\ms{L}}\), 
      \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\), 
      \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\),
    \[
      \upg{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}}{}
      ; \dnt{\haslb{\Gamma}{t}{\ms{L}}}
      = \dnt{\haslb{\Theta}{\exprletsubst{\substctx{\gamma}{\Gamma}}{t}}{\ms{K}}} 
      ; \dnt{\lbsubst{\substlbs{\gamma}{\ms{K}}}{\ms{K}}{\ms{L}}}
      = \dnt{\haslb{\Theta}{[\gamma]t}{\ms{K}}} 
      ; \dnt{\lbsubst{\substlbs{\gamma}{\ms{K}}}{\ms{K}}{\ms{L}}}
    \]
    \item For all 
      \(\lhaslb{\ms{W}}{L}{\ms{L}}\), 
      \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\),
      \(\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}\),
      \[
        \dnt{\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}}
        ; \dnt{\lhaslb{\ms{W'}}{[\gamma]L}{\ms{K}}}
        = \dnt{\lhaslb{\ms{W}}{L}{\ms{L}}}
        ; \dnt{\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}}
        + \dnt{\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}}
      \]
  \end{itemize}  
  Similarly, given \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\),
  \begin{itemize} 
    \item For all \(\haslb{\Gamma}{t}{\ms{L}}\), we have
    \(
      \dnt{\haslb{\Gamma}{t}{\ms{L}}};\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
      = \dnt{\haslb{\Gamma}{[\mc{L}]t}{\ms{K}}}
    \)
    \item For all \(\lhaslb{\ms{W}_{\ms{L}}}{L}{\ms{L}}\), we have
    \[
      \dnt{\lhaslb{\ms{W}_{\ms{L}}}{L}{\ms{L}}};\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
      + \dnt{\lbsubst{\mc{L}}{\ms{W}_{\ms{L}}}{\ms{W}_{\ms{K}}}}
      = \dnt{\lbsubst{\mc{L}}{\ms{W}_{\ms{L}}}{\ms{W}_{\ms{K}}}};\dnt{\lhaslb{\ms{W}_{\ms{K}}}{[\mc{L}]L}{\ms{K}}}
    \]
    where all variables in \(\ms{W}\) not in \(\ms{L}\) are fresh w.r.t \(\mc{L}\).
  \end{itemize}
\end{theorem}

\subsection{Optimisations}

Because we have a denotational semantics, it is easy to validate equations
between terms, blocks, and CFGs. Even the basic proof of the soundness of
substitution already lets us prove the soundness of many fundamental compiler
optimisations, such as global value numbering (let-introduction for pure
expressions), some forms of dead code elimination (i.e., eliminating unused
affine bindings), and strength reduction with loop hoisting (let-introduction
for relevant expressions).

Substitution only applies to central terms, because those are the only terms for
which substitution is sound. Furthermore, there are many other optimisations
that follow from the beta-eta theory of the language, as well as specific
properties of model-dependent operations (such as memory operations). Because we
have a compositional denotational semantics, equations automatically lift into a
full congruence. This lets us justify optimisations such as scalar replacement
of aggregates, eliminating redundant writes, reordering independent memory
operations, and memory-to-register transformations.

Because we have a type-theoretic presentation of control-flow graphs, we also
have a substitution principle and equational theory for them. This lets us show
the soundness of inlining, loop unrolling, loop fusion, dead-code elimination,
and other optimisations which change the structure of the control-flow graph.
Note that we have not yet proved anything about the dataflow analyses used to
find the rewrites -- we have only shown that the rewrites themselves are
semantically justified.

\section{Concrete Models}

In this section, we give cncrete models of the axiomatisation given in Section
\ref{sec:semantics} to with successively more sophisticated features such as
weak memory and separation logic.

\subsection{Trace Models}

\label{ssec:trace-monads}

We will begin by giving a useful foundation for a wide class of \isotopessa
models: the "trace models" over \(\mb{Set}\), which accumulate a (potentially
infinite) nondeterministic trace of monoidal "events". These will later come in
handy for reasoning about output behaviour and weak memory. We start with some
definitions:
\begin{definition}[Relevant, Affine Monad]
  A monad \(\mb{T}\) on \(\ms{Set}\) is \textbf{relevant} if its Kleisli
  category satisfies \(f;\ms{split} = \ms{split};f \ltimes f = \ms{split};f
  \rtimes f\), where \(\ms{split} = \upg{\lambda a. (a, a)}{}\). \(\mb{T}\) is
  \textbf{affine} if its Kleisli category satisfies \(f;\ms{drop} = \ms{drop}\),
  where \(\ms{drop} = \upg{\lambda a.()}{}\). A monad is \textbf{intuitionistic}
  if it is both relevant and affine.
\end{definition}
\begin{definition}[Submonad] 
  A \textbf{submonad} \(\mb{S}\) of a monad \(\mb{T}\) on \(\ms{Set}\) is a
  collection of sets such that, for all \(A\), \(\mb{S}\;A \subseteq
  \mb{T}\;A\), \(\forall a \in A, \eta\;a \in \mb{S}\;A\), and, for all \(s \in
  \mb{S}\;A\), \(f: A \to \mb{S}\;B\), \(\ms{bind}\;s\;f \in \mb{S}\;A\). We say
  that a submonad of \(\mb{T}\) is \textbf{Elgot} if \(\mb{T}\) is Elgot and,
  for all \(f: A \to \mb{S}(B + A)\), \(f^\dagger: A \to \mb{S}\;B\) where
  \(f^\dagger\) is computed as in \(\mb{T}\). Note that \(\mb{S}\) is also an
  (Elgot) monad in its own right, inheriting \(\mu, \eta\) and \(\cdot^\dagger\)
  from \(\mb{T}\). The \textbf{unit submonad} \(\eta\) of every monad \(\mb{T}\)
  is given by the map \(\eta(A) = \{\eta_A\;a \mid a \in A\}\). Given submonads
  \(\mb{S}_i\) of \(\mb{T}\), their elementwise intersections
  \(\bigwedge_i\mb{S}_i = \lambda A. \bigcap_i \mb{S}_i\;A\) are submonads,
  forming a complete meet-semilattice with bottom \(\eta\) and top \(\mb{T}\).
  In particular, the intersection of a relevant/affine/commutative submonad and
  another submonad is relevant/affine/commutative respectively.
\end{definition}
\begin{definition}[Stream Action]
  A \textbf{stream action} of a monoid \(M\) on a set \(I\) equipped with a
  monoid action \(\cdot: M \times I \to I\) is a function \(\Theta: M^\omega \to
  I\) mapping (infinite) streams of \(M\) to elements of \(I\) satisfying
  \(
    m \cdot \Theta(\ell) = \Theta(m \colon \ell)
  \),
  where \(m \colon \ell\) denotes the stream formed by prefixing \(m\) to the
  stream \(\ell\).
  Some basic examples include:
  \begin{enumerate}
    \item For any monoid \(M\), the unique map \(\Theta: M^\omega \to \mathbf{1}\)
    forms a stream action on the unit set.
    % \item In particular, given monoids \(M_i\) acting on sets \(I_i\) with action
    % \(\cdot_i\) and stream actions \(\Theta_i: M_i^\omega \to I_i\), \(\Theta =
    % \langle \Theta_i \circ \ms{fmap}\;\pi_i\rangle_i: (\Pi_iM_i)^\omega \to
    % \Pi_iI_i\) is a stream action compatible with the action \(m \cdot l = (m_i
    % \cdot_i l_i)_i\).
    \item The concatenation map \(\Theta: (A^*)^\omega \to A^\omega\) is a
    stream action of \(A^*\), the set of (finite) lists of \(A\), on \(A^{\leq
    \omega}\), the set of (\textit{potentially} infinite) sequences of \(A\),
    where \(\ell \cdot s\) prepends the \(\ell\) to \(s\).
    % \item The sum map \(\Sigma: \nats^\omega \to \nats \cup \{\infty\}\) is a
    % stream action of \(\nats\) on \(\nats \cup \{\infty\}\), where \(\nats\) acts
    % on \(\nats \cup \{\infty\}\) by addition.
    % \item Given a monoid \(M\) acting on sets \(I_i\) with action \(\cdot_i\) and
    % stream actions \(\Theta_i: M^\omega \to I_i\), \(\Theta =
    % \langle\Theta_i\rangle_i: M^\omega \to \Pi_iI_i\) is a stream action
    % compatible with the action \(m \cdot l = (m \cdot_i l_i)_i\).
  \end{enumerate}
\end{definition}
% We define the \textbf{stream center} of \(M\) with respect to \(I\), denoted
% \(\mc{SZ}(M, I)\), as the subset of the center of \(M\), \(Z(M)\), consisting of
% those elements \(m \in Z(M)\) such that for all \(i \in I\), \(m \cdot i = i\).
% For instance, \(\mc{SZ}(M, \mathbf{1}) = Z(M)\), and \(\mc{SZ}(A^*, A^{\leq
% \omega}) = \{[\,]\}\). Generally, \(\mc{SZ}(M, I)\) always contains
% the unit of \(M\). 
Using stream actions to determine what it means to perform a monoidal action
"infinitely many times," we can now define the \textit{trace monad} of a stream
action as follows:
\begin{definition}[Trace Monad]
  Given a stream action \(\Theta: M^\omega \to I\), we can define the
  \textbf{nondeterministic trace monad} \(\ms{Traces}\;\Theta\;A = \mc{P}^+(A
  \times M + I)\) with unit \(\eta_A = \lambda a.\iota_0 (a, 1)\) and
  multiplication
  \(
    \mu_A\;T = \{[\lambda (a, m). m \cdot a, \ms{id}](t) \mid t \in T\}
  \),
  where the action of \(M\) on traces is given by \(m \cdot \iota_0 (a, m') =
  \iota_0 (a, mm')\) and \(m \cdot \iota_1 t = \iota_1 (m \cdot t)\). We can
  equip this monad with an iteration operator \(f^\dagger(a) = f^\infty(a) \cup
  \bigcup_{n \in \nats}\{\iota_0 (\iota_0\;b, m) \in f_n(a)\} \cup \{\iota_1 t
  \in f_n(a)\}\), yielding an Elgot structure, where
  \[
    f_0 = f, \quad f_{n + 1} = f \gg [\upg{\iota_0}{}, f_n], \quad
    f^\infty(a) = \{\Theta (\lambda n. m_n) \mid a_0 = a \land \forall n, \exists a_{n + 1}. \iota_0 (\iota_1\;a_{n + 1}, m_n) \in f(a_n)\}
  \]
\end{definition}
We can use the trace monad described above to derive some interesting examples
of submonads:
\begin{enumerate}
  \item If \(C \subseteq Z(M)\) is submonoid of \(M\), then \(A \mapsto
  \mc{P}^+(A \times C + \mb{0}) \subseteq \ms{Traces}\;\Theta\;A\) is a submonad.
  \item In particular, if \(M_C \subseteq Z(M)\) is a central submonoid of \(M\)
  such that \(\forall m \in M_C, \forall i \in I, m \cdot i = i\), then \(A
  \mapsto \mc{P}^+(A \times M_C + \mb{0}) \subseteq \ms{Traces}\;\Theta\;A\) is
  a submonad whose Kleisli category is a central subcategory of the Kleisli
  category of \(\ms{Traces}\;\Theta\;A\). Note \(M_C = \{1\}\) is always such a
  submonoid.
  \label{item:central-submonad}
  \item Similarly, if \(M_R \subseteq M_C\) is a submonoid such that \(\forall m
  \in M_R, m \cdot m = m\), then \(A \mapsto \{S \in \mc{P}^+(A \times M_R + I)
  \mid |S| = 1\} \subseteq \ms{Traces}\;\Theta\;A\) is a \textit{relevant}
  submonad. Note \(M_R = \{1\}\) is always such a submonoid.
  \label{item:relevant-submonad}
  \item Finally, \(A \mapsto \mc{P}^+(A \times \{1\} + \mb{0}) \subseteq
  \ms{Traces}\;\Theta\;A\) is an \textit{affine} submonad.
  \label{item:affine-submonad}
\end{enumerate}
Since every commutative submonad induces a central subcategory of the Kleisli
category of the parent monad, it follows that a triple of a relevant, affine,
and commutative submonad make the Kleisli category of a monad \(\mb{T}\) into a
substructural effectful category with all objects both affine and relevant. In
particular, \ref{item:central-submonad}, \ref{item:relevant-submonad},
\ref{item:affine-submonad} equip the Kleisli category of
\(\ms{Traces}\;\Theta\;A\) with the structure of a substructural effectful
category given choices of \(M_C\), \(M_R\).

\subsection{Heaps, Printing, and Separation Logic}

\label{ssec:separation}

We can now introduce the \textbf{printing monad}, which is simply defined as
\(\ms{Print}\;A \equiv \ms{Traces}\;\Sigma\), where \(\Sigma:
(\ms{Byte}^*)^\omega \to \ms{Byte}^{\leq \omega}\) denotes concatenation of
bytestrings into a potentially infinite bytestream. This can be made into a
substructural effectful category by choosing \(M_C = M_R = \{[]\}\), giving us
an \isotopessa model supporting instructions \(\ms{print} \in
\mc{I}^\varnothing_0(\ms{Byte}^*, \mb{1})\) and \(\ms{nondet} \in
\mc{I}^\taff_1(\mb{1}, A)\) (for \(A\) nonempty), with semantics
\(\dnt{\ms{print}} = \lambda b.\{\iota_0 ((), b)\}\), \(\dnt{\ms{nondet}} =
\lambda (). \{\iota_0 (a, []) \mid a \in A\}\).

We can make things a little more interesting by adding a heap to the mix,
defining our monad \(\ms{Comp} = \ms{StateT}\;\ms{Heap}\;\ms{Print}\), where
\(\ms{Heap} = \nats \rightharpoonup \nats\) is simply a partial function with
finite support. The Kleisli category of this monad is a substructural effectful
category and inherits an Elgot structure from that on \(\ms{Set}_{\ms{Print}}\);
we therefore have an \isotopessa model supporting the instructions \(\ms{set}:
\mc{I}^\varnothing_0(\nats \times \nats, \mb{1})\), \(\ms{get}:
\mc{I}^\varnothing_0(\nats, \nats)\), \(\ms{alloc}: \mc{I}^\varnothing_0(\nats,
\nats)\), and \(\ms{free}: \mc{I}^\varnothing_0(\nats, \mb{1})\). We can assign
semantics to \(\ms{set}\) in the standard fashion, with \(\dnt{\ms{set}} =
\lambda (p, v)\;h. \{\iota_0((), [p \mapsto v]h, [])\}\). \(\ms{get}\) is a bit
more tricky, since it is unclear what to do when we try to access uninitialized
memory: one option is simply to return an arbitrary value, with \(\dnt{\ms{get}}
= \lambda p\;h. \{\iota_0(v, h, []) \mid v = h\;p \lor p \notin h\}\). Finally,
\(\dnt{\ms{alloc}} = \lambda v\;h.\{\iota_0\;(p, h', []) \mid h' = h \sqcup p
\mapsto v\}\) simply fills a random empty heap cell with the provided value,
returning a pointer to the cell, while \(\dnt{\ms{free}} = \lambda p\;h.
\{\iota_0\;((), h \setminus p, [])\}\).

Building on the ideas in \citet{mellies-ftrs}, we show how a simple separation
logic based on \citet{reynolds-separation-2002} slightly adapted to the
categorical setting gives us a very rich \isotopessa model supporting a basic
notion of refinement types. Taking as objects pairs \((A, \Phi)\), where \(\Phi: A \to
\ms{Heap} \to \ms{Prop}\), we can define the category \(\ms{Sep}\) to have
morphisms
\begin{equation}
  \ms{Sep}((A, \Phi), (B, \Psi)) = \{f: A \to \ms{Comp}\;B 
    \mid \forall \phi: \ms{Heap} \to \ms{Prop}, \ms{hoare}\;f\;(\Phi \ast \phi)\;(\Psi \ast \phi)\}
\end{equation}
where
\(
  \ms{hoare}\;f\;\Phi\;\Psi = \forall a\;h. \Phi\;a\;h \implies \forall \iota_0\;(b, h') \in f\;a\;h. \Psi\;b\;h'
\)
corresponds to the usual notion of a partial correctness triple, and the
separating conjunction of heap-predicates is defined as
\begin{equation}
  \begin{gathered}
  \varphi \ast \psi = \lambda h. \exists h_1, h_2. h = h_1 \sqcup h_2 \land \varphi h_1 \land \psi h_2 
  \qquad
  \varphi \ast \Phi = \lambda a. \varphi \ast \Phi\;a
  \qquad
  \Phi \ast \varphi = \lambda a. \Phi\;a \ast \varphi \\
  \Phi \ast \Psi = \lambda (a, b). \Phi\;a \ast \Psi\;b: A \times B \to \ms{Heap} \to \ms{Prop}
  \end{gathered}
\end{equation}
We can naturally define a tensor product on objects \((A, \Phi) \otimes (B,
\Psi) = (A \times B, \Phi \ast \Psi)\), making \(\ms{Sep}\) into a substructural
effectful category with:
\begin{itemize}
  \item Monoidal unit \((\mb{1}, \ms{emp})\), where \(\ms{emp}\) is the
  predicate which holds only for the empty heap
  \item Associators and inverse associators given by \(\alpha_{(A, \Phi), (B,
  \Psi), (C, \Theta)} = \alpha_{A, B, C}\)
  %, which are valid since \(\ast\) is associative
  \item Unitors and inverse unitors given by \(\lambda_{(A, \Phi)} =
  \lambda_A\), \(\rho_{(A, \Phi)} = \rho_A\)
  %, which are valid since \(\ms{emp}\)
  is the unit of \(\ast\)
  \item \((C, \Theta) \otimes f = C \times f\) and \(f \otimes (C, \Theta) = f
  \times C\), which are valid definitions by the frame rule
  \item Affine objects those for which \(\ms{drop} = \upg{\lambda a. ()}{} \in
  \ms{Sep}((A, \Phi), (\mb{1}, \ms{emp}))\), and relevant objects those for
  which \(\ms{split} = \upg{\lambda a. (a, a)} \in \ms{Sep}((A, \Phi), (A, \Phi
  * \Phi))\), which for this logic means pairs \((A, \Phi)\) where \(\Phi\) does
  not depend on the heap.
  \item Central morphisms to be those which do not print, diverge, \textit{or}
  modify the heap. % We will consider how to relax this condition later.
  \item Relevant and affine morphisms to be the lifts of relevant/affine
  morphisms in \(\ms{Print}\)
\end{itemize}
\(\ms{Sep}\) also has coproducts given by \((A, \Phi) + (B, \Psi) = (A + B, [\Phi, \Psi])\), which allow us to equip it with an Elgot structure by taking advantage of the fact that
\begin{equation}
  f \in \ms{Sep}((A, \Phi), (B + A, [\Psi, \Phi])) \implies f^\dagger \in \ms{Sep}((A, \Phi), (B, \Psi))
\end{equation}
This implies that \(\ms{Sep}\) is an \isotopessa model supporting the following
interesting operations:
\begin{itemize}
  \item \(\ms{coe}: \mc{I}^\tint_1((A, \Phi), (A, \Psi))\) where \(\forall a,
  \Phi\;a \implies \Psi\;a\), implemented by the identity morphism
  \item \(\ms{print}: \mc{I}^\varnothing_0((\ms{Byte}^*, \ms{emp}), (\mb{1},
  \ms{emp}))\) and \(\ms{nondet}: \mc{I}^\taff_1((\mb{1}, \ms{emp}), (A, \ms{emp}))\) for
  \(A\) nonempty
  \item \(\ms{set}: \mc{I}^\varnothing_0((\nats \times \nats, \lambda (\ell,
  \cdot). \exists v. \ell \mapsto v), (\nats, \lambda \ell. \exists v. \ell
  \mapsto v))\) and \(\ms{get}: \mc{I}^\varnothing_0((\nats \times \nats,
  \lambda (-, \ell). \exists v. \ell \mapsto v), (\nats, \lambda \ell. \exists
  v. \ell \mapsto v))\), which are modified slightly to return the input address.
  \item \(\ms{alloc}: \mc{I}^\varnothing_0((\nats, \ms{emp}), (\nats, \lambda
  \ell. \exists v. \ell \mapsto v))\) and \(\ms{free}: \mc{I}^\varnothing_0((\nats, \lambda \ell. \exists v. \ell \mapsto v), (\mb{1}, \ms{emp}))\)
\end{itemize}
We can clearly see that this is a refinement type system in the sence of
\cite{mellies-ftrs}: in particular, there is a forgetful functor \((A, \Phi)
\mapsto A\) which is simply the identity on morphisms. Unfortunately, this means
that our notion of equality is a little too stringent, since we need to consider
equality on inputs which do not match our preconditions. Thankfully, we can fix
this problem quite easily by quotienting morphisms over the equivalence relation
\begin{equation}
  \forall f, g: \ms{Sep}((A, \Phi), (B, \Psi)).
  f \simeq g \iff \forall a, h. \Phi\;a\;h \implies f\;a\;h = g\;a\;h
\end{equation}
yielding the new category \(\ms{SComp}\). In particular, this allows us to widen
our definition of central morphisms to include all morphisms which do not print
or diverge, including in particular \(\ms{set}, \ms{get}, \ms{alloc}\) and
\(\ms{free}\), which we can in fact mark relevant and affine as well, allowing
us, for example, to make use of intermediate allocations in an otherwise pure
computation.

\subsection{Weak Memory}

Another interesting use of the trace monad to build a sophisticated \isotopessa
model is to build a model of TSO-style weak memory based on \citet{sparky}. For
simplicity, we will consider an infinite set of named locations \(x, y, z \in
\ms{Loc}\) which are subject to concurrent modification by all threads via TSO
reads, writes, and fences. We begin with the following definitions:
\begin{definition}[Pomset] 
  A \textbf{pomset} \(\alpha\) over a set of actions \(\mc{A}\) with a
  distinguished null action "\textbf{tick}" \(\delta \in \mc{A}\) is a
  \textit{nonempty} partially-ordered \textbf{carrier set} \(P\) such that every
  \(p \in P\) has finitely many predecessors equipped with a mapping \(\alpha: P
  \to \mc{A}\) quotiented over the removal of finitely many \(\delta\). A pomset
  is \textbf{finite} if its carrier set is. We will write unordered pomsets
  using multiset notation, e.g. \(\{a, a, b\}\), and linearly ordered pomsets
  using list notation, e.g. \([a, a, b]\).
\end{definition}
(Finite) pomsets form a monoid under sequential composition \(\alpha;\beta\),
which is defined by the function \([\alpha, \beta]: \ms{trim}(P + Q) \to
\mc{A}\), where \(P + Q\) is given the lexicographic ordering and
\(\ms{trim}(P)\) removes all elements of \(P\) with infinitely many
predecessors. They also form a monoid under parallel composition, defining
\(\alpha || \beta = [\alpha, \beta]: P + Q \to \mc{A}\) where \(P + Q\) is given
the standard partial ordering (with elements of \(P\) and \(Q\) incomparable).
Note in both cases the monoidal unit is \(\{\delta\}\), since we only consider
nonempty pomsets but allow the removal of finitely many ticks. Given any
partially ordered set \(N\) and a family of pomsets \(\alpha_n\) for \(n \in
N\), we can define their \textit{sum} \(\Sigma_n\alpha_n\) to be given by the
function \((n, a) \mapsto \alpha_n\;a: \ms{trim}(\Sigma_nP_n) \to \mc{A}\),
where the dependent product \(\Sigma_nP_n\) is given the lexicographic order. In
particular, choosing \(N = \nats\) makes \(\Sigma: \ms{Pom}_{\ms{fin}}^\omega
\to \ms{Pom}\) into a stream action w.r.t the sequential composition monoid on
finite pomsets.

We define a \textbf{program order pomset} to be a pomset with \(\mc{A}_{\ms{PO}}
= \mc{A}_w \cup \mc{A}_r \cup \{\delta\}\), where \(\mc{A}_r\) consists of
\textit{reads} of the form \(x = v\) and \(\mc{A}_w\) consists of
\textit{writes} of the form \(x := v\) for locations \(x\) and values \(v \in
\ms{Word}\). We define the \textbf{program order monad} \(\ms{PO}\;A =
\ms{Traces}\;\Sigma\;A\), where \(\Sigma\) is taken as a stream action on finite
program order pomsets. This yields an \isotopessa model with support for
concurrent read and write operations \(\ms{read}_x: \mc{I}^\varnothing_0(\mb{1},
\ms{Word})\), \(\ms{write}_x: \mc{I}^\varnothing_0(\ms{Word}, \mb{1})\) with
semantics
\(
  \dnt{\ms{read}_x} = R_x^{\ms{PO}} = \lambda (). \{(v, \{x = v\}) | v \in \ms{Word}\}
\),
\(
  \dnt{\ms{write}_x} = W_x^{\ms{PO}} =  \lambda v. \{((), \{x := v\})\}
\).
% The semantics of \(\ms{read}\) in particular give a hint as to how this model
% works: rather than tracking the state of the heap, we simply \textit{emit} a
% pomset of the events that would have be generated by a given execution (in the
% case of a read, a read event \(x = v\) whenever the read returns \(v\), and in
% the case of a write, the single write event \(x := v\) for a write of \(v\)),
% and later post-filter to obtain a set of valid executions. By doing so, our
% semantics remains compositional, allowing us to reason about each program
% fragment individually, while still allowing other program fragments to perform
% concurrent operations affecting our potential executions. 
% On that note, 
We can define the \textit{parallel execution} of two morphisms as follows:
\begin{equation}
  \begin{aligned}
  f_0 || f_1 = \lambda (a_0, a_1). 
  & \{\iota_0 ((b_0, b_1), \alpha_0 || \alpha_1) 
    \mid \iota_0 (b_i, \alpha_i) \in f_i\;a_i\} 
  \\ & \cup \{\iota_1 (\alpha_0 || \alpha_1) \mid (\iota_0 (b_0, \alpha_0) \in f_0\;a \lor \iota_0\;\alpha_0 \in f_0\;a_0) \land \iota_1 \alpha_1 \in f_1\;a_1\} 
  \\ & \cup \{\iota_1 (\alpha_0 || \alpha_1) \mid \iota_1\;\alpha_0 \in f_0\;a_0 \land \iota_0 (a_1, \alpha_1) \in f_1\;a_1\}: \ms{Set}_{\ms{PO}}(A \otimes A', B \otimes B')
  \end{aligned}
\end{equation}
It's tempting to try to use this as a tensor product of morphisms to obtain a
monoidal (rather than preomonoidal) category, but unfortunately, sliding still
fails: \((W_x^{\ms{PO}} || \ms{id}) ; (\ms{id} || W_y^{\ms{PO}})\) emits the
pomset \([x := a, y := b]\), while \(\ms{id} || W_y^{\ms{PO}}) ; (W_x^{\ms{PO}}
|| \ms{id})\) emits the pomset \([y := b, x := a]\). 
% If we did want this behaviour, we'd need to have a significantly more
% sophisticated model of composition which effectively takes nontermination into
% account, which we will leave to future work.

To graduate from sequential consistency to a genuine, if maximally simple, weak
memory model, we introduce \textit{load buffering} of write actions. We will
implement TSO ordering by buffering all our writes, in which case they are only
visible to the local thread. On the other hand, read events will first attempt
to read from the buffer, and, if there is no corresponding write in the buffer,
will read an arbitrary value, in both cases pushing an event to the global
pomset. At any point, we may choose to \textit{flush} some of the writes in
buffer. We introduce the set \(\mc{A}_b = \{(\bufloc{x} := v)\}\) of
\textit{buffer write actions}, where an action \(\bufloc{x} := v\) by a thread
denotes adding a write \(x := v\) to the thread's write buffer. We may then
define the set of TSO actions \(\mc{A}_{\ms{TSO}} = \mc{A}_{\ms{PO}} \cup
\mc{A}_b\); a pomset over this set is called a \textbf{TSO pomset}. A
buffer will be defined to be a list of write actions \(\ms{Buf} = \mc{A}_b^*\),
which we will interpret as linear pomsets over \(\mc{A}_{\ms{TSO}}\) ordered by
index (with the empty list corresponding to \(\{\delta\}\)). In particular, we
define the monad \(\ms{TSO} = \ms{StateT}\;\ms{Buf}\;(\ms{Trace}\;\Sigma)\),
where \(\Sigma\) is taken as a the stream action of finite TSO pomsets on TSO
pomsets. We can view \(\ms{PO}\) as a submonad of this monad.

Given a buffer \(\ms{Buf}\), we can define the result of reads \([\cdot]_x:
\ms{Buf} \to \ms{Word} \sqcup \{\bot\}\) from the buffer by induction to be:
\(
  (L;\{\bufloc{x} := v\})[x] = v,
\),
\(
  (L;\{\bufloc{y} := v\})[x] = L[x]
\),
\(
  [][x] = \bot
\).
% Note that writes at the \textit{end} of the buffer are prioritized, since
% later writes overwrite earlier ones! 
The semantics of reads and writes can then be given by
\begin{equation}
  \begin{aligned}
  \dnt{\ms{read}_x} = R_x^{\ms{TSO}} 
    &= \ms{pflush};(\lambda ()\;L. \{(v, L, \{x = v\}) \mid L[x] = v \lor L[x] = \bot\});\ms{pflush} \\
  \dnt{\ms{write}_x} = W_x^{\ms{TSO}}
    &= \ms{pflush};(\lambda v\;L. \{(v, (L;\{\bufloc{x} := v\}), \{x := v\})\});\ms{pflush}
  \end{aligned}
\end{equation}
where buffer flushing is implemented via the morphism 
\(
  \ms{pflush}_A = \lambda a\;L. \{(a, R, \alpha) | L = \alpha;R\}
  : \ms{Set}_{\ms{TSO}}(A, A)
\)
(called \(\ms{split}\) in \cite{sparky}). To be able to perform synchronization,
we will also need to introduce a \(\ms{fence}: \mc{I}^\varnothing_0(\mb{1},
\mb{1})\) instruction, which simply causes all actions before the fence to be
observed before any actions after the fence. For \(\ms{TSO}\), implementing this
is as simple as flushing the buffer, i.e., we define
\(
  \dnt{\ms{fence}} = \lambda ()\;L. \{((), [], L;\{\delta\})\}
\).
We can now define the parallel composition of morphisms in a "fork-join" style
as follows: we first flush the buffer completely (i.e., taking it and sticking
it at the beginning of our pomset), then execute \(f\) and \(g\) in parallel
with separate buffers, \textit{filtering out executions which completely flush
the buffer}. Since both threads end up with an empty buffer, the resulting
joined buffer is also empty, giving us the following definition:
\begin{equation}
  \begin{aligned}
    f_0 || f_1 = \lambda (a_0, a_1)\;L. 
    & \{\iota_0 ((b_0, b_1), [], L;(\alpha_0 || \alpha_1)) 
      \mid \iota_0 (b_i, [], \alpha_i) \in f_i\;a_i\;[]\} 
    \\ & \cup \{\iota_1 (\alpha_0 || \alpha_1) \mid (\iota_0 (b_0, [], \alpha_0) \in f_0\;a \lor \iota_0\;\alpha_0 \in f_0\;a_0\;[]) \land \iota_1 \alpha_1 \in f_1\;a_1\;[]\} 
    \\ & \cup \{\iota_1 (\alpha_0 || \alpha_1) \mid \iota_1\;\alpha_0 \in f_0\;a_0\;[] \land \iota_0 (a_1, [], \alpha_1) \in f_1\;a_1\;[]\}
    \end{aligned}
\end{equation}

\section{Related Work and Discussion}

For space reasons we, restrict our discussion to the most closely related work
on SSA.

Originally, \citet{kelsey-ssa-cps} showed how to embed SSA into a CPS
representation, and \citet{appel-ssa} showed informally how SSA can be seen as a
collection of mutually-tail-recursive procedures.
\citet{chakravarty-functional-ssa-2003} made Appel's observation formal, by
showing how to translate SSA into A-normal form~\cite{anf}, and used it to prove
the correctness of a constant propagation analysis. These translations were for
an untyped source and target language, and
\citet{typed-effect-ssa-rigon-torrens-vasconcellos-20} give a typed translation
from SSA into the lambda calculus, into a type-and-effect system. In contrast,
we directly track the centrality and duplicability of terms, without tracking
specific effects.

All of these papers give SSA semantics by translation -- that is, they defined
the semantics of an SSA program to be the semantics of the image of the
translation, with the idea that the simpler semantics of the functional language
can make proofs easier. The Grail language of \citet{beringer-imp-fun} makes
this idea explicit by giving both a functional and imperative semantics to a
low-level language, and showing the conditions under which they coincide.
\citet{schneider-imp-fun} prove a similar result, and further mechanise the
proof in Coq.

There have also been a number of papers which have studied the semantics of SSA
directly.

\citet{ssa-types-matsuno-ohori-06} give a type theory for what appear to be
ordinary three-address code programs. However, every well-typed program can be
placed into SSA-form by inserting $\phi$-nodes in a fully type-directed way.
This lets them model SSA without any $\phi$-nodes, lets them use the standard
semantics for three-address code.
%
\citet{hua-explicit-ssa-2010} give another type system and direct operational
semantics for SSA, and prove type safety for it.

Many operational semantics for SSA have arisen from compiler verification
efforts. \citet{barthe-compcert-ssa-2014} give an operational semantics as part
of the CompCertSSA project, and give a semantics-preserving translation from
three-address code into SSA. \citet{herklotz-gsa-2023} formalise "gated SSA" and
give semantics-preserving translations between it and ordinary SSA. Going beyond
CompCertSSA, \citet{vellum} have studied the semantics of the LLVM IR itself.

There has been much less work on denotational semantics for SSA, or directly on
its equational theory. \citet{pop-ssa-inout-2009} give an unusual denotational
model of SSA in terms of the iteration structure of a program, which they use to
better understand the loop-closing $\phi$-nodes found both in the gated SSA
representation as well as practical compilers such as GCC.
\citet{garbuzov-structural-cfg-2018} exhibit a correspondence between an
operational semantics for SSA, and an operational semantics for
call-by-push-value~\cite{cbpv}. They use the normal-form bisimulations of
\citet{lassen-bisim} to derive an equational theory for use in justifying
optimisations. Their operational semantics is effect-free, which means that it
cannot justify optimisations such as eliminating redundant writes.

Our work is the first to validate the equational theory of SSA against a weak
memory semantics. We focused on denotational models of weak memory, since they
make it easier to establish equational properties. However, we need a number of
features of the semantics which are sometimes difficult to establish in the weak
memory world. First, sequencing must be associative (i.e., $(t_1; t_2); t_3
\equiv t_1; (t_2; t_3)$).  Second, the semantics must support loops, including
the equation that a while-loop is equal to its unrolling. Finally, sums need to
be distributive (e.g., in the boolean case, $(A \otimes \mb{2}) \simeq (A +
A)$).

As a result, we focused on the TSO semantics of \citet{sparky}, which validates
the properties we need. For ARM-style concurrency, which admits load buffering,
we only found models which are "near misses": they have a subset of the
properties we need. The event structure model of
\citet{paviotti-modular-relaxed-dep-20} is formulated in terms of step-indexing,
so satisfies all of our needed properties except the loop unrolling property --
step-indexing means it is a refinement rather than an equation. The pomsets with
preconditions model of \citet{jagadeesan-pwp-20} does not exactly satisfy the
associativity of semicolon, and while the pomsets with predicate transformers
model of \citet{leaky-semicolon} does satisfy associativity, it does not yet
support loops.

We do not think it would be prohibitively difficult to extend these models to
support the required features, and we believe our results suggest that it would
be have a high payoff: those features would suffice to build a model of our IR,
giving us tight results validating the transformations compilers want against
the low-level specification of the actual machine.

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\clearpage

\appendix

\begin{figure}
  \begin{center}
    \begin{grammar}
      <\(A, B, C\)> ::= 
      \(X\)
      \;|\; \(\mathbf{1}\)
      \;|\; \(\mathbf{2}\)
      \;|\; \(A \otimes B\)

      <\(a, b, c, e\)> ::= \(x\) 
      \;|\; \(f\;a\)
      \;|\; \((a, b)\) 
      \;|\; \(()\) 
      \;|\; \(\ctt\) 
      \;|\; \(\cff\)
      \;|\; \(\letexpr{x}{a}{e}\)
      \;|\; \(\letexpr{(x, y)}{a}{e}\)
      % \;|\; \(\lbsplice{\ell}{x: A}{t}\)

      <\(s, t\)> ::= \(\lbrb{\ell}{a}\) 
      \;|\; \(\ite{e}{s}{t}\)
      \;|\; \(\letstmt{x}{a}{t}\)
      \;|\; \(\letstmt{(x, y)}{a}{t}\)
      \;|\; \(\ewhere{t}{L}\)

      <\(L\)> ::= \(\cdot\) \;|\; \(\lwbranch{\ell}{x: A}{t}, L\)

      <\(\Gamma\)> ::= \(\cdot\) \;|\; \(\Gamma, \thyp{x}{A}{q}\)

      <\(\ms{L}\)> ::= \(\cdot\) \;|\; \(\ms{L}, \lbl{\ell}[\Gamma](x: A)\)

      <\(p\)> ::= 0 \;|\; 1

      <\(q\)> ::= \(\varnothing\) 
      \;|\; \(\taff\) 
      \;|\; \(\trel\) 
      \;|\; \(\tint\)
    \end{grammar}
  \end{center}
  \caption{Grammar for \isotopessa, parametrized over a set of instructions \(f \in \mc{I}\)}
  \Description{Grammar for isotope-SSA}
  \label{fig:ssa-grammar}
\end{figure}

% \section{Proofs}

% \subsection{Weakening and Substitution}

% \begin{theorem}[Weakening] \
%   \begin{itemize}
%     \item If \(\cwk{\Gamma}{\Delta}\) and \(q \subseteq r\),
%     \(\hasty{\Delta}{p}{a}{A}{r}\), then \(\hasty{\Gamma}{p}{a}{A}{q}\)
%     \item If \(\cwk{\Gamma}{\Delta}\), \(\haslb{\Delta}{t}{\ms{L}}\), then
%     \(\haslb{\Gamma}{t}{\ms{L}}\)
%     \item If \(\lwk{\ms{L}}{\ms{K}}\), \(\haslb{\Gamma}{t}{\ms{L}}\) then
%     \(\haslb{\Gamma}{t}{\ms{K}}\)
%   \end{itemize}
% \end{theorem}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{lemma}[Substitution Splitting] 
%   If \(\csplits{\Gamma}{\Delta}{\Xi}\) and \(\issubst{\gamma}{\Theta}{\Gamma}\),
%   then there exist \(\Theta_\Delta\), \(\Theta_\Xi\) such that:
%   \begin{itemize}
%     \item \(\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}\)
%     \item \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\)
%     \item \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\)
%   \end{itemize}
% \end{lemma}

% \begin{proof} 
%   We proceed by induction on derivations \(\csplits{\Gamma}{\Delta}{\Xi}\). In
%   the \rle{split-nil} case, we have that \(\issubst{\gamma}{\cdot}{\cdot}\) by
%   \rle{subst-nil}. For all other cases, since \(\issubst{\gamma}{\Theta}{\Gamma,
%   \thyp{x}{A}{q}}\), we have by \rle{subst-cons} that
%   \(\csplits{\Theta}{\Theta_\Gamma}{\Theta_x}\),
%   \(\issubst{\gamma}{\Theta_\Gamma}{\Gamma}\),
%   \(\hasty{\Theta_\Gamma}{1}{\gamma(x)}{A}{q}\). Since
%   \(\csplits{\Gamma}{\Delta}{\Xi}\), by induction, we have that
%   \(\csplits{\Theta_\Gamma}{\Theta_\Delta}{\Theta_\Xi}\) where \(\Theta_\Delta\)
%   and \(\Theta_\Xi\) have the desired properties above.
%   \begin{itemize}
%     \item \rle{split-left}: by left split association, we have that there exists
%     \(\Theta_{\Delta, \thyp{x}{A}{q}}\), \(\csplits{\Theta}{\Theta_{\Delta,
%     \thyp{x}{A}{q}}}{\Theta_\Xi}\), \(\csplits{\Theta_{\Delta,
%     \thyp{x}{A}{q}}}{\Theta_\Delta}{\Theta_x}\). Using the fact that
%     \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\), we can conclude by
%     \rle{subst-cons} that \(\issubst{\gamma}{\Theta_{\Delta,
%     \thyp{x}{A}{q}}}{\Delta, \thyp{x}{A}{q}}\); the desired result hence
%     follows.
%     \item \rle{split-right}: by right split association, we have that there
%     exists \(\Theta_{\Xi, \thyp{x}{A}{q}}\),
%     \(\csplits{\Theta}{\Theta_\Delta}{\Theta_{\Xi, \thyp{x}{A}{q}}}\),
%     \(\csplits{\Theta_{\Xi, \thyp{x}{A}{q}}}{\Theta_\Xi}{\Theta_x}\). Using the
%     fact that \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\), we can conclude by
%     \rle{subst-cons} that \(\issubst{\gamma}{\Theta_{\Xi, \thyp{x}{A}{q}}}{\Xi,
%     \thyp{x}{A}{q}}\); the desired result hence follows.
%     \item \rle{split-dup}: since \(\tlin{A}{\trel}\), it follows by
%     \rle{subst-cons} that \(\tlin{\Theta_x}{\trel}\), and hence that
%     \(\csplits{\Theta_x}{\Theta_x}{\Theta_x}\). Therefore, by split association,
%     we have that there exist \(\Theta_{\Delta, \thyp{x}{A}{q}}\), \(\Theta_{\Xi,
%     \thyp{x}{A}{q}}\) such that \(\csplits{\Theta}{\Theta_{\Delta,
%     \thyp{x}{A}{q}}}{\Theta_{\Xi, \thyp{x}{A}{q}}}\). Using the facts that
%     \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\),
%     \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\), we can conclude by \rle{subst-cons}
%     that \(\issubst{\gamma}{\Theta_{\Delta, \thyp{x}{A}{q}}}{\Delta,
%     \thyp{x}{A}{q}}\), \(\issubst{\gamma}{\Theta_{\Xi, \thyp{x}{A}{q}}}{\Xi,
%     \thyp{x}{A}{q}}\); the desired result hence follows.
%     \item \rle{split-drop}: since \(\tlin{A}{\taff}\), it follows by
%     \rle{subst-cons} that \(\tlin{\Theta_x}{\taff}\), and hence that
%     \(\cwk{\Theta_x}{\cdot}\). Therefore, by split association,
%     \(\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}\), giving the desired result.
%   \end{itemize}
% \end{proof}

% \begin{lemma}[Substitution] 
%   Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
%   \begin{itemize}
%     \item If \(\hasty{\Gamma}{p}{a}{A}{q}\), then \(\hasty{\Theta}{p}{[\gamma]a}{A}{q}\)
%     \item If \(\haslb{\Gamma}{t}{\ms{L}}\), then there exists \(\ms{K}\) such that \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) and \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\)
%     \item If \(\lhaslb{\ms{W}}{L}{\ms{L}}\), then there exists \(\ms{K}, \ms{W}'\) such that 
%     \begin{itemize}
%       \item \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) 
%       \item \(\lbsubst{\substlbs{\gamma}{\ms{W}'}}{\ms{W}}{\ms{W}'}\) 
%       \item \(\lhaslb{\ms{W}'}{[\gamma]L}{\ms{K}}\)
%     \end{itemize}
%   \end{itemize}
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{lemma}[Label Substitution] 
%   Given \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\), 
%   \begin{itemize}
%     \item For all \(\haslb{\Gamma}{t}{\ms{L}}\), \(\haslb{\Gamma}{[\mc{L}]t}{\ms{K}}\)
%     \item For all \(\lhaslb{\ms{W}}{L}{\ms{L}}\), \(\lhaslb{\ms{W}}{[\mc{L}]L}{\ms{K}}\)
%   \end{itemize}
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \subsection{Semantic Metatheory}

% \begin{lemma}[Semantic Upgrade]
%   For all \(\Gamma, a, A\), if \(\hasty{\Gamma}{1}{a}{A}{q}\), then
%   \[\dnt{\hasty{\Gamma}{0}{a}{A}{q}} = \upg{\dnt{\hasty{\Gamma}{1}{a}{A}{q}}}{}\]
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{lemma}[Semantic Weakening] 
%   If \(\cwk{\Gamma}{\Delta}\), then 
%   \begin{itemize}
%     \item For all \(\hasty{\Delta}{p}{a}{A}{q}\),
%     \[
%       \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
%       ; \dnt{\hasty{\Delta}{p}{a}{A}{q}}
%       = \dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
%     \]
%     \item For all \(\haslb{\Delta}{t}{\ms{L}}\),
%     \[
%       \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
%       ; \dnt{\haslb{\Delta}{t}{\ms{L}}}
%       = \dnt{\haslb{\Gamma}{t}{\ms{L}}}  
%     \]
%   \end{itemize}

%   Similarly, if \(\lwk{\ms{L}}{\ms{K}}\) and \(\haslb{\Gamma}{t}{\ms{L}}\), then
%   \[
%     \dnt{\haslb{\Gamma}{t}{\ms{L}}}
%     ; \dnt{\lwk{\ms{L}}{\ms{K}}}
%     = \dnt{\haslb{\Gamma}{t}{\ms{K}}}
%   \]
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{lemma}[Substitution Coherence] \
%   \begin{itemize}
%     \item Given any two derivations \(D_1: \issubst{\gamma}{\Theta}{\Gamma}\), \(D_2: \issubst{\gamma'}{\Theta}{\Gamma}\) such that \(\forall x \in \Theta, \gamma(x) = \gamma'(x)\), we have \(\dnt{D_1} = \dnt{D_2}\)
%     \item Given any two derivations \(D_1: \lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\), \(D_2: \lbsubst{\mc{L}'}{\ms{L}}{\ms{K}}\) such that \(\forall \lbl{\ell} \in \ms{L}, \mc{L}(x) = \mc{L}'(x)\), we have \(\dnt{D_1} = \dnt{D_2}\)
%   \end{itemize}
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{lemma}[Semantic Substitution Splitting] 
%   Given \(\csplits{\Gamma}{\Delta}{\Xi}\) and \(\issubst{\gamma}{\Delta}{\Xi}\),
%   if \(\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}\) and \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\), \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\), then
%   \[
%     \dnt{\issubst{\gamma}{\Theta}{\Gamma}};\dnt{\csplits{\Gamma}{\Delta}{\Xi}}
%     = \dnt{\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}};
%       \dnt{\issubst{\gamma}{\Theta_\Delta}{\Delta}}
%       \otimes \dnt{\issubst{\gamma}{\Theta_\Xi}{\Xi}}
%   \]
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{theorem}[Semantic Substitution] 
%   Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
%   \begin{itemize}
%     \item For all \(\hasty{\Gamma}{p}{a}{A}{q}\), 
%     \[
%       \upg{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}}{p}
%       ;\dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
%       = \dnt{\hasty{\Theta}{p}{[\gamma]a}{A}{q}}
%     \]
%     \item For all 
%       \(\haslb{\Gamma}{t}{\ms{L}}\), 
%       \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\), 
%       \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\),
%     \[
%       \upg{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}}{}
%       ; \dnt{\haslb{\Gamma}{t}{\ms{L}}}
%       = \dnt{\haslb{\Theta}{[\gamma]t}{\ms{K}}} 
%       ; \dnt{\lbsubst{\substlbs{\gamma}{\ms{K}}}{\ms{K}}{\ms{L}}}
%     \]
%     \item For all 
%       \(\lhaslb{\ms{W}}{L}{\ms{L}}\), 
%       \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\),
%       \(\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}\),
%       \[
%         \dnt{\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}}
%         ; \dnt{\lhaslb{\ms{W'}}{[\gamma]L}{\ms{K}}}
%         = \dnt{\lhaslb{\ms{W}}{L}{\ms{L}}}
%         ; \dnt{\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}}
%         + \dnt{\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}}
%       \]
%   \end{itemize}
%   Similarly, given \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\),
%   \begin{itemize} 
%     \item For all \(\haslb{\Gamma}{t}{\ms{L}}\), we have
%     \[
%       \dnt{\haslb{\Gamma}{t}{\ms{L}}};\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
%       = \dnt{\haslb{\Gamma}{[\mc{L}]t}{\ms{K}}}
%     \]
%     \item For all \(\lhaslb{\ms{W}_{\ms{L}}}{L}{\ms{L}}\), we have
%     \[
%       \dnt{\lhaslb{\ms{W}_{\ms{L}}}{L}{\ms{L}}};\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
%       + \dnt{\lbsubst{\mc{L}}{\ms{W}_{\ms{L}}}{\ms{W}_{\ms{K}}}}
%       = \dnt{\lbsubst{\mc{L}}{\ms{W}_{\ms{L}}}{\ms{W}_{\ms{K}}}};\dnt{\lhaslb{\ms{W}_{\ms{K}}}{[\mc{L}]L}{\ms{K}}}
%     \]
%     where all variables in \(\ms{W}\) not in \(\ms{L}\) are fresh w.r.t \(\mc{L}\).
%   \end{itemize}
% \end{theorem}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \section{Rewriting}

% \label{apx:rewriting}

% \subsection{Syntax}

% To properly be able to reason about congruence and congruence-based
% optimistion, we need to introduce the concept of \textit{terms with holes},
% which are defined using the grammar in Figure
% \ref{fig:blocks-with-holes-grammar}. This simply extends the three main
% syntactic categories with \textit{holes} of the form \(\lhole{X}\),
% \(\lhole{T}\), \(\lhole{W}\). To \textit{type} terms with holes, we need to
% introduce the concept of a \textit{hole-context} \(H\), which is a list of holes
% annotated as corresponding to one of:
% \begin{itemize}
%   \item A term \(\tyhole{\lhole{X}}{\Delta}{p}{A}{q}\) of type \(A\), quantity
%   \(q\), and purity \(p\) using the variables in \(\Delta\). Such a hole can be
%   filled by any term \(a\) satisfying \(\hasty{\Delta}{p}{a}{A}{q}\)
%   \item A block \(\blkhole{\lhole{T}}{\Delta}{\ms{L}}\) targeting \(\ms{L}\)
%   using the variables in \(\Delta\). Such a hole can be filled by any block
%   \(t\) satisfying \(\haslb{\Delta}{t}{\ms{L}}\).
%   \item A control-flow graph \(\cfghole{\lhole{W}}{\ms{L}}{\ms{K}}\) with input
%   labels \(\ms{L}\) and output labels \(\ms{K}\). Such a hole can be filled by
%   any CFG satisfying \(\lhaslb{\ms{L}}{L}{\ms{K}}\).
% \end{itemize}

% \begin{figure}
%   \begin{grammar}
%     <\(\mhole{a}, \mhole{b}, \mhole{c}, \mhole{e}\)> ::= \(\lhole{X}\)
%     \;|\; \(x\) 
%     \;|\; \(f\;\mhole{a}\)
%     \;|\; \((\mhole{a}, \mhole{b})\) 
%     \;|\; \(()\) 
%     \;|\; \(\ctt\) 
%     \;|\; \(\cff\)
%     \;|\; \(\letexpr{x}{\mhole{a}}{\mhole{e}}\)
%     \;|\; \(\letexpr{(x, y)}{\mhole{a}}{\mhole{e}}\)
%     % \alt \(\lbsplice{\ell}{x: A}{\mhole{t}}\)

%     <\(\mhole{s}, \mhole{t}\)> ::= \(\lhole{T}\) 
%     \;|\; \(\lbrb{\ell}{\mhole{a}}\) 
%     \;|\; \(\ite{\mhole{e}}{\mhole{s}}{\mhole{t}}\)
%     \;|\; \(\letstmt{x}{\mhole{a}}{\mhole{t}}\)
%     \;|\; \(\letstmt{(x, y)}{\mhole{a}}{\mhole{t}}\)
%     \;|\; \(\ewhere{\mhole{t}}{\mhole{L}}\)

%     <\(\mhole{L}\)> ::= \(\lhole{W}\) \;|\; \(\cdot\) \;|\; \(\lwbranch{\ell}{x: A}{\mhole{t}}, \mhole{L}\)

%     <\(H\)> ::= \(\cdot\) 
%     \;|\; \(H, \tyhole{\lhole{X}}{\Gamma}{p}{A}{q}\)
%     \;|\; \(H, \blkhole{\lhole{T}}{\Gamma}{\ms{L}}\)
%     \;|\; \(H, \cfghole{\lhole{W}}{\ms{L}}{\ms{K}}\)
%   \end{grammar}
%   \caption{Grammar for \isotopessa expressions and blocks with holes}
%   \Description{Grammar for isotope-SSA blocks with holes}
%   \label{fig:blocks-with-holes-grammar}
% \end{figure}

% We introduce new typing judgements for terms, blocks, and CFGs with holes in
% Figure \ref{fig:ssa-holes-judgements}. In each case, they correspond to the
% judgement that, given a collection \(\mc{H}\) terms, blocks, and CFGs to fill
% each hole in \(H\), would satisfy the corresponding judgement without holes in
% \ref{fig:ssa-judgements}.

% \TODO{note on rewrites}

% \begin{figure}
%   \begin{center}        
%     \begingroup
%     \renewcommand{\arraystretch}{1.5}
%     \setlength{\tabcolsep}{2em}
%     \begin{tabular}{rl}
%         \multicolumn{1}{c}{Judgment} & \multicolumn{1}{c}{Meaning} \\ \hline
%         \(\isrw{\mc{H}}{I}{H}\) & \(\mc{H}\) is a rewrite from \(I\) to \(H\),
%         i.e. fills every hole in \(H\) modulo \(I\) \\
%         \(\mhasty{H}{\Gamma}{p}{\mhole{a}}{A}{q}\) & Given
%         \(\isrw{\mc{H}}{\cdot}{H}\), \(\hasty{\Gamma}{p}{[\mc{H}]\mhole{a}}{A}{q}\) \\
%         \(\mhaslb{H}{\Gamma}{\mhole{t}}{\ms{L}}\) & Given
%         \(\isrw{\mc{H}}{\cdot}{H}\), \(\haslb{\Gamma}{[\mc{H}]\mhole{t}}{\ms{L}}\) \\
%         \(\mlhaslb{H}{\ms{L}}{\mhole{L}}{\ms{K}}\) & Given
%         \(\isrw{\mc{H}}{\cdot}{H}\),
%         \(\haslb{\Gamma}{[\mc{H}]\mhole{t}}{\ms{L}}\) \\
%     \end{tabular}
%     \endgroup
%   \end{center}
%   \caption{Typing judgements for \isotopessa terms, blocks, and CFGs with holes}
%   \Description{Typing judgements for isotope-SSA terms, blocks, and CFGs with
%   holes}
%   \label{fig:ssa-holes-judgements}
% \end{figure}

% \begin{figure}
%   \begin{gather*}    
%     \prftree[r]{\rle{hole-tm}}
%       {\tyhole{\lhole{X}}{\Delta}{p}{A}{q}}
%       {\cwk{\Gamma}{\Delta}}
%       {\mhasty{H}{\Gamma}{p}{\lhole{X}}{A}{q}} \qquad
%     \prftree[r]{\rle{var}}
%       {\cwk{\Gamma}{\thyp{x}{A}{q}}}
%       {\mhasty{H}{\Gamma}{p}{x}{A}{q}} \qquad
%     \prftree[r]{\rle{app}}
%       {f \in \mc{I}_p^q(A, B)}
%       {\mhasty{H}{\Gamma}{1}{\mhole{a}}{A}{q}}
%       {\mhasty{H}{\Gamma}{p}{f\;\mhole{a}}{B}{q}} \\
%     \prftree[r]{\rle{pair}}
%       {\csplits{\Gamma}{\Delta}{\Xi}}
%       {\mhasty{H}{\Delta}{1}{\mhole{a}}{A}{q}}
%       {\mhasty{H}{\Xi}{1}{\mhole{b}}{B}{q}}
%       {\mhasty{H}{\Gamma}{p}{(\mhole{a}, \mhole{b})}{A \otimes B}{q}} \\
%     \prftree[r]{\rle{unit}}
%       {\cwk{\Gamma}{\cdot}}
%       {\mhasty{H}{\Gamma}{p}{()}{\mb{1}}{q}} \qquad
%     \prftree[r]{\rle{true}}
%       {\cwk{\Gamma}{\cdot}}
%       {\mhasty{H}{\Gamma}{p}{\ctt}{\mb{2}}{q}} \qquad
%     \prftree[r]{\rle{false}}
%       {\cwk{\Gamma}{\cdot}}
%       {\mhasty{H}{\Gamma}{p}{\cff}{\mb{2}}{q}} \\
%     \prftree[r]{\rle{let}}
%       {\csplits{\Gamma}{\Delta}{\Xi}}
%       {\mhasty{H}{\Delta, \thyp{x}{A}{}}{p}{\mhole{e}}{B}{q}}
%       {\mhasty{H}{\Xi}{1}{\mhole{a}}{A}{q}}
%       {\mhasty{H}{\Gamma}{p}{\letexpr{x}{\mhole{a}}{\mhole{e}}}{B}{q}} \qquad
%     % \prftree[r]{\rle{blk}}
%     %   {\mhaslb{H}{\Gamma}{\mhole{t}}{\llhyp{\ell}{\cdot}{A}}}
%     %   {\mhasty{H}{\Gamma}{0}{\lbsplice{\ell}{A}{\mhole{t}}}{A}{\varnothing}} 
%       \\
%     \prftree[r]{\rle{let2}}
%       {\csplits{\Gamma}{\Delta}{\Xi}}
%       {\mhasty{H}{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{p}{\mhole{e}}{C}{q}}
%       {\mhasty{H}{\Xi}{1}{\mhole{a}}{A \otimes B}{q}}
%       {\mhasty{H}{\Gamma}{p}{\letexpr{(x, y)}{\mhole{a}}{\mhole{e}}}{C}{q}}
%   \end{gather*}
%   \caption{Typing rules for \isotopessa terms with holes}
%   \Description{Typing rules for isotope-SSA terms with holes}
%   \label{fig:ssa-term-holes-typing}
% \end{figure}

% \begin{figure}
%   \begin{gather*}    
%     \prftree[r]{\rle{hole-blk}}
%       {\blkhole{\lhole{T}}{\Delta}{\ms{L}} \in H}
%       {\cwk{\Gamma}{\Delta}}
%       {\mhaslb{H}{\Gamma}{\lhole{T}}{\ms{L}}} 
%       \qquad
%     \prftree[r]{\rle{br}}
%       {\csplits{\Gamma}{\Delta}{\Xi}}
%       {\lwk{\llhyp{\ell}{\Delta}{A}}{\ms{L}}}
%       {\mhasty{H}{\Xi}{1}{\mhole{a}}{A}{\tint}}
%       {\mhaslb{H}{\Gamma}{\lbrb{\ell}{\mhole{a}}}{\ms{L}}} 
%       \\
%     \prftree[r]{\rle{ite}}
%       {\csplits{\Gamma}{\Delta}{\Xi}}
%       {\mhasty{H}{\Xi}{1}{\mhole{e}}{\mb{2}}{\tint}}
%       {\mhaslb{H}{\Delta}{\mhole{s}}{\ms{L}}}
%       {\mhaslb{H}{\Delta}{\mhole{t}}{\ms{L}}}
%       {\mhaslb{H}{\Gamma}{\ite{\mhole{e}}{\mhole{s}}{\mhole{t}}}{\ms{L}}} \\
%     \prftree[r]{\rle{let-blk}}
%       {\csplits{\Gamma}{\Delta}{\Xi}}
%       {\mhaslb{H}{\Delta, \thyp{x}{A}{}}{\mhole{t}}{\ms{L}}}
%       {\mhasty{H}{\Xi}{0}{\mhole{a}}{A}{q}}
%       {\mhaslb{H}{\Gamma}{\letstmt{x}{\mhole{a}}{\mhole{t}}}{\ms{L}}} \\
%     \prftree[r]{\rle{let2-blk}}
%       {\csplits{\Gamma}{\Delta}{\Xi}}
%       {\mhaslb{H}{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{\mhole{t}}{\ms{L}}}
%       {\mhasty{H}{\Xi}{0}{\mhole{a}}{A \otimes B}{q}}
%       {\mhaslb{H}{\Gamma}{\letstmt{(x, y)}{\mhole{a}}{\mhole{t}}}{\ms{L}}} \\
%     \prftree[r]{\rle{where}}
%       {\mhaslb{H}{\Gamma}{\mhole{t}}{\ms{L}}}
%       {\mlhaslb{H}{\ms{L}}{\mhole{L}}{\ms{K}}}
%       {\mhaslb{H}{\Gamma}{\ewhere{\mhole{t}}{\mhole{L}}}{\ms{K}}} 
%       \qquad
%     \prftree[r]{\rle{hole-cfg}}
%       {\cfghole{\lhole{W}}{\ms{L}}{\ms{K}} \in H}
%       {\lwk{\ms{M}}{\ms{K}}}
%       {\mlhaslb{\ms{L}}{H}{\lhole{W}}{\ms{K}}}
%       \\
%     \prftree[r]{\rle{nil-br}}
%       {\mlhaslb{H}{\ms{L}}{\cdot}{\ms{L}}} \qquad
%     \prftree[r]{\rle{cons-br}}
%       {\mlhaslb{H}{\ms{L}}{\mhole{L}}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}}
%       {\mhaslb{H}{\Gamma, \thyp{x}{A}{}}{\mhole{t}}{\ms{L}}}
%       {\mlhaslb{H}{\ms{L}}{\mhole{L}, \lwbranch{\ell}{x: A}{\mhole{t}}}{\ms{K}}}
%   \end{gather*}
%   \caption{Typing rules for \isotopessa blocks with holes}
%   \Description{Typing rules for isotope-SSA blocks with holes}
%   \label{fig:ssa-block-holes-typing}
% \end{figure}

% \begin{figure}
%   \begin{gather*}
%     \boxed{H: \ms{Hole} \to \ms{MExpr} \sqcup \ms{MBlock} \sqcup \ms{MCfg}}
%     \\
%     \prftree[r]{\rle{rw-nil}}
%       {\isrw{\mc{H}}{H}{\cdot}}
%     \qquad
%     \prftree[r]{\rle{rw-tm}}
%       {\isrw{\mc{H}}{I}{H}}
%       {\mhasty{I}{\Gamma}{p}{\mc{H}(\lhole{X})}{A}{q}}
%       {\isrw{\mc{H}}{I}{H, \tyhole{\lhole{X}}{\Gamma}{p}{A}{q}}}
%     \\
%     \prftree[r]{\rle{rw-blk}}
%       {\isrw{\mc{H}}{I}{H}}
%       {\mhaslb{I}{\Gamma}{\mc{H}(\lhole{T})}{\ms{L}}}
%       {\isrw{\mc{H}}{I}{H, \blkhole{\lhole{T}}{\Gamma}{\ms{L}}}}
%     \qquad
%     \prftree[r]{\rle{rw-cfg}}
%       {\isrw{\mc{H}}{I}{H}}
%       {\mlhaslb{I}{\ms{L}}{\mc{H}(\lhole{W})}{\ms{K}}}
%       {\isrw{\mc{H}}{I}{H, \cfghole{\lhole{L}}{\ms{L}}{\ms{K}}}}
%   \end{gather*}
%   \caption{Typing rules for \isotopessa rewrites}
%   \Description{Typing rules for isotope-SSA rewrites}
%   \label{fig:rewrite-typing}
% \end{figure}

% The typing rules for terms, blocks, and CFGs with holes are exactly the same as
% the corresponding rules without holes, modulo the additional rules
% \rle{hole-tm}, \rle{hole-blk}, and \rle{hole-cfg} to type the holes themselves;
% they are given in full in Figures \ref{fig:ssa-term-holes-typing} and
% \ref{fig:ssa-block-holes-typing}. This formalism allows us to state the
% \textit{syntactic rewriting} theorem, which says, in essence, that filling holes
% with valid terms yields valid terms.
% \begin{theorem}[Rewriting]
%   Given \(\isrw{\mc{H}}{H}{I}\), we have
%   \begin{itemize}
%     \item If \(\mhasty{H}{\Gamma}{p}{a}{A}{q}\), then \(\mhasty{I}{\Gamma}{p}{[\mc{H}]a}{A}{q}\)
%     \item If \(\mhaslb{H}{\Gamma}{t}{\ms{L}}\), then \(\mhaslb{I}{\Gamma}{[\mc{H}]t}{\ms{L}}\)
%   \end{itemize}
%   Furthermore, we have that 
%   \begin{itemize}
%     \item \(\mhasty{\cdot}{\Gamma}{p}{a}{A}{q} \iff \hasty{\Gamma}{p}{a}{A}{q}\)
%     \item \(\mhaslb{\cdot}{\Gamma}{t}{\ms{L}} \iff \haslb{\Gamma}{t}{\ms{L}}\)
%   \end{itemize}
% \end{theorem}

% \subsection{Semantics}

% We can interpret a hole context semantically as the \textit{set} of all possible
% tuples of morphisms which would typecheck in place of the holes in the context,
% as specified in Figure \ref{fig:holes-semantics}.
% \begin{figure}
%   \begin{align*}
%       \\
%     \dnt{H, \tyhole{\lhole{X}}{\Gamma}{p}{A}{q}}
%       &= \dnt{H} \times \mc{C}_p^q(\dnt{\Gamma}, \dnt{A}) 
%       \\
%     \dnt{H, \blkhole{\lhole{T}}{\Gamma}{\ms{L}}}
%       &= \dnt{H} \times \mc{C}_0(\dnt{\Gamma}, \dnt{\ms{L}}) 
%       \\
%     \dnt{H, \cfghole{\lhole{W}}{\ms{L}}{\ms{K}}}
%       &= \dnt{H} \times \mc{C}_0(\dnt{\ms{L}}, \dnt{\ms{K}} + \dnt{\ms{L}}) 
%       \\
%     \dnt{\cdot} 
%       &= \mb{1}
%   \end{align*}
%   \caption{Semantics for \isotopessa holes}
%   \Description{Semantics for isotope-SSA holes}
%   \label{fig:holes-semantics}
% \end{figure}
% We can now quite easily give semantics for terms, blocks, and CFGs with holes
% which simply correspond to "filling in" any holes with the appropriate morphism
% from the hole context, with all other rules interpreted identically to their
% regular variants: this is done in Figures \ref{fig:ssa-term-holes-semantics} and
% \ref{fig:ssa-block-holes-semantics}.

% \begin{figure}
%   \begin{gather*}
%     \boxed{
%       \dnt{\mhasty{H}{\Gamma}{p}{\mhole{a}}{A}{q}}
%       : \dnt{H} \to \mc{C}_p^q(\dnt{\Gamma}, \dnt{A})}
%       \\
%     \sorry
%   \end{gather*}
%   \caption{Semantics for \isotopessa terms with holes}
%   \Description{Semantics for isotope-SSA terms with holes}
%   \label{fig:ssa-term-holes-semantics}
% \end{figure}

% \begin{figure}
%   \begin{gather*}
%     \boxed{
%       \dnt{\mhaslb{H}{\Gamma}{\mhole{t}}{\ms{L}}}
%       : \dnt{H} \to \mc{C}_0^\varnothing(\dnt{\Gamma}, \dnt{\ms{L}})}
%       \\
%     \sorry
%       \\  
%     \boxed{
%       \dnt{\mlhaslb{H}{\ms{L}}{\mhole{L}}{\ms{K}}}
%       : \dnt{H} \to \mc{C}_0^\varnothing(\dnt{\ms{L}}, \dnt{\ms{K}} + \dnt{\ms{L}})}
%       \\
%     \sorry
%   \end{gather*}
%   \caption{Semantics for \isotopessa blocks with holes}
%   \Description{Semantics for isotope-SSA blocks with holes}
%   \label{fig:ssa-block-holes-semantics}
% \end{figure}

% \begin{figure}
%   \begin{gather*}
%     \boxed{
%       \dnt{\isrw{\mc{H}}{I}{H}}
%       : \dnt{I} \to \dnt{H} 
%     }
%     \\
%     \dnt{\isrw{\mc{H}}{I}{\cdot}}\;i = ()
%     \qquad
%     \dnt{\isrw{\mc{H}}{I}{H, \tyhole{\lhole{X}}{\Gamma}{p}{A}{q}}}\;i
%     = (\dnt{\isrw{\mc{H}}{I}{H}}\;i, \dnt{\mhasty{I}{\Gamma}{p}{\mc{H}(\lhole{X})}{A}{q}}\;i)
%     \\
%     \dnt{\isrw{\mc{H}}{I}{H, \blkhole{\lhole{T}}{\Gamma}{\ms{L}}}}
%     = (\dnt{\isrw{\mc{H}}{I}{H}}\;i, \dnt{\mhaslb{I}{\Gamma}{\mc{H}(\lhole{T})}{\ms{L}}}\;i)
%     \\
%     \dnt{\isrw{\mc{H}}{I}{H, \cfghole{\lhole{L}}{\ms{L}}{\ms{K}}}}
%     = (\dnt{\isrw{\mc{H}}{I}{H}}\;i, \dnt{\mlhaslb{I}{\ms{L}}{\mc{H}(\lhole{W})}{\ms{K}}}\;i)
%   \end{gather*}
%   \caption{Semantics for \isotopessa rewrites}
%   \Description{Semantics for isotope-SSA rewrites}
%   \label{fig:rewrite-semantics}
% \end{figure}

% % \begin{lemma}[Rewriting Coherence] \
% %   \begin{itemize}
% %     \item Given any two derivations \(D_1: \mhasty{H}{\Gamma}{p}{a}{A}{q}\), \(D_2: \mhasty{\Gamma}{p}{H}{a}{A}{r}\), \(\dnt{D_1} = \dnt{D_2}\)
% %     \item Given any two derivations \(D_1, D_2: \mhaslb{H}{\Gamma}{t}{\ms{L}}\), \(\dnt{D_1} = \dnt{D_2}\)
% %     \item Given any two derivations \(D_1, D_2: \isrw{\mc{H}}{I}{H}\), \(\dnt{D_1} = \dnt{D_2}\)
% %   \end{itemize}
% % \end{lemma}

% \begin{theorem}[Semantic Rewriting]
%   Given \(\isrw{\mc{H}}{I}{H}\), we have that
%   \begin{itemize}
%     \item For all \(\mhasty{H}{\Gamma}{p}{a}{A}{q}\), \(m \in \dnt{I}\), we have
%     \[
%       \dnt{\mhasty{I}{\Gamma}{p}{[\mc{H}]a}{A}{q}}(m)
%       = \dnt{\mhasty{H}{\Gamma}{p}{a}{A}{q}}(\dnt{\isrw{\mc{H}}{I}{H}}(m))
%     \]
%     \item For all \(\mhaslb{H}{\Gamma}{t}{\ms{L}}\), \(m \in \dnt{I}\), we have
%     \[
%       \dnt{\mhaslb{I}{\Gamma}{[\mc{H}]t}{\ms{L}}}(m)
%       = \dnt{\mhaslb{H}{\Gamma}{t}{\ms{L}}}(\dnt{\isrw{\mc{H}}{I}{H}}(m))
%     \]
%   \end{itemize}

%   Furthermore, we have that
%   \[
%     \dnt{\mhasty{\cdot}{\Gamma}{p}{a}{A}{q}}() = \dnt{\hasty{\Gamma}{p}{a}{A}{q}}
%     \qquad
%     \dnt{\mhaslb{\cdot}{\Gamma}{t}{\ms{L}}}() = \dnt{\haslb{\Gamma}{t}{\ms{L}}}
%   \]
% \end{theorem}

% We can further use this example to get a substructural \textit{monoidal} category by considering a primitive form of separation logic: let's introduce a grammar of predicates \(\varphi, \psi ::= \top, \ms{S}, \bot\) for programs which cannot access state, can access state, and are invalid, respectively. The separating conjunction operator on these predicates is then given by
% \begin{equation}
%   \top * \varphi = \varphi * \top = \varphi
%   \qquad
%   \bot * \varphi = \varphi * \bot = \bot
%   \qquad
%   \ms{S} * \ms{S} = \bot
% \end{equation}
% We'll then define a category \(\ms{SRel}\) with objects of the form \((A, \varphi)\) and morphisms
% \begin{equation}
%   \begin{gathered}
%     \boxed{
%       \ms{SRel}((A, \varphi), (B, \psi)) 
%       \subseteq A \to \ms{S} \to \mc{P}(A \times S) 
%     }
%     \\
%     \begin{aligned}
%       \ms{SRel}((A, \top), (B, \top)) &= \{f \in A \to S \to \mc{P}(A \times S) \mid \exists g. f = \lambda a, s. g(a) \times \{s\}\} \\
%       \ms{SRel}((A, \ms{S}), (B, \ms{S})) &= A \to S \to \mc{P}(A \times S) \\
%       \ms{SRel}((A, \varphi), (B, \psi)) &= \{\lambda a, s. \varnothing\} \quad \text{otherwise}
%     \end{aligned}
%   \end{gathered}
% \end{equation}
% Note that it is precisely the \textit{central} morphisms \(\ms{C}_0(A, B)\) which lie in \(\ms{SRel}((A, \top), (B, \top))\). We can then define the tensor product on objects as \((A, \varphi) \otimes (B, \psi) = (A \otimes B, \varphi * \psi)\), and, for \(f \in \mathsf{SRel}((A, \varphi), (B, \psi))\), and tensor product functors
% \begin{equation}
%   \begin{aligned}
%   f \otimes (C, \phi) &= f \otimes C,
%   & (C, \phi) \otimes f &= C \otimes f
%   && \text{if}\; \varphi * \phi \neq \bot, \psi * \phi \neq \bot \\
%   f \otimes (C, \phi) &= \lambda a, s. \varnothing,
%   & (C, \phi) \otimes f &=  \lambda a, s. \varnothing
%   && \text{otherwise}
%   \end{aligned} 
% \end{equation}
% Lifting associators, unitors, and symmetries from \(\mc{C}_0\) in the obvious manner (setting them to \(\lambda a, s. \varnothing\) where necessary), we can verify that \(\ms{SRel}\) is a \textit{premonoidal} category. we can see that it is in fact \textit{monoidal} by nothing that, for any \(f, g\), \(f \ltimes g = f \rtimes g\) since either \(f\) is central, \(g\) is central, or their product is the empty morphism \(\lambda a, s. \varnothing\). Hence, to get a \textit{substructural} monoidal category, we simply need to define:
% \begin{equation}
%   \begin{gathered}
%   \ms{Aff}(\ms{SRel}) = \ms{Rel}(\ms{SRel}) = \{(A, \top) | A \in \ms{Set}\} \neq |\ms{SRel}|
%   \\
%   \begin{aligned}
%     \ms{SRel}^\varnothing((A, \phi), (B, \psi))
%     &= \ms{SRel}((A, \phi), (B, \psi)) \\
%     \ms{SRel}^{\{\ms{a}\}}((A, \phi), (B, \psi)) 
%     &= \{f \in \ms{SRel}((A, \phi), (B, \psi)) \mid \forall a, s. |f(a, s)| \geq 1 \lor \phi \neq \psi \lor \phi = \bot \} \\
%     \ms{SRel}^{\{\ms{r}\}}((A, \phi), (B, \psi)) 
%     &= \{f \in \ms{SRel}((A, \phi), (B, \psi)) \mid \forall a, s. |f(a, s)| \leq 1 \lor \phi \neq \psi \lor \phi = \bot \} \\
%     \implies \ms{SRel}^{\{\ms{a}, \ms{r}\}}((A, \phi), (B, \psi)) 
%     &= \{f \in \ms{SRel}((A, \phi), (B, \psi)) \mid \forall a, s. |f(a, s)| = 1 \lor \phi \neq \psi \lor \phi = \bot \} \\
%   \end{aligned}
%   \end{gathered}
% \end{equation}
% We can verify that the associators, unitors, and symmetry lie in \(\ms{SRel}^{\{\ms{a}, \ms{r}\}}\) and that all other desired axioms are satisfied.

% \TODO{relate separation logic trick to \cite{promonad}, \cite{linear-state-usage}, \cite{mellies-ftrs}}

% \TODO{clean up, and generalize to arbitrary premonoidal categories (w/ initial object? how to deal w/ coproducts?)}

\end{document}
\endinput